---
title: "GAM and Figures"
author: "Natasha Ryan"
output: 
  html_document: default
editor_options: 
  chunk_output_type: inline
---

This RMD reads objects from the global environment, generated immediately before in Wavelet_LowLight.Rmd for generalized additive modelling and generating figures for Natasha Ryan's undergraduate honours thesis at Mount Allison University. 

Two list objects missing???
# Line 411 'diff.sstate' not found;  not generated by running Wavelet_LowLight.Rmd firstly
# Line 533 'temp.sstate' not found;  not generated by running Wavelet_LowLight.Rmd firstly

Should they be 'diff.list' and 'temp.list'?
Code does not run with those names?

Introduction

This script fits Generalized Additive Models (GAMs) using mgcv::gam combined with possibGAM (from purrr::possibly), allowing model fitting without halting execution due to errors. Smooth terms (s) and interaction effects (ti) are used to capture nonlinear relationships.  gam.check is used to run model diagnostics guiding basis dimension (k) selection.

GamParaMod : is the response variable input for the GAM. It is assigned at the start of each GAM chunk.
s(CalcSal) and s(par_ue): main effects
ti(CalcSal, par_ue, k = 3): Tensor product smooth for modeling interaction between CalcSal and par_ue.

Unlike te(), ti() includes only the interaction term (not the main effects), which is appropriate here because main effects are already modeled separately using s(CalcSal) and s(par_ue).
This avoids redundancy and helps reduce overfitting.
The k parameter limits the flexibility of the interaction term. High k = more flexibility for more complex patterns, low k = simple curves.



# Setup
```{r load libraries}
library(tidyverse)
library(metR)
library(mgcv)
library(ggpubr)
library(knitr)
library(broom)
library(patchwork)
library(cowplot)  # for get_legend() necessary for combining plots and use ggdraw
library(grid) # for textGrob necessary for combining axis

# library(commonR)  Natasha defined her own theme_pres(), used here and in other .Rmd;  sloppy handling of functions, variables and environments
```
```{r define paths}

PlotsOut <- file.path("..", "Figures")
```

```{r create ggplot theme}
theme_thesis <- function() {
  ggplot2::theme_bw() +
  ggplot2::theme(text = element_text(family = "sans"), 
        axis.text=element_text(size=11), axis.title=element_text(size=13),
        legend.text=element_text(size=11), legend.title=element_text(size=13),
        strip.background = element_rect(colour="black", fill=NA),
        strip.text = element_text(size=11), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.title.y.right = element_text(size = 12), 
        axis.title.x.top = element_text(size = 12)
  )}
```


# Representative Fluorescence Oscillations

```{r warning=FALSE}
Cp_data <- readRDS("../Data/CleanData/LOW_Chlamydomonas_priscuii_SolFitsTrim.Rds") %>%
  mutate(SState = ifelse(SState == 4, 0, SState), 
         SState = factor(SState, levels = c(1, 2, 3, 0))) %>%
  ungroup()

Cp_data <- Cp_data %>%
    group_by(PulseSpace_s) %>%
    mutate(Approx_Light_Level = mean(as.numeric(Light_Level))) %>%
    mutate(Approx_Light_Level = round(Approx_Light_Level, 3)) %>%
    ungroup()
```


```{r warning=FALSE}
Cp_data %>%
  filter(Filename %in% c("20230918_MaPo1001_12C_1s", "20230918_MaPo1001_12C_4s", "20230919_MaPo1004_12C_16s",
                      "20230918_MaPo1001_4C_1s",  "20230918_MaPo1001_4C_4s", "20230918_MaPo1001_4C_16s")) %>%
  ggplot() +
  geom_point(aes(x=Flashnumber, y=FvFmrunnorm, col=as.factor(SState))) +
  scale_color_manual(values = c(rcartocolor::carto_pal(n=4,"Temps"))) +
  labs(x= "Flash Number", 
       y=expression("Normalized Chlorophyll Fluorescence (F"[v]*" / F"[m]*")"),
       col= "Majority S-State") +
  ggh4x::facet_nested(rows=vars(Approx_Light_Level, PulseSpace_s), cols=vars(Temp_C), 
                      as.table = FALSE) +
  scale_y_continuous(sec.axis=sec_axis(~ . , 
                     name = bquote(Effective ~ Light ~ Level ~ (µmol ~ photons ~ m^-2 ~ s^-1)), 
                     breaks = NULL,labels = NULL)) +
  scale_x_continuous(breaks=c(4,8,12,16,20,24,28,32),
                     sec.axis=sec_axis(~ . , 
                                       name = sym("Measurement Temperature (°C)"), 
                                       breaks = NULL, labels = NULL)) +
  theme_thesis() 

# ggsave("representativeoscillations.png", width = 9, height = 5, dpi = 900,
#       plot=last_plot(),
#       path="../Output")
```



# FFT
```{r}
# cols.2 <- c("#3F3C70","#40538C", "#4C99A0", "#96D2A6")
# 
# Cp_data %>%
#   filter(CultureID == "MaPo1001", 
#          ObsDate == "2023-09-18", 
#          Temp_C == 4, 
#          PulseSpace_s == 1) %>%
#   ggplot() +
#   geom_point(aes(x=Flashnumber, y=FvFmrunnorm, col=as.factor(SState)), size=1.5) +
#   scale_color_manual(values = c(cols.2)) +
#   labs(x= "Flash Number", 
#        y=expression("Normalized Chlorophyll Fluorescence (F"[v]*" / F"[m]*")"),
#        col= "Majority S-State") +
#   scale_x_continuous(breaks=c(4,8,12,16,20,24,28,32)) +
#   #theme_pres() + 
#   theme(legend.position = "none")

# ggsave("wavelet_before.png", width = 4, height = 4, dpi = 900,
#        plot=last_plot(),
#        path="../Output")
```

##Chunk returns error
```{r}
# wave_data <- Cp_wavelets[["Cp_1_4_0.77084"]]
# 
# 
# WaveletComp::reconstruct(wave_data, only.sig=FALSE,
#             plot.waves = TRUE, plot.rec = FALSE,
#             spec.time.axis = list(at = c(0,4,8,12,16,20,24,28,32)),
#             timelab = "Flash Number",
#             verbose = FALSE)
```

##Chunk returns error, no file 'LOW_Wavelets.Rds' in directory

# Wavelet Periodicity 

```{r}
# wavelets <- readRDS("../Data/ProcessedData/LOW_Wavelets.Rds") 
# specieslists <- split(wavelets, sub("_.*", "_wavelets", names(wavelets)))
# list2env(specieslists, envir = .GlobalEnv)
# 
# FcG0_wavelets <- Fc_wavelets[grepl(paste("_0$", collapse = "|"), names(Fc_wavelets))]
# FcG6_wavelets <- Fc_wavelets[grepl(paste("_6$", collapse = "|"), names(Fc_wavelets))]
```

##No wavelet data
```{r periodicity function}
# plot.wt.grid <- function(wavelets) {
#   result_list <- lapply(names(wavelets), function(condition_name) {
#     condition <- wavelets[[condition_name]]
#     Power.avg <- condition$Power.avg
#     Power.avg.pval <- condition$Power.avg.pval
#     Period <- condition$Period
#     Condition <- rep(condition_name, length(Power.avg))  # Repeat condition name for each row
#     df <- data.frame(Power.avg, Power.avg.pval, Period, Condition) %>%
#       separate(Condition, into = c("Strain", "PulseSpace_s", "Temp_C", "LightLevel"), sep = "_") %>%
#       mutate(PulseSpace_s = factor(PulseSpace_s, levels = c("1", "2","4","8","16")),
#              Temp_C = factor(Temp_C, levels = 
#                            c("0","2","4","6","8","10","12","14","16","18","20","22","24","26","28"))) %>%
#       mutate(Power.avg.pval = ifelse(Power.avg.pval <= 0.05, 0, 1))  %>%
#     
#      return(df)
#   })
# 
#   result_df <- bind_rows(result_list) %>%
#     group_by(PulseSpace_s) %>%
#     mutate(Light_Level = mean(as.numeric(LightLevel))) %>%
#     mutate(Light_Level = round(Light_Level, 3)) %>%
#     ungroup()
#   
#   plot <- result_df %>%
#     ggplot()+ 
#     geom_path(aes(x= Power.avg, y= Period, col=Power.avg.pval), linewidth=0.8) +
#     scale_color_gradient(low="#4AC0AD", high="black", breaks= c(0.05, 1), 
#                          labels= c("≤ 0.05", "> 0.05"), guide="legend") +
#     labs(x= "Average Wavelet Power", colour= "P-Value") +
#     ggh4x::facet_nested(rows=vars(Light_Level, PulseSpace_s), cols = vars(Temp_C),
#                         as.table = FALSE) +
#     scale_y_continuous(limits= c(0,10),
#                        breaks= c(0,4,8), 
#                  sec.axis=sec_axis(~ . , 
#                              name = bquote(Effective ~ Light ~ Level ~ (µmol ~ photons ~ m^-2 ~ s^-1)),
#                                    breaks = NULL,labels = NULL)) +
#     scale_x_continuous(limits= c(0,1), 
#                        breaks=c(0, 0.4, 0.8),
#     sec.axis=sec_axis(~ . , name = sym("Measurement Temperature (°C)"), breaks = NULL, labels = NULL)) +
#     theme_thesis() 
#   return(plot)
# }
```

##No wavelet data
```{r Periodicity Test, warning=FALSE}
# plot.wt.grid(Ci_wavelets)
# plot.wt.grid(Cm_wavelets)
# plot.wt.grid(Cr_wavelets)
# plot.wt.grid(Cv_wavelets)
# plot.wt.grid(Tp_wavelets)
# plot.wt.grid(FcG6_wavelets)
# plot.wt.grid(FcG0_wavelets)
# 
# plot.wt.grid(Cr_wavelets)
#  # ggsave("temperate_wtavg.png", width = 9.5, height = 5, dpi = 900,
#  #      plot=last_plot(),
#  #      path="../Output")
# 
# plot.wt.grid(Cp_wavelets)
#  # ggsave("polar_wtavg.png", width = 9.5, height = 5, dpi = 900,
#  #      plot=last_plot(),
#  #      path="../Output")
```

# Reconstruction 

```{r}
# reconstructions <- readRDS(file = file.path("..", "Data", "ProcessedData", "LOW_Reconstructions.Rds" )) 
# reconstruction.list <- split(reconstructions, sub("_.*", "_reconstruction", names(reconstructions)))
# list2env(reconstruction.list, envir = .GlobalEnv)
# 
# FcG0_reconstruction <- Fc_reconstruction[grepl(paste("_0$", collapse = "|"), names(Fc_reconstruction))]
# FcG6_reconstruction <- Fc_reconstruction[grepl(paste("_6$", collapse = "|"), names(Fc_reconstruction))]
```


```{r}
# plot.reconstruction <- function(reconstructed) {
#   result_list <- Map(function(condition_name, condition_data) {
#     series <- as.data.frame(condition_data$series)
#     series <- mutate(series, Condition = condition_name) %>%
#       mutate(Flashnumber = row_number())  %>%
#       separate(Condition, into = c("Strain", "PulseSpace_s", "Temp_C", "LightLevel"),
#                 sep = "([\\/\\_])") %>%
#        mutate(PulseSpace_s = factor(PulseSpace_s, levels = c("1","2","4","8","16")),
#              Temp_C = factor(Temp_C, levels = 
#                              c("0","2","4","6","8","10","12","14","16","18","20","22","24","26","28")))
#     return(series)
#   }, names(reconstructed), reconstructed)
#   
#   result_df <- bind_rows(result_list) %>%
#     group_by(PulseSpace_s) %>%
#     mutate(Light_Level = mean(as.numeric(LightLevel))) %>%
#     mutate(Light_Level = round(Light_Level, 3)) %>%
#     ungroup()
#   
#   plot <- result_df %>%
#     ggplot()+
#     geom_line(aes(x=Flashnumber, y=FvFmrunnorm, col = "Original")) +
#     geom_line(aes(x=Flashnumber, y=FvFmrunnorm.r, col= "Reconstructed"))  +
#     scale_colour_manual(values=c("black", "deepskyblue3"), 
#                         labels = c("Original", expression(paste("Reconstructed (", alpha, " = 0.05)")))) +
#     labs(x= "Flash Number", 
#          y=expression("Normalized Chlorophyll Fluorescence (F"[v]*" / F"[m]*")"),
#          col= "Wave Function") +
#     ggh4x::facet_nested(rows=vars(Light_Level, PulseSpace_s), cols=vars(Temp_C), as.table = FALSE) +
#     scale_y_continuous(sec.axis=sec_axis(~ . , 
#                            name = bquote(Effective ~ Light ~ Level ~ (µmol ~ photons ~ m^-2 ~ s^-1)), 
#                                          breaks = NULL,labels = NULL)) +
#     scale_x_continuous(sec.axis=sec_axis(~ . , name = sym("Measurement Temperature (°C)"), 
#                                          breaks = NULL, labels = NULL)) +
#     theme_thesis()
#   
#   return(plot)
# }
```


```{r warning=FALSE}
# plot.reconstruction(Ci_reconstruction)
# plot.reconstruction(Cm_reconstruction)
# plot.reconstruction(Cr_reconstruction)
# plot.reconstruction(Cv_reconstruction)
# plot.reconstruction(Tp_reconstruction)
# plot.reconstruction(FcG0_reconstruction)
# plot.reconstruction(FcG6_reconstruction)
# 
# 
# plot.reconstruction(Cp_reconstruction)
 # ggsave("representative_reconstruction.png", width = 10, height = 6, dpi = 900,
 #      plot=last_plot(),
 #      path="../Output")
```


# Damping Across Taxa

```{r import & wrangle data}
sstate_damping <- readRDS(file = file.path("..", "Data", "ProcessedData", "LOW_SState_Damping.Rds" )) %>%
  mutate(Strain = factor(Strain), 
  ThermalTolerance = case_when(
     Strain %in% c("Fragilariopsis_cylindrus", "Chlamydomonas_ICEMDV","Chlamydomonas_priscuii",
                   "Chlamydomonas_malina") ~ "Polar",
     Strain %in% c("Chlamydomonas_reinhardtii", "Chlorella_vulgaris",
                   "Thalassiosira_pseudonana") ~ "Temperate"),  
  Type = case_when(
     Strain %in% c("Fragilariopsis_cylindrus","Thalassiosira_pseudonana") ~ "Diatom", 
     Strain %in% c("Chlamydomonas_ICEMDV","Chlamydomonas_priscuii", "Chlamydomonas_malina",
                   "Chlamydomonas_reinhardtii", "Chlorella_vulgaris") ~ "Green Algae"),
  l.strain = case_when(
    Strain == "Chlamydomonas_ICEMDV" ~ "~italic('C. ICEMDV')", 
    Strain == "Chlamydomonas_malina" ~"~italic('C. malina')",
    Strain == "Chlamydomonas_priscuii" ~ "~italic('C. priscuii')",
    Strain == "Chlamydomonas_reinhardtii" ~ "~italic('C. reinhardtii')",
    Strain == "Chlorella_vulgaris" ~ "~italic('C. vulgaris')",
    Strain == "Fragilariopsis_cylindrus" ~ "~italic('F. cylindrus')", 
    Strain == "Thalassiosira_pseudonana" ~ "~italic('T. pseudonana')"),
  l.temp = paste0("Grown~at~", GrowthTemp_C,"*degree*C"),
  FacetLabel_temp = paste0("grown~at~", GrowthTemp_C,"*degree*C"),
  )

# diff.list <- doBy::splitBy(sstate_damping, formula = "Strain")
# temp.list <- doBy::splitBy(sstate_damping, formula = c("Strain", "GrowthTemp_C"))
```


```{r Plot Setup}
# # Vertical Lines for Growth Temperatures 
# diatom.vlines <- data.frame(
#   l.strain = c("~italic('F. cylindrus')", "~italic('F. cylindrus')", "~italic('T. pseudonana')"),
#   l.temp = c("Grown~at~0*degree*C", "Grown~at~6*degree*C","Grown~at~22*degree*C"), 
#   GrowthTemp_C = c(0.1, 6, 22))
# 
# algae.vlines <- data.frame(
#   l.strain = c("~italic('C. ICEMDV')",  "~italic('C. malina')", "~italic('C. priscuii')", 
#              "~italic('C. vulgaris')", "~italic('C. reinhardtii')"),
#   ThermalTolerance = c("Polar", "Polar", "Polar", "Temperate", "Temperate"), 
#   GrowthTemp_C = c(4.1, 4.1, 4.1, 22, 23.85))
# 
# # Computing flash spacing averages for secondary axes of heatmaps
# spacing.diatom <- doBy::summaryBy(Light_Level ~ PulseSpace_s, data = subset(sstate_damping, Type == "Diatom"))
# spacing.algae <- doBy::summaryBy(Light_Level ~ PulseSpace_s, data = subset(sstate_damping, Type == "Green Algae"))
```


### Significance Matrix
Figure out what to do with growth temps
Figure out how to change facet widths to scale with temperature ranges
https://stackoverflow.com/questions/52341385/how-to-automatically-adjust-the-width-of-each-facet-for-facet-wrap
https://teunbrand.github.io/ggh4x/articles/Facets.html#position-scales
```{r Plot significance matrix, warning=FALSE}
mtrx.data <- sstate_damping %>%
  mutate(DampingIndex = ifelse(DampingIndex > 0, 1, 0),
  Strain = factor(Strain, levels=c("Fragilariopsis_cylindrus", "Thalassiosira_pseudonana", "Chlamydomonas_ICEMDV", 
            "Chlamydomonas_malina", "Chlamydomonas_priscuii", "Chlorella_vulgaris", "Chlamydomonas_reinhardtii")))

mtrx.data  %>%
  ggplot() +
  geom_tile(aes(x= as.factor(Temp_C), y=as.factor(PulseSpace_s), fill=as.factor(DampingIndex))) +
  scale_fill_manual(values= c("black","skyblue"), 
                      name= bquote(atop("Significance of", "4-Step Periodicity")), 
                      labels =c("p > 0.05", "p ≤ 0.05")) +
  labs(x = "Measurement Temperature (°C)", y= "Flash Spacing (s)") +
  geom_vline(xintercept = as.factor(mtrx.data$GrowthTemp_C), colour = "green", size = 1) +
  ggh4x::facet_nested_wrap(vars(ThermalTolerance, l.strain), nrow=2, scales="free_x", 
                             labeller = label_parsed) +
  # ggh4x::force_panelsizes(cols = c(10, 8, 8, 8, 18, 16, 12), respect = TRUE, total_width = unit(12, "cm")) +
  theme_thesis()





# +
#   theme(legend.title = element_text(size=11), strip.text = element_text(face="italic"),
#         legend.position = c(0.97,0.1), legend.justification = c(1,0))

#, GrowthTemp_C add to ggh4x::facet_nested_wrap?

#free_x

 # ggsave("periodicity_mtrx.png", width = 7 , height = 5, dpi = 900,
 #        plot=last_plot(),
 #        path="../Output")
```



```{r}
mtrx.data  %>%
  filter(Strain == 	"Fragilariopsis_cylindrus") %>%
  ggplot() +
  geom_tile(aes(x= as.factor(Temp_C), y=as.factor(PulseSpace_s), fill=as.factor(DampingIndex))) +
  scale_fill_manual(values= c("black","skyblue"), 
                      name= bquote(atop("Significance of", "4-Step Periodcity")), 
                      labels =c("p > 0.05", "p ≤ 0.05")) +
  labs(x = "Measurement Temperature (°C)", 
       y= "Flash Spacing (s)") +
  facet_wrap(vars(l.temp), labeller = label_parsed) +
  theme_thesis() +
  theme(legend.title = element_text(size=11))

# ggsave("Fc_mtrx.png", width = 6 , height = 3, dpi = 900,
#        plot=last_plot(),
#        path="../Output")
```


### GAM By Deviation From Growth Temp

```{r Functions for GAM by Diff}
# predict.diff.gam <- function(diff.list) {
#   combineddata <- lapply(diff.list, function(damping){
#   gam <- mgcv::gam(DampingIndex ~ te(TempDiff_C, Light_Level, k= n_distinct(damping$TempDiff_C)), 
#                    data = damping, method = "REML")
# 
#   data <- with(damping, expand.grid( Light_Level = seq(min(Light_Level), max(Light_Level), 0.05),
#     TempDiff_C = seq(min(TempDiff_C), max(TempDiff_C), 0.1)))
#   
#   data <- data %>%
#     mutate(DampingIndex = predict(gam, newdata = data), 
#            DampingIndex = ifelse(DampingIndex < 0, 0, DampingIndex),
#            Strain = factor(unique(damping$Strain)), 
#            l.strain = unique(damping$l.strain),
#            ThermalTolerance = factor(unique(damping$ThermalTolerance))) 
#       return(data) })
#   combineddata <- do.call(bind_rows, combineddata)
#   return(combineddata)
# }
```


```{r Functions for GAM by Diff}
# predict.gam.d <- predict.diff.gam(diff.list)
```


```{r Functions for GAM by Diff}
# summary.diff.gam <- function(damping) {
#   knots <- n_distinct(damping$TempDiff_C)
#   gam <- mgcv::gam(DampingIndex ~ te(TempDiff_C, Light_Level, k=knots), data = damping, method = "REML")
#   summary <- summary(gam)
#   shapiro <- shapiro.test(resid(gam))
#   k <- mgcv::k.check(gam)
#   vis.fit <- mgcViz::getViz(gam)
#   check <- mgcViz::check.gamf0iz(vis.fit,
#                                 a.qq = list(method = "tnorm", 
#                                             a.cipoly = list(fill = "light blue")), 
#                                 a.respoi = list(size = 1), 
#                                 a.hist = list(bins = 4))
#   combined <- list(summary = summary, norm = shapiro, check.k = k, plot.assumptions = check)
#   return(combined)
# }
```


```{r Create Predicted df}
# predict.gam.d <- lapply(diff.sstate, predict.diff.gam)
# predict.gam.d <- bind_rows(predict.gam.d)
```


```{r Generate model summaries}
#summary.gam.d <- lapply(diff.sstate, summary.diff.gam)
```


##### Diatoms
```{r}
# predict.gam.d %>%
#   filter(Strain %in% c("Fragilariopsis_cylindrus", "Thalassiosira_pseudonana")) %>%
#   ggplot(aes(x=TempDiff_C, y=Light_Level, z= DampingIndex)) +
#   geom_contour_fill() + 
#   geom_contour2(color = "white", linemitre = 10, linewidth=0.25) +
#   geom_text_contour(stroke = 0.1, family = "Times New Roman", size = 3.5) +
#   geom_vline(xintercept=0, col= "white", linetype = "dashed", linewidth = 1.3) +
#   labs(x= "Deviation from Growth Temperature (°C)",
#        y= expression("Light Level ("*mu*"mol photons m"^"-2"~s^-1*")"),
#        fill= "Damping Index") +
#   scale_fill_gradientn(colors = pals::ocean.deep(13)) +
#   coord_cartesian(ylim = c(0.03, 0.53)) +
#   scale_x_continuous(breaks= c(-12,-8,-4,0,4,8), expand=c(0,0)) +
#   scale_y_continuous(expand=c(0,0), sec.axis = dup_axis(
#     breaks = c(spacing.diatom$Light_Level.mean), labels = c(spacing.diatom$PulseSpace_s),
#     name = sym("Equivalent Flash Spacing (s)"))) +
#   ggh4x::facet_nested_wrap(vars(ThermalTolerance, l.strain), ncol = 1, strip.position = "right",
#                              labeller = label_parsed)+
#   theme_thesis() +
#   theme(strip.placement = "outside",
#         panel.spacing.y = unit(4, "mm"))

# ggsave("diatoms_diff.png", width = 8.5 , height = 5.5, dpi = 900,
#        plot=last_plot(),
#        path="../Output")
```

##### Algae
```{r}
# predict.gam.d %>%
#   filter(Strain %in% c("Chlamydomonas_ICEMDV", "Chlamydomonas_malina", "Chlamydomonas_priscuii", 
#              "Chlorella_vulgaris", "Chlamydomonas_reinhardtii")) %>%
#   ggplot(aes(x=TempDiff_C, y=Light_Level, z= DampingIndex)) +
#   geom_contour_fill() + 
#   geom_contour2(color = "white", linemitre = 10, linewidth=0.25) +
#   geom_text_contour(stroke = 0.1, family = "Times New Roman", size = 3.5) +
#   geom_vline(xintercept=0, col= "white", linetype = 1, linewidth = 1.5) +  
#   geom_vline(xintercept=0, col= "black", linetype = 2, linewidth = 0.9) +  
#   labs(x= "Deviation from Growth Temperature (°C)",
#        y= expression("Light Level ("*mu*"mol photons m"^"-2"~s^-1*")"),
#        fill= "Damping Index") + 
#   scale_fill_gradientn(colors = pals::ocean.deep(13), breaks = seq(0, 13, by = 3)) +
#   coord_cartesian(ylim = c(0.03, 0.66)) +
#   scale_x_continuous(breaks= c(-12,-8,-4,0,4,8), expand=c(0,0)) +
#   scale_y_continuous(expand=c(0,0), sec.axis = dup_axis(
#     breaks = c(spacing.algae$Light_Level.mean), labels = c(spacing.algae$PulseSpace_s),
#     name = sym("Equivalent Flash Spacing (s)"))) +
#   ggh4x::facet_nested_wrap(vars(ThermalTolerance, l.strain), ncol = 1, strip.position = "right",
#                              labeller = label_parsed)+
#   theme_thesis() +
#   theme(strip.placement = "outside",
#         panel.spacing.y = unit(4, "mm"))


# ggsave("algae_diff.png", width = 8 , height = 10.5, dpi = 900,
#        plot=last_plot(),
#        path="../Output")
```


### GAMs By Measurement Temp 

```{r Functions for GAM by Temp}
# predict.temp.gam <- function(damping) {
#   
#   knots <- n_distinct(damping$Temp_C)
#   
#   gam <- mgcv::gam(DampingIndex ~ te(Temp_C, Light_Level, k=knots), data = damping, method = "REML")
# 
#   data <- with(damping, expand.grid(
#     Light_Level = seq(min(Light_Level), max(Light_Level), 0.05),
#     Temp_C = seq(min(Temp_C), max(Temp_C), 0.1)
#   ))
#   
#   data$DampingIndex <- predict(gam, newdata = data)
#   
#   data <- data %>%
#     mutate(DampingIndex = ifelse(DampingIndex < 0, 0, DampingIndex)) %>%
#     mutate(Strain = factor(unique(damping$Strain)), 
#            l.strain = unique(damping$l.strain),
#            GrowthTemp_C = factor(unique(damping$GrowthTemp_C)), 
#            ThermalTolerance = factor(unique(damping$ThermalTolerance))) 
#   
#   return(data) 
# }
# 
# summary.temp.gam <- function(damping) {
#   knots <- n_distinct(damping$Temp_C)
#   gam <- mgcv::gam(DampingIndex ~ te(Temp_C, Light_Level, k=knots), data = damping, method = "REML")
#   summary <- summary(gam)
#   shapiro <- shapiro.test(resid(gam))
#   k <- mgcv::k.check(gam)
#   vis.fit <- mgcViz::getViz(gam)
#   check <- mgcViz::check.gamViz(vis.fit,
#                                 a.qq = list(method = "tnorm", 
#                                             a.cipoly = list(fill = "light blue")), 
#                                 a.respoi = list(size = 1), 
#                                 a.hist = list(bins = 4))
#   combined <- list(summary = summary, norm = shapiro, check.k = k, plot.assumptions = check)
#   return(combined)
# }
```

XXXXSubstitute Mirelle GAM code
Run on 'sstate_damping' dataframe
Response variable 'DampingIndex'
Predictors are 'Temp_C' and 'Light_Level'
Grouping variables are 'Strain' and GrowthTemp_C'

```{r DampingIndex GAM}

GamParaMod = "DampingIndex" 

#https://stackoverflow.com/questions/49570452/is-it-possible-to-specify-lower-bound-in-response-variable-during-smooth-with-ga 
possibGAM = possibly(.f = mgcv::gam, otherwise = NULL)
possibPRED = possibly(.f = mgcv::predict.gam, otherwise = NULL)

# build a grid that includes measured temperatures (0,2,4,6,8,10,12,14,16,18,20,22,24,26,28)
# Define important Temp_C values


# Build Temp_C vector that includes both the full sequence and key values
Temp_C_vals <- seq(0, 28, by = 0.05)

# Same for Light_Level if needed (optional: if you want specific values there too)
Light_vals <- seq(0.02, 1.0, by = 0.02)

# Build the grid
PAR_Temp_Model <- expand_grid(
  Light_Level = Light_vals,
  Temp_C = Temp_C_vals
) %>% 
  as_tibble()


# this generates a grid but the exact measured temperatures are excluded therefore may show a gap at axis limits (example, for x-axis 4 to 12 has Temp data starting at 4.13 so white space from 4.13 to 4.0 on plot)
# PAR_Temp_Model <- expand_grid(
#   Light_Level = seq(from=min(0.02),
#               to=max(1.0),
#               length.out = 150),
#   Temp_C = seq(from=min(0),
#               to=max(28),
#               length.out = 150)
#   ) %>% as_tibble()


summary(PAR_Temp_Model) 


DampingIndex_Gam <- sstate_damping %>% 
   ungroup() %>%
  group_by(Strain, GrowthTemp_C, ThermalTolerance, Type, l.strain, l.temp, FacetLabel_temp) %>%
  nest() %>%
  mutate(
    GamModel = map(data, ~possibGAM(!!sym(GamParaMod) ~  s(Light_Level, k = 3) + s(Temp_C, k = 3) + ti(Temp_C, Light_Level, k = 3), data = .x,
                                    family = "gaussian"(link = "log"), 
                                    method = "REML")),
    predicted = map(GamModel, ~possibPRED(.x, newdata = PAR_Temp_Model, type = "response", se.fit = TRUE)),
    tidied = map(GamModel, ~tidy(.x)),
    augmented = map(GamModel, ~augment(.x)),
    glance = map(GamModel, ~glance(.x))
  ) %>% 
  ungroup()

DampingIndex_Pred <- DampingIndex_Gam %>%
  dplyr::select(predicted, Strain, GrowthTemp_C, ThermalTolerance, Type, l.strain, l.temp, FacetLabel_temp) %>%
  dplyr::filter(predicted != "NULL") %>%
  rowwise() %>%
  mutate(db = list(cbind(predicted, PAR_Temp_Model))) %>%
  dplyr::select(-predicted) %>%
  unnest(col = db) 

# GAM model results into table
kable(DampingIndex_Gam %>% 
  dplyr::select(Strain, tidied) %>% 
  unnest(tidied), 
  digits = 3)

```


```{r check DampingIndex GAM}

#check that EDF (Effective Degrees of Freedom)
# Note: Low p-value in gam.check() for a smooth term means k may be too small
# not that the term is statistically significant, but that it may be under fitting.
purrr::walk(DampingIndex_Gam$GamModel, gam.check)

# GAM CHECK RESULTS

# in s(Light_Level, k = 3) , k is too low for F.c, C.ICEMDV and C.priscuii; however, k = 4 returns 'NULL' for all GAMs 
# in s(CalcSal, k = 3) , k is good
# in ti(CalcSal, par_ue, k = 3) k is too low for C.priscuii; however, k = 4 returns 'NULL' for all GAMs 
```

```{r Compute flash spacing}

# Computing flash spacing averages for secondary axes of heatmaps for algae
spacing.algae <- sstate_damping %>%
  filter(Type == "Green Algae") %>%
  group_by(PulseSpace_s) %>%
  summarise(
    Light_Level.mean = mean(Light_Level, na.rm = TRUE),
    Light_Level.sd = sd(Light_Level, na.rm = TRUE),
    .groups = "drop"
  )

# Computing flash spacing averages for secondary axes of heatmaps for diatoms
spacing.diatom <- sstate_damping %>%
  filter(Type == "Diatom") %>%
  group_by(PulseSpace_s) %>%
  summarise(
    Light_Level.mean = mean(Light_Level, na.rm = TRUE),
    Light_Level.sd = sd(Light_Level, na.rm = TRUE),
    .groups = "drop"
  )


```



```{r Plot Algae DampingIndex GAM}

# Run this chunk each time for each strain to generate strain specific hypervolume 
# plotting all strains and adding a facet-wrap does not work as the contour lines are determined for each individual strain.
GamStrain = "Chlorella_vulgaris"
#Chlamydomonas_ICEMDV 
#Chlamydomonas_malina 
#Chlamydomonas_priscuii 
#Chlorella_vulgaris 
#Chlamydomonas_reinhardtii

ErrorTol = 170


DampingIndex_Hyper <- DampingIndex_Pred %>% 
  filter(Strain %in% GamStrain) %>%
  group_by(Strain, GrowthTemp_C) %>% 
  mutate(IsFitGood = case_when(se.fit/fit*100 < ErrorTol ~ (1),#Assigns '0' to faulty data where Standarderror larger +/- Errortolerance
                    TRUE ~ (0))) %>% 
  filter(IsFitGood == 1) %>%  # filters faulty data where Standarderror larger +/- Errortolerance
  ungroup()
 
       
maxDI <- DampingIndex_Hyper %>%
  filter(fit == max(fit, na.rm = TRUE))
maxDI <- maxDI$fit

minDI <- DampingIndex_Hyper %>%
  filter(fit == min(fit, na.rm = TRUE))
minDI <- minDI$fit


ContourBreaks <- DampingIndex_Hyper %>%
  group_by(Strain) %>% 
  mutate(upper10 = quantile(fit, 0.9, na.rm = TRUE),
         upper25 = quantile(fit, 0.75, na.rm = TRUE),
         median50 = quantile(fit, 0.5, na.rm = TRUE),
         lower25 = quantile(fit, 0.25, na.rm = TRUE),
         lower10 = quantile(fit, 0.1, na.rm = TRUE)) %>%
  ungroup() %>%
  dplyr::select(Strain, upper10, upper25, median50, lower25, lower10) %>%
  distinct() %>%
  pivot_longer(cols = c(upper10, upper25, median50, lower25, lower10), names_to = 'quantile', values_to = 'quantileValue') %>%
  mutate(quantileValue = round(quantileValue, digits = 1))


###plot
#https://stackoverflow.com/questions/28775036/ggplot-line-graph-with-na-values
#https://stackoverflow.com/questions/6906661/ggplot2-make-missing-value-in-geom-tile-not-blank
#https://stackoverflow.com/questions/24822621/multiple-legends-in-a-plot-with-geom-tile
#https://stackoverflow.com/questions/43395219/plotting-multiple-layers-with-geom-raster-or-geom-tile-or-geom-rect
#https://stackoverflow.com/questions/59949635/ggplot-align-grid-lines-with-end-of-geom-tile
#https://stackoverflow.com/questions/22945651/remove-space-between-plotted-data-and-the-axes
#https://stackoverflow.com/questions/5293715/how-to-use-greek-symbols-in-ggplot2


ContourLines <- ContourBreaks 
ContourLines <- ContourLines$quantileValue 

(ContourDI <- ggplot(na.omit(DampingIndex_Hyper,spacing.algae), aes(Temp_C, Light_Level, z = fit, fill = fit)) +
  geom_tile() +
  scale_fill_gradientn(name = bquote("Damping Index"), colors = pals::ocean.matter(14), limits  = range(0, 13), na.value = "pink") +
    metR::geom_contour2(aes(label = stat(level)),
                        breaks = ContourLines,
                        skip = 0,
                        color = "white",
                        linemitre = 10,
                        linewidth = 0.5,
                        stroke = 0.5,
                        size = 3.3,
                        label.placer = label_placer_fraction(frac = 0.4)) +
    # geom_text_contour(breaks = ContourLines, stroke = 0.1, size = 3.3,
    #               label.placer = label_placer_fraction(frac = 0.4)) +
    scale_x_continuous(limits = c(10, 26), breaks = seq(10, 26, by = 4), expand = c(0, 0)) + # For C.vulgaris & T.p
    # scale_x_continuous(limits = c(4, 12), breaks = seq(4, 12, by = 4), expand = c(0, 0)) + # For C.ICE, C.malina & C.priscuii
     # scale_x_continuous(limits = c(12, 24), breaks = seq(12, 24, by = 4), expand = c(0, 0)) + # For C.reinhardtii
    scale_y_continuous(limits = c(0, 0.7), breaks = seq(0.2, 0.6, by = 0.2), expand=c(0,0), 
                       sec.axis = dup_axis(breaks = c(spacing.algae$Light_Level.mean), 
                                           labels = c(spacing.algae$PulseSpace_s),
                                           name = sym("Equivalent Flash Spacing (s)"))) +
    coord_cartesian(expand = FALSE, clip = "on") +
    # labs(title = GamStrain) +
    xlab("Measurement Temperature (°C)") +
    ylab("Light Level (µmol photons"~ m^-2~ ~s^-1~ ")") +
    # ylab(NULL) +
  ggh4x::facet_nested_wrap(vars(l.strain), nrow=1, scales="free_x",
                             labeller = label_parsed) +
  geom_vline(aes(xintercept = GrowthTemp_C + 0.025),
             col= "black", linetype = 2, linewidth = 1.7) +
  # ggh4x::force_panelsizes(cols = c(10, 8, 8, 8, 18, 16, 12), respect = TRUE, total_width = unit(12, "cm")) +
  theme_thesis()+
    theme(legend.position = "right")  +
  theme(panel.spacing.x = unit(4, "mm")))
        
        # ,
        # legend.position = c(0.89, 0.07), legend.justification = c(1,0)))

ContourDI

# Dynamically name the plot with strain variable
assign(paste0("ti_plot_", GamStrain), ContourDI)

```

```{r Combine and save Algae Damping Index GAMs}

Filename <- "AlgaeGAM"

# Remove individual legends
p1 <- ti_plot_Chlamydomonas_ICEMDV + theme(legend.position = "none",
    axis.title.y.right = element_blank()) +
  labs(x = NULL, y = NULL)
p2 <- ti_plot_Chlamydomonas_malina + theme(legend.position = "none",
    axis.title.y.right = element_blank()) +
  labs(x = NULL, y = NULL)
p3 <- ti_plot_Chlamydomonas_priscuii+ theme(legend.position = "none",
    axis.title.y.right = element_blank()) +
  labs(x = NULL, y = NULL)
p4 <- ti_plot_Chlamydomonas_reinhardtii + theme(legend.position = "none",
    axis.title.y.right = element_blank()) +
  labs(x = NULL, y = NULL)
p5 <- ti_plot_Chlorella_vulgaris + theme(legend.position = "none",
    axis.title.y.right = element_blank()) +
  labs(x = NULL, y = NULL)

# Extract vertical legend from p1
legend_grob <- get_legend(
  p1 + theme(legend.position = "right", legend.direction = "vertical")
)

# Combine with manual layout
CombineAlgae <- patchwork::wrap_plots(
  p1, p2, p3,
 p4, p5, gridExtra::arrangeGrob(legend_grob),
  design = c(
    area(1, 1), area(1, 2), area(1, 3),  # row 1
    area(2, 1), area(2, 2), area(2, 3)   # row 2 (legend under p3)
  )
)

CombineAlgae

# full layout
# ggdraw uses the 'cowplot' package
AlgaeGAM <- ggdraw() +
  draw_plot(CombineAlgae, x = 0.1, y = 0.1, width = 0.85, height = 0.85) +
  draw_label("Measurement Temperature (°C)", x = 0.5, y = 0.04, size = 20) +
  draw_label("Light Level (µmol photons"~ m^-2~ ~s^-1~ ")", x = 0.05, y = 0.5, angle = 90, size = 20) +
  draw_label("Equivalent Flash Spacing (s)", x = 0.97, y = 0.5, angle = 270, size = 20)
    
AlgaeGAM

# must use cowplot to save as x and y axisa are lost using ggsave 
cowplot::save_plot(
  filename = file.path(PlotsOut, paste0(Filename, ".png")),
  plot = AlgaeGAM,
  base_height = 10,
  base_width = 15,
  dpi = 300,
  bg ="white" # need a white background for axis visibility
)
```


### Next Combine Diatoms

```{r Plot Diatoms DampingIndex GAM}

# Run this chunk each time for each strain to generate strain specific hypervolume 
# plotting all strains and adding a facet-wrap does not work as the contour lines are determined for each individual strain.

 GamStrain = "Fragilariopsis_cylindrus"
#Fragilariopsis_cylindrus
#Thalassiosira_pseudonana 

GamTemp = 6 #0 #6 #22 for T.p

ErrorTol = 310


DampingIndex_Hyper <- DampingIndex_Pred %>% 
  filter(Strain %in% GamStrain) %>%
  filter(GrowthTemp_C %in% GamTemp) %>% 
  group_by(Strain, GrowthTemp_C) %>% 
  mutate(IsFitGood = case_when(se.fit/fit*100 < ErrorTol ~ (1),#Assigns '0' to faulty data where Standarderror larger +/- Errortolerance
                    TRUE ~ (0))) %>% 
  filter(IsFitGood == 1) %>%  # filters faulty data where Standarderror larger +/- Errortolerance
  mutate(FacetLabel = paste(l.strain, FacetLabel_temp, sep = " ~ ")) %>% 
  ungroup()
 
       
maxDI <- DampingIndex_Hyper %>%
  filter(fit == max(fit, na.rm = TRUE))
maxDI <- maxDI$fit

minDI <- DampingIndex_Hyper %>%
  filter(fit == min(fit, na.rm = TRUE))
minDI <- minDI$fit


ContourBreaks <- DampingIndex_Hyper %>%
  group_by(Strain) %>% 
  mutate(upper10 = quantile(fit, 0.9, na.rm = TRUE),
         upper25 = quantile(fit, 0.75, na.rm = TRUE),
         median50 = quantile(fit, 0.5, na.rm = TRUE),
         lower25 = quantile(fit, 0.25, na.rm = TRUE),
         lower10 = quantile(fit, 0.1, na.rm = TRUE)) %>%
  ungroup() %>%
  dplyr::select(Strain, upper10, upper25, median50, lower25, lower10) %>%
  distinct() %>%
  pivot_longer(cols = c(upper10, upper25, median50, lower25, lower10), names_to = 'quantile', values_to = 'quantileValue') %>%
  mutate(quantileValue = round(quantileValue, digits = 1))


###plot
#https://stackoverflow.com/questions/28775036/ggplot-line-graph-with-na-values
#https://stackoverflow.com/questions/6906661/ggplot2-make-missing-value-in-geom-tile-not-blank
#https://stackoverflow.com/questions/24822621/multiple-legends-in-a-plot-with-geom-tile
#https://stackoverflow.com/questions/43395219/plotting-multiple-layers-with-geom-raster-or-geom-tile-or-geom-rect
#https://stackoverflow.com/questions/59949635/ggplot-align-grid-lines-with-end-of-geom-tile
#https://stackoverflow.com/questions/22945651/remove-space-between-plotted-data-and-the-axes
#https://stackoverflow.com/questions/5293715/how-to-use-greek-symbols-in-ggplot2


ContourLines <- ContourBreaks 
ContourLines <- ContourLines$quantileValue 

(ContourDI <- ggplot(na.omit(DampingIndex_Hyper, spacing.diatom), aes(Temp_C, Light_Level, z = fit, fill = fit)) +
  geom_tile() +
  scale_fill_gradientn(name = bquote("Damping Index"), colors = pals::ocean.matter(14), limits  = range(0, 16.5), na.value = "pink") +
    metR::geom_contour2(aes(label = stat(level)),
                        breaks = ContourLines,
                        skip = 0,
                        color = "white",
                        linemitre = 10,
                        linewidth = 0.5,
                        stroke = 0.5,
                        size = 3.3,
                        label.placer = label_placer_fraction(frac = 0.4)) +
    # geom_text_contour(breaks = ContourLines, stroke = 0.1, size = 3.3,
    #               label.placer = label_placer_fraction(frac = 0.4)) +
    # scale_x_continuous(limits = c(10, 28), breaks = seq(10, 28, by = 4), expand = c(0, 0)) + # For T.pseudonana
    scale_x_continuous(limits = c(0, 10), breaks = seq(2, 10, by = 4), expand = c(0, 0)) + # For F.c
    scale_y_continuous(limits = c(0, 0.7), breaks = seq(0.2, 0.6, by = 0.2), expand=c(0,0), 
                       sec.axis = dup_axis(breaks = c(spacing.diatom$Light_Level.mean), 
                                           labels = c(spacing.diatom$PulseSpace_s),
                                           name = sym("Equivalent Flash Spacing (s)"))) +
 coord_cartesian(expand = FALSE, clip = "on") +
  #   geom_text(aes(x = 0.1, y = 0.60, label = paste("Grown at", GrowthTemp_C, "°C", sep = " "), 
  #                 show.legend = FALSE),
  #   color = "white", 
  #   hjust = 0,
  #   vjust = 1,
  #   size = 5.0
  # )  +
    xlab("Measurement Temperature (°C)") +
    ylab("Light Level (µmol photons"~ m^-2~ ~s^-1~ ")") +
    # ylab(NULL) +
  ggh4x::facet_nested_wrap(vars(FacetLabel), nrow=1, scales="free_x",
                             labeller = label_parsed) +
  geom_vline(aes(xintercept = GrowthTemp_C + 0.025),
             col= "black", linetype = 2, linewidth = 1.7) +
  # ggh4x::force_panelsizes(cols = c(10, 8, 8, 8, 18, 16, 12), respect = TRUE, total_width = unit(12, "cm")) +
  theme_thesis()+
    theme(legend.position = "right")  +
  theme(panel.spacing.x = unit(4, "mm")))

ContourDI

# Dynamically name the plot with strain variable
assign(paste0("ti_plot_", GamStrain, GamTemp), ContourDI)

```


```{r Combine and save Algae Damping Index GAMs}

Filename <- "DiatomsGAM"

# Remove individual legends
p1 <- ti_plot_Fragilariopsis_cylindrus0 + theme(legend.position = "none",
    axis.title.y.right = element_blank()) +
  labs(x = NULL, y = NULL)
p2 <- ti_plot_Fragilariopsis_cylindrus6 + theme(legend.position = "none",
    axis.title.y.right = element_blank()) +
  labs(x = NULL, y = NULL)
p3 <- ti_plot_Thalassiosira_pseudonana22 + theme(legend.position = "none",
    axis.title.y.right = element_blank()) +
  labs(x = NULL, y = NULL)


# Extract vertical legend from p1
legend_grob <- get_legend(
  p1 + theme(legend.position = "right", legend.direction = "vertical")
)

# Combine with manual layout
CombineDiatoms <- patchwork::wrap_plots(
  p1, p2,
 p3, gridExtra::arrangeGrob(legend_grob),
  design = c(
    area(1, 1), area(1, 2),  # row 1
    area(2, 1), area(2, 2)   # row 2 (legend under p2)
  )
)

CombineDiatoms

# full layout
# ggdraw uses the 'cowplot' package
DiatomGAM <- ggdraw() +
  draw_plot(CombineDiatoms, x = 0.1, y = 0.1, width = 0.85, height = 0.85) +
  draw_label("Measurement Temperature (°C)", x = 0.5, y = 0.04, size = 20) +
  draw_label("Light Level (µmol photons"~ m^-2~ ~s^-1~ ")", x = 0.05, y = 0.5, angle = 90, size = 20) +
  draw_label("Equivalent Flash Spacing (s)", x = 0.97, y = 0.5, angle = 270, size = 20)
    
DiatomGAM

# must use cowplot to save as x and y axisa are lost using ggsave 
cowplot::save_plot(
  filename = file.path(PlotsOut, paste0(Filename, ".png")),
  plot = DiatomGAM,
  base_height = 10,
  base_width = 15,
  dpi = 300,
  bg ="white" # need a white background for axis visibility
)

```






```{r}
predict.gam.t <- lapply(temp.sstate, predict.temp.gam)


predict.gam.t <- bind_rows(predict.gam.t) %>%
  mutate(l.temp = paste0("Grown~at~", GrowthTemp_C,"*degree*C"), 
         l.temp )
```

```{r}
summary.gam.t <- lapply(temp.sstate, summary.temp.gam)
```


##### Diatoms

```{r}
predict.gam.t %>%
  filter(Strain %in% c("Thalassiosira_pseudonana", "Fragilariopsis_cylindrus")) %>%
  ggplot(aes(x=Temp_C, y=Light_Level, z= DampingIndex)) +
  geom_contour_fill() + 
  geom_contour2(color = "white", linemitre = 10, linewidth=0.25) +
  geom_text_contour(stroke = 0.1, family = "Times New Roman", size = 3.3, 
                    label.placer = label_placer_fraction(frac=0.4)) +
  labs(x= "Measurement Temperature (°C)",
       y= expression("Light Level ("*mu*"mol photons m"^"-2"~s^-1*")"),
       fill= "Damping Index") +
  scale_fill_gradientn(colors = pals::ocean.matter(14), breaks = seq(0, 13, by = 3)) +
  coord_cartesian(ylim = c(0.03, 0.53)) +
  scale_x_continuous(breaks = c(2, 6, 10, 14, 18, 22, 26), expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0), sec.axis = dup_axis(
    breaks = c(spacing.diatom$Light_Level.mean), labels = c(spacing.diatom$PulseSpace_s),
    name = sym("Equivalent Flash Spacing (s)"))) +
  ggh4x::facet_nested_wrap(vars(l.strain, l.temp), ncol = 2, 
                           strip.position = "top", scales="free_x", 
                           labeller = label_parsed) +
  geom_vline(data=diatom.vlines, aes(xintercept=GrowthTemp_C), 
             col= "white", linetype = 2, linewidth = 1) +
  theme_thesis() +
  theme(panel.spacing.x = unit(4, "mm"),
        legend.position = c(0.89, 0.07), legend.justification = c(1,0))


# ggsave("diatoms_temp.png", width = 7.5 , height = 5, dpi = 900,
#        plot=last_plot(),
#        path="../Output")
```

```{r}
comp.labs <- str_wrap(c("Thalassiosira_pseudonana_22" = "Thalassiosira pseudonana",
               "Fragilariopsis_cylindrus_6" = "Fragilariopsis cylindrus grown at 6°C",
               "Fragilariopsis_cylindrus_0" = "Fragilariopsis cylindrus grown at 0°C"), width = 25)

predict.gam.t %>%
  filter(Strain %in% c("Thalassiosira_pseudonana", "Fragilariopsis_cylindrus")) %>%
  filter(Temp_C == 10) %>%
  mutate(Culture = paste(Strain, GrowthTemp_C, sep = "_")) %>%
  ggplot() +
  geom_point(aes(x= Light_Level, y=DampingIndex, col= Culture, shape= Culture )) +
  geom_path(aes(x= Light_Level, y=DampingIndex, col= Culture)) +
  labs(x = expression("Light Level ("*mu*"mol photons m"^"-2"~s^-1*")"), y = "Damping Index") +
  scale_color_manual(values=c("#009392","#91B8AB", "#DFA0A0"), labels=c(comp.labs)) +
  scale_shape_discrete(labels=c(comp.labs)) + 
  scale_x_continuous(limits = c(0, 0.66), sec.axis = dup_axis(breaks = c(spacing.diatom$Light_Level.mean), 
                labels = c(spacing.diatom$PulseSpace_s), name = sym("Equivalent Flash Spacing (s)")), 
                expand= c(0.01,0.01)) +
  scale_y_continuous(limits = c(0,12), breaks = c(2,4,6,8,10,12)) +
  theme_thesis() +
  theme(legend.key.size = unit(1, "cm"), legend.text = element_text(face = "italic"))

# ggsave("diatom_comp.png", width = 8 , height = 4, dpi = 900,
#        plot=last_plot(),
#        path="../Output")
```

##### Algae
```{r}
predict.gam.t %>%
  filter(Strain %in% c("Chlamydomonas_ICEMDV", "Chlamydomonas_malina", "Chlamydomonas_priscuii", 
             "Chlorella_vulgaris", "Chlamydomonas_reinhardtii")) %>%
  ggplot(aes(x=Temp_C, y=Light_Level, z= DampingIndex)) +
  geom_contour_fill() + 
  geom_contour2(color = "white", linemitre = 10, linewidth=0.25) +
  geom_text_contour(stroke = 0.1, family = "Times New Roman", size = 3.5, 
                    label.placer = label_placer_random()) +
  labs(x= "Measurement Temperature (°C)",
       y= expression("Light Level ("*mu*"mol photons m"^"-2"~s^-1*")"),
       fill= "Damping Index") +
  scale_fill_gradientn(colors = pals::ocean.matter(14), breaks = seq(0, 13, by = 3)) +
  coord_cartesian(ylim = c(0.03, 0.66)) +
  scale_x_continuous(breaks= c(0,4,8,12,16,20,24,28), expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0),sec.axis = dup_axis(
    breaks = c(spacing.algae$Light_Level.mean), labels = c(spacing.algae$PulseSpace_s),
    name = sym("Equivalent Flash Spacing (s)"))) +
  ggh4x::facet_nested_wrap(vars(ThermalTolerance, l.strain), ncol = 3, strip.position = "top",
                       scales = "free_x", 
                       labeller = label_parsed) +
  geom_vline(data=algae.vlines, aes(xintercept=GrowthTemp_C), 
             col= "white", linetype = 2, linewidth = 1) +
  theme_thesis() +
  theme(strip.placement = "outside",
        panel.spacing.x = unit(4, "mm"),
        legend.position = c(0.93, 0.07), legend.justification = c(1,0))

# ggsave("algae_temp.png", width = 9.3 , height = 6, dpi = 900,
#        plot=last_plot(),
#        path="../Output")
```

```{r}
algae.labs <- c("Chlamydomonas_ICEMDV" = "Chlamydomonas ICEMDV", 
                 "Chlamydomonas_malina" = "Chlamydomonas malina",
                 "Chlamydomonas_priscuii" = "Chlamydomonas priscuii", 
                 "Chlorella_vulgaris" = "Chlorella vulgaris", 
                 "Chlamydomonas_reinhardtii" = "Chlamydomonas reinhardtii")
               

predict.gam.t %>%
  filter(Strain %in% c("Chlamydomonas_ICEMDV", "Chlamydomonas_malina", "Chlamydomonas_priscuii", 
             "Chlorella_vulgaris", "Chlamydomonas_reinhardtii")) %>%
  filter(Temp_C == 12) %>%
  ggplot() +
  geom_point(aes(x= Light_Level, y=DampingIndex, col= Strain, shape= Strain)) +
  geom_path(aes(x= Light_Level, y=DampingIndex, col= Strain)) +
  labs(x = expression("Light Level ("*mu*"mol photons m"^"-2"~s^-1*")"),
       y = "Damping Index", col = "Strain", shape= "Strain") +
  scale_color_manual(values=c(rcartocolor::carto_pal(n=5,"Temps")), labels=c(algae.labs)) +
  scale_shape_manual(values=c(16,3,17,18,8), labels=c(algae.labs)) +
  scale_x_continuous(sec.axis = dup_axis(breaks = c(spacing.algae$Light_Level.mean), 
                labels = c(spacing.algae$PulseSpace_s), name = sym("Equivalent Flash Spacing (s)")), 
                expand= c(0.01,0.01)) +
  scale_y_continuous(limits = c(0,12), breaks = c(2,4,6,8,10,12)) +
  theme_thesis() +
  theme(legend.text = element_text(face="italic"))

# ggsave("algae_comp.png", width = 8 , height = 4, dpi = 900,
#        plot=last_plot(),
#        path="../Output")
```

# Evaluating Outlier in *C. malina*
```{r}
Cm_damping <- subset(sstate_damping, Strain == "Chlamydomonas_malina")

# Original model 
Cm_gam <- gam(DampingIndex ~ te(Temp_C, Light_Level, k=3), data = Cm_damping, method = "REML")
cooksd <- cooks.distance(Cm_gam)
plot(1:length(cooksd), cooksd, type = "h")

# Remove outlier point
Cm_rmoutlier <- Cm_damping %>% filter(!Light_Level == 0.07333)

# Re-run model 
Cm_rm.out <- gam(DampingIndex ~ te(Temp_C, Light_Level, k=3), data = Cm_damping, method = "REML")
summary(Cm_rm.out)

```



# For Presentation 
```{r}
# theme_pres <- function() {
#   theme_bw() +
#   theme(text = element_text(family = "Georgia"), 
#         axis.text=element_text(size=11), axis.title=element_text(size=13),
#         legend.text=element_text(size=11), legend.title=element_text(size=11),
#         strip.background = element_rect(colour="black", fill=NA),
#         strip.text = element_text(size=11), 
#         panel.grid.major = element_blank(), 
#         panel.grid.minor = element_blank()
#   )
# }
```

```{r warning=FALSE}
power_list <- lapply(names(Cp_wavelets), function(condition_name) {
    condition <- Cp_wavelets[[condition_name]]
    Power.avg <- condition$Power.avg
    Power.avg.pval <- condition$Power.avg.pval
    Period <- condition$Period
    Condition <- rep(condition_name, length(Power.avg))
    df <- data.frame(Power.avg, Power.avg.pval, Period, Condition) %>%
      separate(Condition, into = c("Strain", "PulseSpace_s", "Temp_C", "LightLevel"), sep = "_") %>%
      mutate(Power.avg.pval = ifelse(Power.avg.pval <= 0.05, 0, 1)) 
     return(df)
  })
power.data <- bind_rows(power_list)
  
power.data %>%
    filter(Temp_C == "8") %>% 
    group_by(PulseSpace_s) %>%
    mutate(Light_Level = mean(as.numeric(LightLevel))) %>%
    mutate(Light_Level = round(Light_Level, 2)) %>%
    ungroup() %>%
    ggplot()+ 
    geom_path(aes(x= Power.avg, y= Period, col=Power.avg.pval), linewidth=0.8) +
    scale_color_gradient(low="skyblue", high="black", breaks= c(0.05, 1), 
                         labels= c("p ≤ 0.05", "p > 0.05"), guide="legend") +
    labs(x= "Average Wavelet Power", colour= bquote(atop("Significance of", "ChlF Oscillations"))) +
    ggh4x::facet_nested(cols=vars(Light_Level)) +
    scale_y_continuous(limits= c(0,10), breaks= c(0,4,8), expand=c(0,0)) +
    scale_x_continuous(limits= c(0,1), breaks=c(0, 0.4, 0.8),expand=c(0,0), 
          sec.axis=sec_axis(~ . , name = bquote(Light ~ Level ~ (µmol ~ photons ~ m^-2 ~ s^-1)), 
                            breaks = NULL, labels = NULL))
#+
 #   theme_pres() 

# ggsave("wtpower_pres.png", width = 9.5 , height = 4.5, dpi = 900,
#        plot=last_plot(),
#        path="../Output/HonoursDay")
```

```{r warning = false}
pres.reconstruction <- function(reconstructed) {
  result_list <- Map(function(condition_name, condition_data) {
    series <- as.data.frame(condition_data$series)
    series <- mutate(series, Condition = condition_name) %>%
      mutate(Flashnumber = row_number())  %>%
      separate(Condition, into = c("Strain", "PulseSpace_s", "Temp_C", "LightLevel"),
                sep = "([\\/\\_])") %>%
       mutate(PulseSpace_s = factor(PulseSpace_s))
    return(series)
  }, names(reconstructed), reconstructed) }

Cp.recon <- bind_rows(pres.reconstruction(Cp_reconstruction)) %>%
  group_by(PulseSpace_s) %>%
  mutate(LightLevel = round(mean(as.numeric(LightLevel)), 2)) %>%
  ungroup() %>%
  select(-c(FvFmrunnorm.trend)) %>%
  filter(Temp_C == "8", LightLevel == 0.71)

Cp.long <- Cp.recon %>%
  pivot_longer(cols = starts_with("FvFmrunnorm"), names_to = "Series", values_to = "FvFmrunnorm_value") %>%
  mutate(Series = ifelse(Series == "FvFmrunnorm", "Original~Wavelet",
                paste(expression(paste("Reconstruction at ", alpha, " = 0.05")))))
                  
 Cp.long %>%
   ggplot(aes(x=Flashnumber, y=FvFmrunnorm_value, col = Series)) +
   geom_line() + 
   facet_wrap(vars(Series), labeller=label_parsed) + 
   scale_colour_manual(values=c("black", "deepskyblue3")) +
   labs(x= "Flash Number",  y="Transformed ChlF") +
   scale_x_continuous(breaks = seq(0,32,4), expand = c(0,0)) +
   #theme_pres() +
   theme(legend.position = "none", panel.spacing.x = unit(2, "cm"), 
         strip.background = element_rect(colour= NA, fill=NA),
         strip.text = element_text(size=17, margin=margin(b=15)),
         panel.border = element_rect(color = "black", size = 1))
 
# ggsave("reconstruct_pres.png", width = 11 , height = 4.5, dpi = 900,
#        plot=last_plot(),
#        path="../Output/HonoursDay")
```


```{r}
spacing.average <- sstate_damping %>%
  mutate(DampingIndex = ifelse(DampingIndex > 0, 1, 0)) %>%
  group_by(PulseSpace_s) %>%
  mutate(mean.light = mean(Light_Level),
         mean.light = round(mean.light,2)) %>%
  ungroup()

spacing.average  %>%
  ggplot() +
  geom_tile(aes(x= as.factor(Temp_C), y=as.factor(mean.light), fill=as.factor(DampingIndex))) +
  scale_fill_manual(values= c("black","skyblue"), 
                      name= bquote(atop("Significance of", "ChlF Oscillations")), 
                      labels =c("p > 0.05", "p ≤ 0.05")) +
  labs(x = "Measurement Temperature (°C)", y= bquote(Light ~ Level ~ (µmol ~ photons ~ m^-2 ~ s^-1))) +
  facet_grid(cols=vars(ThermalTolerance), scales = "free_x", space = "free_x") +
  scale_x_discrete(expand=c(0,0)) + scale_y_discrete(expand=c(0,0))
#+
 # theme_pres() 

# ggsave("mtrx_pres.png", width = 10 , height = 4, dpi = 900,
#        plot=last_plot(),
#        path="../Output/HonoursDay")
```

```{r}
damping.group <- sstate_damping %>% 
  doBy::splitBy(formula = ~ ThermalTolerance) 
```

```{r}
predict.diff.group <- function(group.list) {
  combineddata <- lapply(group.list, function(damping){
  gam <- mgcv::gam(DampingIndex ~ te(TempDiff_C, Light_Level, k= n_distinct(damping$TempDiff_C)), 
                   data = damping, method = "REML")

  data <- with(damping, expand.grid( Light_Level = seq(0.025, 0.65, 0.01),
    TempDiff_C = seq(min(TempDiff_C), max(TempDiff_C), 0.1)))
  
  data <- data %>%
    mutate(DampingIndex = predict(gam, newdata = data), 
           DampingIndex = ifelse(DampingIndex < 0, 0, DampingIndex),
           ThermalTolerance = factor(unique(damping$ThermalTolerance))) 
      return(data) })
  combineddata <- do.call(bind_rows, combineddata)
  return(combineddata)
}
```

```{r}
predict.group <- predict.diff.group(damping.group)
legend.lab <- str_wrap("Duration of Significant ChlF Oscillations", width = 20)

predict.group %>%
  ggplot(aes(x=TempDiff_C, y=Light_Level, z= DampingIndex)) +
  geom_contour_fill() + 
  geom_contour2(color = "white", linemitre = 10, linewidth=0.25) +
  geom_text_contour(stroke = 0.1, family = "Times New Roman", size = 3.5) +
  labs(x= "Deviation from Growth Temperature During Measurement (°C)",
       y= expression("Light Level ("*mu*"mol photons m"^"-2"~s^-1*")"),
       fill = legend.lab) +
  scale_fill_gradientn(colors = pals::ocean.deep(13), limits=c(0,12), breaks = c(0,3,6,9,12)) +
  scale_x_continuous(expand=c(0,0), limits= c(-6,6)) + scale_y_continuous(expand=c(0,0)) +
  facet_wrap(vars(ThermalTolerance), ncol = 2) +
  #theme_pres() +
  theme(panel.spacing.x = unit(6, "mm"))

# ggsave("between_pres.png", width = 9 , height = 4, dpi = 900,
#        plot=last_plot(),
#        path="../Output/HonoursDay")
```

```{r}
gam.within <- mgcv::gam(DampingIndex ~ te(Temp_C, Light_Level), 
                   data = sstate_damping, method = "REML")

data.within <- with(sstate_damping, expand.grid( Light_Level = seq(0.025, 0.65, 0.01),
    Temp_C = seq(min(Temp_C), max(Temp_C), 0.1)))
  
data.within <- data.within %>%
    mutate(DampingIndex = predict(gam.within, newdata = data.within), 
           DampingIndex = ifelse(DampingIndex < 0, 0, DampingIndex))

data.within %>%
  ggplot(aes(x=Temp_C, y=Light_Level, fill= DampingIndex)) +
  geom_tile() + 
  labs(x= "Measurement Temperature (°C)", y= expression("Light Level ("*mu*"mol photons m"^"-2"~s^-1*")"),
       fill= legend.lab) +
  scale_fill_gradientn(colors = pals::ocean.matter(16272),limits=c(0,12), breaks = c(0,3,6,9,12)) +
  scale_x_continuous(expand=c(0,0), breaks=c(0,4,8,12,16,20,24,28)) + scale_y_continuous(expand=c(0,0))

#+
 # theme_pres() 

# ggsave("within_pres.png", width = 7.5 , height = 4.2, dpi = 900,
#        plot=last_plot(),
#        path="../Output/HonoursDay")
```

