---
title: "GAM and Figures"
author: "Natasha Ryan"
output: 
  html_document: default
editor_options: 
  chunk_output_type: inline
---

This RMD imports processed generated in Wavelet_LowLight.Rmd for generalized additive modelling and generating figures for Natasha Ryan's undergraduate honours thesis at Mount Allison University. 

# Setup
```{r load libraries}
library(tidyverse)
library(metR)
library(mgcv)
library(ggpubr)
```

```{r create ggplot theme}
theme_thesis <- function() {
  ggplot2::theme_bw() +
  ggplot2::theme(text = element_text(family = "Times New Roman"), 
        axis.text=element_text(size=11), axis.title=element_text(size=13),
        legend.text=element_text(size=11), legend.title=element_text(size=13),
        strip.background = element_rect(colour="black", fill=NA),
        strip.text = element_text(size=11), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.title.y.right = element_text(size = 12), 
        axis.title.x.top = element_text(size = 12)
  )}
```


# Representative Fluorescence Oscillations

```{r warning=FALSE}
Cp_data <- readRDS("../Data/CleanData/LOW_Chlamydomonas_priscuii_SolFitsTrim.Rds") %>%
  mutate(SState = ifelse(SState == 4, 0, SState), 
         SState = factor(SState, levels = c(1, 2, 3, 0))) %>%
  ungroup()

Cp_data <- Cp_data %>%
    group_by(PulseSpace_s) %>%
    mutate(Approx_Light_Level = mean(as.numeric(Light_Level))) %>%
    mutate(Approx_Light_Level = round(Approx_Light_Level, 3)) %>%
    ungroup()
```


```{r warning=FALSE}
Cp_data %>%
  filter(Filename %in% c("20230918_MaPo1001_12C_1s", "20230918_MaPo1001_12C_4s", "20230919_MaPo1004_12C_16s",
                      "20230918_MaPo1001_4C_1s",  "20230918_MaPo1001_4C_4s", "20230918_MaPo1001_4C_16s")) %>%
  ggplot() +
  geom_point(aes(x=Flashnumber, y=FvFmrunnorm, col=as.factor(SState))) +
  scale_color_manual(values = c(rcartocolor::carto_pal(n=4,"Temps"))) +
  labs(x= "Flash Number", 
       y=expression("Normalized Chlorophyll Fluorescence (F"[v]*" / F"[m]*")"),
       col= "Majority S-State") +
  ggh4x::facet_nested(rows=vars(Approx_Light_Level, PulseSpace_s), cols=vars(Temp_C), 
                      as.table = FALSE) +
  scale_y_continuous(sec.axis=sec_axis(~ . , 
                     name = bquote(Effective ~ Light ~ Level ~ (µmol ~ photons ~ m^-2 ~ s^-1)), 
                     breaks = NULL,labels = NULL)) +
  scale_x_continuous(breaks=c(4,8,12,16,20,24,28,32),
                     sec.axis=sec_axis(~ . , 
                                       name = sym("Measurement Temperature (°C)"), 
                                       breaks = NULL, labels = NULL)) +
  theme_thesis() 

# ggsave("representativeoscillations.png", width = 9, height = 5, dpi = 900,
#       plot=last_plot(),
#       path="../Output")
```



# FFT
```{r}
cols.2 <- c("#3F3C70","#40538C", "#4C99A0", "#96D2A6")

Cp_data %>%
  filter(CultureID == "MaPo1001", 
         ObsDate == "2023-09-18", 
         Temp_C == 4, 
         PulseSpace_s == 1) %>%
  ggplot() +
  geom_point(aes(x=Flashnumber, y=FvFmrunnorm, col=as.factor(SState)), size=1.5) +
  scale_color_manual(values = c(cols.2)) +
  labs(x= "Flash Number", 
       y=expression("Normalized Chlorophyll Fluorescence (F"[v]*" / F"[m]*")"),
       col= "Majority S-State") +
  scale_x_continuous(breaks=c(4,8,12,16,20,24,28,32)) +
  theme_pres() + 
  theme(legend.position = "none")

# ggsave("wavelet_before.png", width = 4, height = 4, dpi = 900,
#        plot=last_plot(),
#        path="../Output")
```

```{r}
wave_data <- Cp_wavelets[["Cp_1_4_0.77084"]]


WaveletComp::reconstruct(wave_data, only.sig=FALSE,
            plot.waves = TRUE, plot.rec = FALSE,
            spec.time.axis = list(at = c(0,4,8,12,16,20,24,28,32)),
            timelab = "Flash Number",
            verbose = FALSE)
```


# Wavelet Periodicity 

```{r}
wavelets <- readRDS("../Data/ProcessedData/LOW_Wavelets.Rds") 
specieslists <- split(wavelets, sub("_.*", "_wavelets", names(wavelets)))
list2env(specieslists, envir = .GlobalEnv)

FcG0_wavelets <- Fc_wavelets[grepl(paste("_0$", collapse = "|"), names(Fc_wavelets))]
FcG6_wavelets <- Fc_wavelets[grepl(paste("_6$", collapse = "|"), names(Fc_wavelets))]
```


```{r periodicity function}
plot.wt.grid <- function(wavelets) {
  result_list <- lapply(names(wavelets), function(condition_name) {
    condition <- wavelets[[condition_name]]
    Power.avg <- condition$Power.avg
    Power.avg.pval <- condition$Power.avg.pval
    Period <- condition$Period
    Condition <- rep(condition_name, length(Power.avg))  # Repeat condition name for each row
    df <- data.frame(Power.avg, Power.avg.pval, Period, Condition) %>%
      separate(Condition, into = c("Strain", "PulseSpace_s", "Temp_C", "LightLevel"), sep = "_") %>%
      mutate(PulseSpace_s = factor(PulseSpace_s, levels = c("1", "2","4","8","16")),
             Temp_C = factor(Temp_C, levels = 
                           c("0","2","4","6","8","10","12","14","16","18","20","22","24","26","28"))) %>%
      mutate(Power.avg.pval = ifelse(Power.avg.pval <= 0.05, 0, 1))  %>%
    
     return(df)
  })

  result_df <- bind_rows(result_list) %>%
    group_by(PulseSpace_s) %>%
    mutate(Light_Level = mean(as.numeric(LightLevel))) %>%
    mutate(Light_Level = round(Light_Level, 3)) %>%
    ungroup()
  
  plot <- result_df %>%
    ggplot()+ 
    geom_path(aes(x= Power.avg, y= Period, col=Power.avg.pval), linewidth=0.8) +
    scale_color_gradient(low="#4AC0AD", high="black", breaks= c(0.05, 1), 
                         labels= c("≤ 0.05", "> 0.05"), guide="legend") +
    labs(x= "Average Wavelet Power", colour= "P-Value") +
    ggh4x::facet_nested(rows=vars(Light_Level, PulseSpace_s), cols = vars(Temp_C),
                        as.table = FALSE) +
    scale_y_continuous(limits= c(0,10),
                       breaks= c(0,4,8), 
                 sec.axis=sec_axis(~ . , 
                             name = bquote(Effective ~ Light ~ Level ~ (µmol ~ photons ~ m^-2 ~ s^-1)),
                                   breaks = NULL,labels = NULL)) +
    scale_x_continuous(limits= c(0,1), 
                       breaks=c(0, 0.4, 0.8),
    sec.axis=sec_axis(~ . , name = sym("Measurement Temperature (°C)"), breaks = NULL, labels = NULL)) +
    theme_thesis() 
  return(plot)
}
```


```{r Periodicity Test, warning=FALSE}
plot.wt.grid(Ci_wavelets)
plot.wt.grid(Cm_wavelets)
plot.wt.grid(Cr_wavelets)
plot.wt.grid(Cv_wavelets)
plot.wt.grid(Tp_wavelets)
plot.wt.grid(FcG6_wavelets)
plot.wt.grid(FcG0_wavelets)

plot.wt.grid(Cr_wavelets)
 # ggsave("temperate_wtavg.png", width = 9.5, height = 5, dpi = 900,
 #      plot=last_plot(),
 #      path="../Output")

plot.wt.grid(Cp_wavelets)
 # ggsave("polar_wtavg.png", width = 9.5, height = 5, dpi = 900,
 #      plot=last_plot(),
 #      path="../Output")
```

# Reconstruction 

```{r}
reconstructions <- readRDS("~/LowLightPhotosynth/Data/ProcessedData/LOW_Reconstructions.Rds") 
reconstruction.list <- split(reconstructions, sub("_.*", "_reconstruction", names(reconstructions)))
list2env(reconstruction.list, envir = .GlobalEnv)

FcG0_reconstruction <- Fc_reconstruction[grepl(paste("_0$", collapse = "|"), names(Fc_reconstruction))]
FcG6_reconstruction <- Fc_reconstruction[grepl(paste("_6$", collapse = "|"), names(Fc_reconstruction))]
```


```{r}
plot.reconstruction <- function(reconstructed) {
  result_list <- Map(function(condition_name, condition_data) {
    series <- as.data.frame(condition_data$series)
    series <- mutate(series, Condition = condition_name) %>%
      mutate(Flashnumber = row_number())  %>%
      separate(Condition, into = c("Strain", "PulseSpace_s", "Temp_C", "LightLevel"),
                sep = "([\\/\\_])") %>%
       mutate(PulseSpace_s = factor(PulseSpace_s, levels = c("1","2","4","8","16")),
             Temp_C = factor(Temp_C, levels = 
                             c("0","2","4","6","8","10","12","14","16","18","20","22","24","26","28")))
    return(series)
  }, names(reconstructed), reconstructed)
  
  result_df <- bind_rows(result_list) %>%
    group_by(PulseSpace_s) %>%
    mutate(Light_Level = mean(as.numeric(LightLevel))) %>%
    mutate(Light_Level = round(Light_Level, 3)) %>%
    ungroup()
  
  plot <- result_df %>%
    ggplot()+
    geom_line(aes(x=Flashnumber, y=FvFmrunnorm, col = "Original")) +
    geom_line(aes(x=Flashnumber, y=FvFmrunnorm.r, col= "Reconstructed"))  +
    scale_colour_manual(values=c("black", "deepskyblue3"), 
                        labels = c("Original", expression(paste("Reconstucted (", alpha, " = 0.05)")))) +
    labs(x= "Flash Number", 
         y=expression("Normalized Chlorophyll Fluorescence (F"[v]*" / F"[m]*")"),
         col= "Wave Function") +
    ggh4x::facet_nested(rows=vars(Light_Level, PulseSpace_s), cols=vars(Temp_C), as.table = FALSE) +
    scale_y_continuous(sec.axis=sec_axis(~ . , 
                           name = bquote(Effective ~ Light ~ Level ~ (µmol ~ photons ~ m^-2 ~ s^-1)), 
                                         breaks = NULL,labels = NULL)) +
    scale_x_continuous(sec.axis=sec_axis(~ . , name = sym("Measurement Temperature (°C)"), 
                                         breaks = NULL, labels = NULL)) +
    theme_thesis()
  
  return(plot)
}
```


```{r warning=FALSE}
plot.reconstruction(Ci_reconstruction)
plot.reconstruction(Cm_reconstruction)
plot.reconstruction(Cr_reconstruction)
plot.reconstruction(Cv_reconstruction)
plot.reconstruction(Tp_reconstruction)
plot.reconstruction(FcG0_reconstruction)
plot.reconstruction(FcG6_reconstruction)


plot.reconstruction(Cp_reconstruction)
 # ggsave("representative_reconstruction.png", width = 10, height = 6, dpi = 900,
 #      plot=last_plot(),
 #      path="../Output")
```


# Damping Across Taxa

```{r import & wrangle data}
sstate_damping <- readRDS("~/LowLightPhotosynth/Data/ProcessedData/LOW_SState_Damping.Rds") %>%
  mutate(Strain = factor(Strain), 
  ThermalTolerance = case_when(
     Strain %in% c("Fragilariopsis_cylindrus", "Chlamydomonas_ICEMDV","Chlamydomonas_priscuii",
                   "Chlamydomonas_malina") ~ "Polar",
     Strain %in% c("Chlamydomonas_reinhardtii", "Chlorella_vulgaris",
                   "Thalassiosira_pseudonana") ~ "Temperate"),  
  Type = case_when(
     Strain %in% c("Fragilariopsis_cylindrus","Thalassiosira_pseudonana") ~ "Diatom", 
     Strain %in% c("Chlamydomonas_ICEMDV","Chlamydomonas_priscuii", "Chlamydomonas_malina",
                   "Chlamydomonas_reinhardtii", "Chlorella_vulgaris") ~ "Green Algae"),
  l.strain = case_when(
    Strain == "Chlamydomonas_ICEMDV" ~ "~italic('C. ICEMDV')", 
    Strain == "Chlamydomonas_malina" ~"~italic('C. malina')",
    Strain == "Chlamydomonas_priscuii" ~ "~italic('C. priscuii')",
    Strain == "Chlamydomonas_reinhardtii" ~ "~italic('C. reinhardtii')",
    Strain == "Chlorella_vulgaris" ~ "~italic('C. vulgaris')",
    Strain == "Fragilariopsis_cylindrus" ~ "~italic('F. cylindrus')", 
    Strain == "Thalassiosira_pseudonana" ~ "~italic('T. pseudonana')"),
  l.temp = paste0("Grown~at~", GrowthTemp_C,"*degree*C"))

diff.list <- doBy::splitBy(sstate_damping, formula = "Strain")
temp.list <- doBy::splitBy(sstate_damping, formula = c("Strain", "GrowthTemp_C"))
```


```{r Plot Setup}
# Vertical Lines for Growth Temperatures 
diatom.vlines <- data.frame(
  l.strain = c("~italic('F. cylindrus')", "~italic('F. cylindrus')", "~italic('T. pseudonana')"),
  l.temp = c("Grown~at~0*degree*C", "Grown~at~6*degree*C","Grown~at~22*degree*C"), 
  GrowthTemp_C = c(0.1, 6, 22))

algae.vlines <- data.frame(
  l.strain = c("~italic('C. ICEMDV')",  "~italic('C. malina')", "~italic('C. priscuii')", 
             "~italic('C. vulgaris')", "~italic('C. reinhardtii')"),
  ThermalTolerance = c("Polar", "Polar", "Polar", "Temperate", "Temperate"), 
  GrowthTemp_C = c(4.1, 4.1, 4.1, 22, 23.85))

# Computing flash spacing averages for secondary axes of heatmaps
spacing.diatom <- doBy::summaryBy(Light_Level ~ PulseSpace_s, data = subset(sstate_damping, Type == "Diatom"))
spacing.algae <- doBy::summaryBy(Light_Level ~ PulseSpace_s, data = subset(sstate_damping, Type == "Green Algae"))
```


### Significance Matrix
```{r Plot significance matrix, warning=FALSE}
mtrx.data <- sstate_damping %>%
  mutate(DampingIndex = ifelse(DampingIndex > 0, 1, 0),
  Strain = factor(Strain, levels=c("Fragilariopsis_cylindrus", "Thalassiosira_pseudonana", "Chlamydomonas_ICEMDV", 
            "Chlamydomonas_malina", "Chlamydomonas_priscuii", "Chlorella_vulgaris", "Chlamydomonas_reinhardtii")))

mtrx.data  %>%
  ggplot() +
  geom_tile(aes(x= as.factor(Temp_C), y=as.factor(PulseSpace_s), fill=as.factor(DampingIndex))) +
  scale_fill_manual(values= c("black","skyblue"), 
                      name= bquote(atop("Significance of", "4-Step Periodcity")), 
                      labels =c("p > 0.05", "p ≤ 0.05")) +
  labs(x = "Measurement Temperature (°C)", y= "Flash Spacing (s)") +
  ggh4x::facet_nested_wrap(vars(ThermalTolerance, l.strain), ncol=4, nrow=2, scales="free_x",
                             labeller = label_parsed) +
  theme_thesis() +
  theme(legend.title = element_text(size=11), strip.text = element_text(face="italic"),
        legend.position = c(0.97,0.1), legend.justification = c(1,0))

# ggsave("periodicity_mtrx.png", width = 7 , height = 5, dpi = 900,
#        plot=last_plot(),
#        path="../Output")
```

```{r}
mtrx.data  %>%
  filter(Strain == 	"Fragilariopsis_cylindrus") %>%
  ggplot() +
  geom_tile(aes(x= as.factor(Temp_C), y=as.factor(PulseSpace_s), fill=as.factor(DampingIndex))) +
  scale_fill_manual(values= c("black","skyblue"), 
                      name= bquote(atop("Significance of", "4-Step Periodcity")), 
                      labels =c("p > 0.05", "p ≤ 0.05")) +
  labs(x = "Measurement Temperature (°C)", 
       y= "Flash Spacing (s)") +
  facet_wrap(vars(l.temp), labeller = label_parsed) +
  theme_thesis() +
  theme(legend.title = element_text(size=11))

# ggsave("Fc_mtrx.png", width = 6 , height = 3, dpi = 900,
#        plot=last_plot(),
#        path="../Output")
```


### GAM By Deviation From Growth Temp

```{r}
# come back to list later 
predict.gam <- function(list, predictor) {
      
  combineddata <- lapply(list, function(damping){
  gam <- mgcv::gam(DampingIndex ~ te(get(predictor), Light_Level, k= n_distinct(damping[[predictor]])), 
                   data = damping, method = "REML")

  data <- with(damping, expand.grid( Light_Level = seq(min(Light_Level), max(Light_Level), 0.05),
    predictor = seq(min(damping[[predictor]]), max(damping[[predictor]]), 0.1)))
  
  data <- data %>%
    mutate(DampingIndex = predict(gam, newdata = data), 
           DampingIndex = ifelse(DampingIndex < 0, 0, DampingIndex),
           Strain = factor(unique(damping$Strain)),
           l.strain = unique(damping$l.strain),
           GrowthTemp_C = factor(unique(damping$GrowthTemp_C)), 
           l.temp = unique(damping$l.temp),
           ThermalTolerance = factor(unique(damping$ThermalTolerance))) 
      return(data) })
  combineddata <- do.call(bind_rows, combineddata)
  return(combineddata)
}
```


```{r}
predict.gam.d <- predict.gam(diff.list, TempDiff_C)
```






```{r Functions for GAM by Diff}
predict.diff.gam <- function(diff.list) {
  combineddata <- lapply(diff.list, function(damping){
  gam <- mgcv::gam(DampingIndex ~ te(TempDiff_C, Light_Level, k= n_distinct(damping$TempDiff_C)), 
                   data = damping, method = "REML")

  data <- with(damping, expand.grid( Light_Level = seq(min(Light_Level), max(Light_Level), 0.05),
    TempDiff_C = seq(min(TempDiff_C), max(TempDiff_C), 0.1)))
  
  data <- data %>%
    mutate(DampingIndex = predict(gam, newdata = data), 
           DampingIndex = ifelse(DampingIndex < 0, 0, DampingIndex),
           Strain = factor(unique(damping$Strain)), 
           l.strain = unique(damping$l.strain),
           ThermalTolerance = factor(unique(damping$ThermalTolerance))) 
      return(data) })
  combineddata <- do.call(bind_rows, combineddata)
  return(combineddata)
}
```


```{r Functions for GAM by Diff}
predict.gam.d <- predict.diff.gam(diff.list)
```


```{r Functions for GAM by Diff}
summary.diff.gam <- function(damping) {
  knots <- n_distinct(damping$TempDiff_C)
  gam <- mgcv::gam(DampingIndex ~ te(TempDiff_C, Light_Level, k=knots), data = damping, method = "REML")
  summary <- summary(gam)
  shapiro <- shapiro.test(resid(gam))
  k <- mgcv::k.check(gam)
  vis.fit <- mgcViz::getViz(gam)
  check <- mgcViz::check.gamf0iz(vis.fit,
                                a.qq = list(method = "tnorm", 
                                            a.cipoly = list(fill = "light blue")), 
                                a.respoi = list(size = 1), 
                                a.hist = list(bins = 4))
  combined <- list(summary = summary, norm = shapiro, check.k = k, plot.assumptions = check)
  return(combined)
}
```


```{r Create Predicted df}
predict.gam.d <- lapply(diff.sstate, predict.diff.gam)
predict.gam.d <- bind_rows(predict.gam.d)
```

```{r Generate model summaries}
summary.gam.d <- lapply(diff.sstate, summary.diff.gam)
```


##### Diatoms
```{r}
predict.gam.d %>%
  filter(Strain %in% c("Fragilariopsis_cylindrus", "Thalassiosira_pseudonana")) %>%
  ggplot(aes(x=TempDiff_C, y=Light_Level, z= DampingIndex)) +
  geom_contour_fill() + 
  geom_contour2(color = "white", linemitre = 10, linewidth=0.25) +
  geom_text_contour(stroke = 0.1, family = "Times New Roman", size = 3.5) +
  geom_vline(xintercept=0, col= "white", linetype = "dashed", linewidth = 1.3) +
  labs(x= "Deviation from Growth Temperature (°C)",
       y= expression("Light Level ("*mu*"mol photons m"^"-2"~s^-1*")"),
       fill= "Damping Index") +
  scale_fill_gradientn(colors = pals::ocean.deep(13)) +
  coord_cartesian(ylim = c(0.03, 0.53)) +
  scale_x_continuous(breaks= c(-12,-8,-4,0,4,8), expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0), sec.axis = dup_axis(
    breaks = c(spacing.diatom$Light_Level.mean), labels = c(spacing.diatom$PulseSpace_s),
    name = sym("Equivalent Flash Spacing (s)"))) +
  ggh4x::facet_nested_wrap(vars(ThermalTolerance, l.strain), ncol = 1, strip.position = "right",
                             labeller = label_parsed)+
  theme_thesis() +
  theme(strip.placement = "outside",
        panel.spacing.y = unit(4, "mm"))

# ggsave("diatoms_diff.png", width = 8.5 , height = 5.5, dpi = 900,
#        plot=last_plot(),
#        path="../Output")
```

##### Algae
```{r}
predict.gam.d %>%
  filter(Strain %in% c("Chlamydomonas_ICEMDV", "Chlamydomonas_malina", "Chlamydomonas_priscuii", 
             "Chlorella_vulgaris", "Chlamydomonas_reinhardtii")) %>%
  ggplot(aes(x=TempDiff_C, y=Light_Level, z= DampingIndex)) +
  geom_contour_fill() + 
  geom_contour2(color = "white", linemitre = 10, linewidth=0.25) +
  geom_text_contour(stroke = 0.1, family = "Times New Roman", size = 3.5) +
  geom_vline(xintercept=0, col= "white", linetype = 1, linewidth = 1.5) +  
  geom_vline(xintercept=0, col= "black", linetype = 2, linewidth = 0.9) +  
  labs(x= "Deviation from Growth Temperature (°C)",
       y= expression("Light Level ("*mu*"mol photons m"^"-2"~s^-1*")"),
       fill= "Damping Index") + 
  scale_fill_gradientn(colors = pals::ocean.deep(13), breaks = seq(0, 13, by = 3)) +
  coord_cartesian(ylim = c(0.03, 0.66)) +
  scale_x_continuous(breaks= c(-12,-8,-4,0,4,8), expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0), sec.axis = dup_axis(
    breaks = c(spacing.algae$Light_Level.mean), labels = c(spacing.algae$PulseSpace_s),
    name = sym("Equivalent Flash Spacing (s)"))) +
  ggh4x::facet_nested_wrap(vars(ThermalTolerance, l.strain), ncol = 1, strip.position = "right",
                             labeller = label_parsed)+
  theme_thesis() +
  theme(strip.placement = "outside",
        panel.spacing.y = unit(4, "mm"))


# ggsave("algae_diff.png", width = 8 , height = 10.5, dpi = 900,
#        plot=last_plot(),
#        path="../Output")
```
```{r}
sstate_damping %>%
  filter(Type== "Green Algae",
         Temp_C %in% c(10,12))
```


### GAMs By Measurement Temp 

```{r Functions for GAM by Temp}
predict.temp.gam <- function(damping) {
  
  knots <- n_distinct(damping$Temp_C)
  
  gam <- mgcv::gam(DampingIndex ~ te(Temp_C, Light_Level, k=knots), data = damping, method = "REML")

  data <- with(damping, expand.grid(
    Light_Level = seq(min(Light_Level), max(Light_Level), 0.05),
    Temp_C = seq(min(Temp_C), max(Temp_C), 0.1)
  ))
  
  data$DampingIndex <- predict(gam, newdata = data)
  
  data <- data %>%
    mutate(DampingIndex = ifelse(DampingIndex < 0, 0, DampingIndex)) %>%
    mutate(Strain = factor(unique(damping$Strain)), 
           l.strain = unique(damping$l.strain),
           GrowthTemp_C = factor(unique(damping$GrowthTemp_C)), 
           ThermalTolerance = factor(unique(damping$ThermalTolerance))) 
  
  return(data) 
}

summary.temp.gam <- function(damping) {
  knots <- n_distinct(damping$Temp_C)
  gam <- mgcv::gam(DampingIndex ~ te(Temp_C, Light_Level, k=knots), data = damping, method = "REML")
  summary <- summary(gam)
  shapiro <- shapiro.test(resid(gam))
  k <- mgcv::k.check(gam)
  vis.fit <- mgcViz::getViz(gam)
  check <- mgcViz::check.gamViz(vis.fit,
                                a.qq = list(method = "tnorm", 
                                            a.cipoly = list(fill = "light blue")), 
                                a.respoi = list(size = 1), 
                                a.hist = list(bins = 4))
  combined <- list(summary = summary, norm = shapiro, check.k = k, plot.assumptions = check)
  return(combined)
}
```

```{r}
predict.gam.t <- lapply(temp.sstate, predict.temp.gam)
predict.gam.t <- bind_rows(predict.gam.t) %>%
  mutate(l.temp = paste0("Grown~at~", GrowthTemp_C,"*degree*C"), 
         l.temp )
```

```{r}
summary.gam.t <- lapply(temp.sstate, summary.temp.gam)
```


##### Diatoms

```{r}
predict.gam.t %>%
  filter(Strain %in% c("Thalassiosira_pseudonana", "Fragilariopsis_cylindrus")) %>%
  ggplot(aes(x=Temp_C, y=Light_Level, z= DampingIndex)) +
  geom_contour_fill() + 
  geom_contour2(color = "white", linemitre = 10, linewidth=0.25) +
  geom_text_contour(stroke = 0.1, family = "Times New Roman", size = 3.3, 
                    label.placer = label_placer_fraction(frac=0.4)) +
  labs(x= "Measurement Temperature (°C)",
       y= expression("Light Level ("*mu*"mol photons m"^"-2"~s^-1*")"),
       fill= "Damping Index") +
  scale_fill_gradientn(colors = pals::ocean.matter(14), breaks = seq(0, 13, by = 3)) +
  coord_cartesian(ylim = c(0.03, 0.53)) +
  scale_x_continuous(breaks = c(2, 6, 10, 14, 18, 22, 26), expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0), sec.axis = dup_axis(
    breaks = c(spacing.diatom$Light_Level.mean), labels = c(spacing.diatom$PulseSpace_s),
    name = sym("Equivalent Flash Spacing (s)"))) +
  ggh4x::facet_nested_wrap(vars(l.strain, l.temp), ncol = 2, 
                           strip.position = "top", scales="free_x", 
                           labeller = label_parsed) +
  geom_vline(data=diatom.vlines, aes(xintercept=GrowthTemp_C), 
             col= "white", linetype = 2, linewidth = 1) +
  theme_thesis() +
  theme(panel.spacing.x = unit(4, "mm"),
        legend.position = c(0.89, 0.07), legend.justification = c(1,0))


# ggsave("diatoms_temp.png", width = 7.5 , height = 5, dpi = 900,
#        plot=last_plot(),
#        path="../Output")
```

```{r}
comp.labs <- str_wrap(c("Thalassiosira_pseudonana_22" = "Thalassiosira pseudonana",
               "Fragilariopsis_cylindrus_6" = "Fragilariopsis cylindrus grown at 6°C",
               "Fragilariopsis_cylindrus_0" = "Fragilariopsis cylindrus grown at 0°C"), width = 25)

predict.gam.t %>%
  filter(Strain %in% c("Thalassiosira_pseudonana", "Fragilariopsis_cylindrus")) %>%
  filter(Temp_C == 10) %>%
  mutate(Culture = paste(Strain, GrowthTemp_C, sep = "_")) %>%
  ggplot() +
  geom_point(aes(x= Light_Level, y=DampingIndex, col= Culture, shape= Culture )) +
  geom_path(aes(x= Light_Level, y=DampingIndex, col= Culture)) +
  labs(x = expression("Light Level ("*mu*"mol photons m"^"-2"~s^-1*")"), y = "Damping Index") +
  scale_color_manual(values=c("#009392","#91B8AB", "#DFA0A0"), labels=c(comp.labs)) +
  scale_shape_discrete(labels=c(comp.labs)) + 
  scale_x_continuous(limits = c(0, 0.66), sec.axis = dup_axis(breaks = c(spacing.diatom$Light_Level.mean), 
                labels = c(spacing.diatom$PulseSpace_s), name = sym("Equivalent Flash Spacing (s)")), 
                expand= c(0.01,0.01)) +
  scale_y_continuous(limits = c(0,12), breaks = c(2,4,6,8,10,12)) +
  theme_thesis() +
  theme(legend.key.size = unit(1, "cm"), legend.text = element_text(face = "italic"))

# ggsave("diatom_comp.png", width = 8 , height = 4, dpi = 900,
#        plot=last_plot(),
#        path="../Output")
```

##### Algae
```{r}
predict.gam.t %>%
  filter(Strain %in% c("Chlamydomonas_ICEMDV", "Chlamydomonas_malina", "Chlamydomonas_priscuii", 
             "Chlorella_vulgaris", "Chlamydomonas_reinhardtii")) %>%
  ggplot(aes(x=Temp_C, y=Light_Level, z= DampingIndex)) +
  geom_contour_fill() + 
  geom_contour2(color = "white", linemitre = 10, linewidth=0.25) +
  geom_text_contour(stroke = 0.1, family = "Times New Roman", size = 3.5, 
                    label.placer = label_placer_random()) +
  labs(x= "Measurement Temperature (°C)",
       y= expression("Light Level ("*mu*"mol photons m"^"-2"~s^-1*")"),
       fill= "Damping Index") +
  scale_fill_gradientn(colors = pals::ocean.matter(14), breaks = seq(0, 13, by = 3)) +
  coord_cartesian(ylim = c(0.03, 0.66)) +
  scale_x_continuous(breaks= c(0,4,8,12,16,20,24,28), expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0),sec.axis = dup_axis(
    breaks = c(spacing.algae$Light_Level.mean), labels = c(spacing.algae$PulseSpace_s),
    name = sym("Equivalent Flash Spacing (s)"))) +
  ggh4x::facet_nested_wrap(vars(ThermalTolerance, l.strain), ncol = 3, strip.position = "top",
                       scales = "free_x", 
                       labeller = label_parsed) +
  geom_vline(data=algae.vlines, aes(xintercept=GrowthTemp_C), 
             col= "white", linetype = 2, linewidth = 1) +
  theme_thesis() +
  theme(strip.placement = "outside",
        panel.spacing.x = unit(4, "mm"),
        legend.position = c(0.93, 0.07), legend.justification = c(1,0))

# ggsave("algae_temp.png", width = 9.3 , height = 6, dpi = 900,
#        plot=last_plot(),
#        path="../Output")
```

```{r}
algae.labs <- c("Chlamydomonas_ICEMDV" = "Chlamydomonas ICEMDV", 
                 "Chlamydomonas_malina" = "Chlamydomonas malina",
                 "Chlamydomonas_priscuii" = "Chlamydomonas priscuii", 
                 "Chlorella_vulgaris" = "Chlorella vulgaris", 
                 "Chlamydomonas_reinhardtii" = "Chlamydomonas reinhardtii")
               

predict.gam.t %>%
  filter(Strain %in% c("Chlamydomonas_ICEMDV", "Chlamydomonas_malina", "Chlamydomonas_priscuii", 
             "Chlorella_vulgaris", "Chlamydomonas_reinhardtii")) %>%
  filter(Temp_C == 12) %>%
  ggplot() +
  geom_point(aes(x= Light_Level, y=DampingIndex, col= Strain, shape= Strain)) +
  geom_path(aes(x= Light_Level, y=DampingIndex, col= Strain)) +
  labs(x = expression("Light Level ("*mu*"mol photons m"^"-2"~s^-1*")"),
       y = "Damping Index", col = "Strain", shape= "Strain") +
  scale_color_manual(values=c(rcartocolor::carto_pal(n=5,"Temps")), labels=c(algae.labs)) +
  scale_shape_manual(values=c(16,3,17,18,8), labels=c(algae.labs)) +
  scale_x_continuous(sec.axis = dup_axis(breaks = c(spacing.algae$Light_Level.mean), 
                labels = c(spacing.algae$PulseSpace_s), name = sym("Equivalent Flash Spacing (s)")), 
                expand= c(0.01,0.01)) +
  scale_y_continuous(limits = c(0,12), breaks = c(2,4,6,8,10,12)) +
  theme_thesis() +
  theme(legend.text = element_text(face="italic"))

# ggsave("algae_comp.png", width = 8 , height = 4, dpi = 900,
#        plot=last_plot(),
#        path="../Output")
```

# Evaluating Outlier in *C. malina*
```{r}
Cm_damping <- subset(sstate_damping, Strain == "Chlamydomonas_malina")

# Original model 
Cm_gam <- gam(DampingIndex ~ te(Temp_C, Light_Level, k=3), data = Cm_damping, method = "REML")
cooksd <- cooks.distance(Cm_gam)
plot(1:length(cooksd), cooksd, type = "h")

# Remove outlier point
Cm_rmoutlier <- Cm_damping %>% filter(!Light_Level == 0.07333)

# Re-run model 
Cm_rm.out <- gam(DampingIndex ~ te(Temp_C, Light_Level, k=3), data = Cm_damping, method = "REML")
summary(Cm_rm.out)

```



# For Presentation 
```{r}
theme_pres <- function() {
  theme_bw() +
  theme(text = element_text(family = "Georgia"), 
        axis.text=element_text(size=11), axis.title=element_text(size=13),
        legend.text=element_text(size=11), legend.title=element_text(size=13),
        strip.background = element_rect(colour="black", fill=NA),
        strip.text = element_text(size=11), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()
  )
}
```

```{r}
damping.group <- sstate_damping %>% 
   mutate(ThermalTolerance= ifelse(Strain %in% c("Fragilariopsis_cylindrus", 
                           "Chlamydomonas_ICEMDV","Chlamydomonas_priscuii", "Chlamydomonas_malina"),
                           "Polar", 
                           ifelse(Strain %in% c("Chlamydomonas_reinhardtii", "Chlorella_vulgaris",
                                               "Thalassiosira_pseudonana"), "Temperate", NA)), 
         Type = ifelse(Strain %in% c("Chlamydomonas_reinhardtii", "Chlorella_vulgaris",
                           "Chlamydomonas_ICEMDV","Chlamydomonas_priscuii", "Chlamydomonas_malina"),
                           "Algae", 
                           ifelse(Strain %in% c("Fragilariopsis_cylindrus", "Thalassiosira_pseudonana"),
                                  "Diatom", NA)), 
         Group = paste(ThermalTolerance, Type, sep = "_")) %>%
  doBy::splitBy(formula = ~ Group)
```

```{r}
general.group <- sstate_damping %>% 
  doBy::splitBy(formula = ~ ThermalTolerance)
```

```{r}
predict.diff.group <- function(damping) {
  
  knots <- n_distinct(damping$TempDiff_C)
  
  gam <- mgcv::gam(DampingIndex ~ te(TempDiff_C, Light_Level, k=3), data = damping)

  data <- with(damping, expand.grid(
    Light_Level = seq(0.025, 0.65, 0.01),
    TempDiff_C = seq(min(TempDiff_C), max(TempDiff_C), 0.05)
  ))
  
  data$DampingIndex <- predict(gam, newdata = data)
  
  data <- data %>%
    mutate(DampingIndex = ifelse(DampingIndex < 0, 0, DampingIndex)) %>%
    mutate(ThermalTolerance = unique(damping$ThermalTolerance))
  
  return(data) 
}
```


```{r}
predict.group <- lapply(general.group, predict.diff.group)
predict.group <- bind_rows(predict.group) 
```


```{r}
predict.group %>%
  ggplot(aes(x=TempDiff_C, y=Light_Level, z= DampingIndex)) +
  geom_contour_fill() + 
  geom_contour2(color = "white", linemitre = 10, linewidth=0.25) +
  geom_text_contour(stroke = 0.1, family = "Times New Roman", size = 3.5) +
  geom_vline(xintercept=0, col= "white", linetype = 1, linewidth = 1.3) +  
  geom_vline(xintercept=0, col= "black", linetype = 2, linewidth = 0.8) +  
  labs(x= "Deviation from Growth Temperature (°C)",
       y= expression("Light Level ("*mu*"mol photons m"^"-2"~s^-1*")"),
       fill= "Damping Index") + 
  scale_fill_gradientn(colors = pals::ocean.deep(13)) +
  scale_x_continuous(expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0)) +
  ggh4x::facet_nested_wrap(vars(ThermalTolerance), ncol = 1, strip.position = "right") +
  theme_pres() 

# ggsave("between_pres.png", width = 10 , height = 4.5, dpi = 900,
#        plot=last_plot(),
#        path="~/Downloads")
```

```{r}
predict.temp.group <- function(damping) {
  
  knots <- n_distinct(damping$Temp_C)
  
  gam <- mgcv::gam(DampingIndex ~ te(Temp_C, Light_Level, k=knots), data = damping)

  data <- with(damping, expand.grid(
    Light_Level = seq(0.025, 0.7, 0.01),
    Temp_C = seq(min(Temp_C), max(Temp_C), 0.05)
  ))
  
  data$DampingIndex <- predict(gam, newdata = data)
  
  data <- data %>%
    mutate(DampingIndex = ifelse(DampingIndex < 0, 0, DampingIndex)) %>%
    mutate(ThermalTolerance = unique(damping$ThermalTolerance))
  
  return(data) 
}
```




```{r}
predict.group.temp <- lapply(general.group, predict.temp.group)
predict.group.df <- bind_rows(predict.group.temp) 
```



```{r warning=FALSE}
predict.group.df %>%
  ggplot(aes(x=Temp_C, y=Light_Level, fill= DampingIndex)) +
  geom_tile() + 
  labs(x= "Measurement Temperature (°C)",
       y= expression("Light Level ("*mu*"mol photons m"^"-2"~s^-1*")"),
       fill= "Damping Index") +
  scale_fill_gradientn(colors = pals::ocean.matter(26600)) +
  scale_x_continuous(expand=c(0,0)) + scale_y_continuous(expand=c(0,0)) +
  facet_wrap(vars(ThermalTolerance), ncol = 2, scales= "free_x") +
  theme_pres() +
  theme(panel.spacing.x = unit(6, "mm"))

# ggsave("within_pres.png", width = 10 , height = 4.5, dpi = 900,
#        plot=last_plot(),
#        path="~/Downloads")
```
