---
title: "GAM and Figures"
author: "Mireille Savoie, Natasha Ryan"
output:
  word_document: default
  html_document: default
editor_options:
  chunk_output_type: inline
---


Introduction

This script fits Generalized Additive Models (GAMs) using mgcv::gam combined with possibGAM (from purrr::possibly), allowing model fitting without halting execution due to errors. Smooth terms (s) and interaction effects (ti) are used to capture nonlinear relationships.  gam.check is used to run model diagnostics guiding basis dimension (k) selection.

GamParaMod : is the response variable input for the GAM. It is assigned at the start of each GAM chunk.
s(CalcSal) and s(par_ue): main effects
ti(CalcSal, par_ue, k = 3): Tensor product smooth for modeling interaction between CalcSal and par_ue.

Unlike te() (main effects and interactions), ti() includes only the interaction term (not the main effects), which is appropriate here because main effects are already modeled separately using s(CalcSal) and s(par_ue).
This avoids redundancy and helps reduce overfitting.
The k parameter limits the flexibility of the interaction term. High k = more flexibility for more complex patterns, low k = simple curves.



# Setup
```{r load libraries}
library(tidyverse)
#library(metR) unused library loaded; used for meteorological data
library(mgcv)
library(ggpubr)
library(knitr)
library(broom)
# library(commonR)  Natasha defined her own theme_pres(), used here and in other .Rmd;  sloppy handling of functions, variables and environments
```


```{r define paths}
DataOut <- file.path("..", "Data", "ProcessedData")
PlotsOut <- file.path("..", "Figures")
```

# Damping Across Taxa
#add new taxa
```{r import & wrangle data}
`%notin%` <- negate(`%in%`)

# Remove Chlamydomonas_malina and Chlamydomonas_priscui grown at salinities not at or near optimal salinity of 500mM. Note: Chlamydomonas_ICEMDV optimal salinity is in theory 500mM but cultures did not grow at time of experiments; therefore culture maintenance salinity of 70 mM is considered optimal 
sstate_damping_2025 <- readRDS(file = file.path("..", "Data", "ProcessedData", "LOW_Chlamy_dampingYear.Rds")) %>%
  mutate(NaCl_opt = as.numeric(NaCl_opt)) %>%
 filter(Year != 2025 | NaCl_opt == NaCl_mM) %>% 
 filter(!(Strain == "Chlamydomonas_priscui" & NaCl_mM == 70)) %>% 
   filter(!(Strain == "Chlamydomonas_malina" & NaCl_mM == 70)) %>% 
 mutate(IsOptSalinity = case_when(NaCl_opt == NaCl_mM ~ 1,
                                  TRUE ~ 0)) %>% 
   select((-c(NaCl_mM))) 

glimpse(sstate_damping_2025)

# Remove all Chlamydomonas from this data frame as the same data are included in the 2025 dataset
sstate_damping_2023 <- readRDS(file = file.path("..", "Data", "ProcessedData", "LOW_SState_Damping.Rds" )) %>% 
  filter(!grepl("Chlamydomonas", Strain)) %>% 
  mutate(Year = as.character(2023)) %>% 
  rename(LightLevel_ue = 'Light_Level') %>% 
  mutate(IsOptSalinity = 1,
         NaCl_opt = case_when(Strain == "Thalassiosira_pseudonana" ~ 0,
                              Strain == "Chlorella_vulgaris" ~ 0,
                              Strain == "Fragilariopsis_cylindrus" ~ 500))

glimpse(sstate_damping_2023)

sstate_damping <- bind_rows (sstate_damping_2023, sstate_damping_2025)%>%
  mutate(Strain = factor(Strain),
         NaCl_opt = round(NaCl_opt, 0),
  ThermalTolerance = case_when(
     Strain %in% c("Fragilariopsis_cylindrus",
                   "Chlamydomonas_ICEMDV",
                   "Chlamydomonas_priscui",
                   "Chlamydomonas_malina",
                   "Chlamydomonas_klinobasis") ~ "Polar",
     Strain %in% c("Chlamydomonas_reinhardtii",
                   "Chlamydomonas_euryale",
                   "Chlorella_vulgaris",
                   "Thalassiosira_pseudonana") ~ "Temperate"),  
  Type = case_when(
     Strain %in% c("Fragilariopsis_cylindrus",
                   "Thalassiosira_pseudonana") ~ "Diatom", 
     Strain %in% c("Chlamydomonas_ICEMDV",
                   "Chlamydomonas_priscui", 
                   "Chlamydomonas_malina", 
                   "Chlamydomonas_reinhardtii", 
                   "Chlamydomonas_euryale", 
                   "Chlamydomonas_klinobasis", 
                   "Chlorella_vulgaris") ~ "Green Algae"),
  l.strain = case_when(
    Strain == "Chlamydomonas_ICEMDV" ~ "~italic('C. ICEMDV')", 
    Strain == "Chlamydomonas_malina" ~"~italic('C. malina')",
    Strain == "Chlamydomonas_priscui" ~ "~italic('C. priscui')",
    Strain == "Chlamydomonas_reinhardtii" ~ "~italic('C. reinhardtii')",
    Strain == "Chlorella_vulgaris" ~ "~italic('C. vulgaris')",
    Strain == "Fragilariopsis_cylindrus" ~ "~italic('F. cylindrus')", 
    Strain == "Thalassiosira_pseudonana" ~ "~italic('T. pseudonana')",
    Strain == "Chlamydomonas_klinobasis" ~ "~italic('C. klinobasis')",            
    Strain == "Chlamydomonas_euryale" ~ "~italic('C. euryale')"),
  l.temp = paste0("Grown~at~", GrowthTemp_C,"*degree*C"),
  FacetLabel_temp = paste0("grown~at~", GrowthTemp_C,"*degree*C"),
  )

```


Run GAM on 'sstate_damping' dataframe
Response variable 'DampingIndex'
Predictors are 'Temp_C' and 'LightLevel_ue'
Grouping variables are 'Strain' and GrowthTemp_C'
```{r DampingIndex GAM}

GamParaMod = "DampingIndex" 

#https://stackoverflow.com/questions/49570452/is-it-possible-to-specify-lower-bound-in-response-variable-during-smooth-with-ga 
possibGAM = possibly(.f = mgcv::gam, otherwise = NULL)
possibPRED = possibly(.f = mgcv::predict.gam, otherwise = NULL)

# Build Temp_C vector 
Temp_C_vals <- seq(0, 28, by = 0.05)

# Same for LightLevel_ue
Light_vals <- seq(0.02, 1.0, by = 0.02)

# Build the grid to fill in predicted data
PAR_Temp_Model <- expand_grid(
  LightLevel_ue = Light_vals,
  Temp_C = Temp_C_vals
) %>% 
  as_tibble()


summary(PAR_Temp_Model) 


DampingIndex_Gam <- sstate_damping %>% 
   ungroup() %>%
  group_by(Strain, GrowthTemp_C, NaCl_opt, ThermalTolerance, Type, l.strain, l.temp, FacetLabel_temp) %>%
  nest() %>%
  mutate(
    GamModel = map(data, ~possibGAM(!!sym(GamParaMod) ~  s(LightLevel_ue, k = 3) + s(Temp_C, k = 3) + ti(Temp_C, LightLevel_ue, k = 3), data = .x,
                                    family = "gaussian"(link = "log"), 
                                    method = "REML")),
    predicted = map(GamModel, ~possibPRED(.x, newdata = PAR_Temp_Model, type = "response", se.fit = TRUE)),
    tidied = map(GamModel, ~tidy(.x)),
    augmented = map(GamModel, ~augment(.x)),
    glance = map(GamModel, ~glance(.x))
  ) %>% 
  ungroup() %>% 
  mutate(StrainLabel = case_when(
    Strain == "Chlamydomonas_ICEMDV" ~ "C. ICEMDV", 
    Strain == "Chlamydomonas_malina" ~"C. malina",
    Strain == "Chlamydomonas_priscui" ~ "C. priscui",
    Strain == "Chlamydomonas_reinhardtii" ~ "C. reinhardtii",
    Strain == "Chlorella_vulgaris" ~ "C. vulgaris",
    Strain == "Fragilariopsis_cylindrus" ~ "F. cylindrus", 
    Strain == "Thalassiosira_pseudonana" ~ "T. pseudonana",
    Strain == "Chlamydomonas_klinobasis" ~ "C. klinobasis",            
    Strain == "Chlamydomonas_euryale" ~ "C. euryale"))

DampingIndex_Pred <- DampingIndex_Gam %>%
  dplyr::select(predicted, Strain, GrowthTemp_C, NaCl_opt, ThermalTolerance, Type, l.strain, l.temp, FacetLabel_temp) %>%
  dplyr::filter(predicted != "NULL") %>%
  rowwise() %>%
  mutate(db = list(cbind(predicted, PAR_Temp_Model))) %>%
  dplyr::select(-predicted) %>%
  unnest(col = db) 

# GAM model results into table
GAMStats<- DampingIndex_Gam %>% 
  dplyr::select(Strain, GrowthTemp_C, NaCl_opt, tidied) %>% 
  unnest(tidied) %>% 
  mutate(edf = round(edf, digit = 4),
         ref.df = round(ref.df, digit = 4),
         statistic = round(statistic, digit = 4),
         p.value = round(p.value, digit = 4))

saveRDS(object = GAMStats, file = file.path(DataOut, "GAMStats.Rds"), ascii = FALSE, version = NULL, refhook = NULL)

saveRDS(object = DampingIndex_Gam, file = file.path(DataOut, "DampingIndex_Gam.Rds"), ascii = FALSE, version = NULL, refhook = NULL)

```


```{r check DampingIndex GAM}

#check that EDF (Effective Degrees of Freedom)
# Note: Low p-value in gam.check() for a smooth term means k may be too small not that the term is statistically significant, but that it may be under fitting.
purrr::walk(DampingIndex_Gam$GamModel, gam.check)



# GAM CHECK RESULTS

# in s(LightLevel_ue, k = 3) , k is too low for F.c, and C.priscui; however, k = 4 returns 'NULL' for all GAMs 
# in s(CalcSal, k = 3) , k is good
# in ti(CalcSal, par_ue, k = 3) k is too low for C.priscui; however, k = 4 returns 'NULL' for all GAMs 


# explanation for manuscript:

# Basis dimensions (k) were intentionally kept low to ensure stable model fitting across all strains, given limited sample sizes within some strain subsets. Although gam.check() suggested that higher basis dimensions may be warranted for some smooth terms in some strains (low k-index and significant p.values), increasing k resulted in non-convergence or unstable fits for the majority of strains. Therefore, low k values were used to prioritize model stability and comparability among strains.
```

```{r gammodeldiagnostic,  fig.cap = "GAM model diagnostic summary for polar and temperate diatoms, of consecutive flashes before damping of SState induced chlorophyll fluorescence oscillations. GAM model predicted by the temperature (°C) imposed during measurements and the effective light level (µmol photons m^-2^s^-1^, estimated from flash spacings in seconds and 𝛔~PSII~).", out.height =  "100%", out.width = "100%", align = 'c', echo = FALSE, warnings = FALSE, comment = NA, results='asis'}

DampingIndex_Gam <- readRDS(file = file.path("..", "Data", "ProcessedData", "DampingIndex_Gam.RDS"))


DampingIndex_Gam$GamModel |>
  rlang::set_names(
    paste0(
      "*", DampingIndex_Gam$StrainLabel, "*",
      ", Growth Temperature ",
      DampingIndex_Gam$GrowthTemp_C, " °C",
      ", Salinity ",
      DampingIndex_Gam$NaCl_opt, " mM"
    )
  ) |>
  purrr::iwalk(~ {
    cat("\n\n**Strain:** ", .y, sep = "")
    mgcv::gam.check(.x)
  })



```


```{r save rds}

 saveRDS(object = DampingIndex_Pred, file = file.path(DataOut, "DampingIndex_GAMPred.Rds"), ascii = FALSE, version = NULL, refhook = NULL)

```

