---
title: "GAM and Figures"
author: "Mireille Savoie, Natasha Ryan"
output: 
  html_document: default
editor_options: 
  chunk_output_type: inline
---


Introduction

This script generates a strain-specific contour plot ("hypervolume")
to visualize the predicted damping index response across measured temperature and equivalent
light conditions. It uses a Generalized Additive Model (GAM)
output generated from 'GAM_MS.Rmd' to create smooth interpolations and overlays contour lines
at key quantiles (10%, 25%, 50%, 75%, and 90%) of the predicted damping index.

Key points:
- The script is meant to be re-run for each strain individually by
  selecting the 'GamStrain' variable and selecting the corresponding 'scale_x_continuous'
  axis limits to represent each strain's experimental range.
- Filters out predictions with high relative uncertainty (SE/fit > ErrorTol).
- Computes quantile contour levels dynamically from predicted damping index values.
- Plots a color-coded of the predicted damping index (fit),
  overlaid with white contour lines at quantile breaks.

Note: Individual strain ggplot generations is necessary as faceting is not possible because each strain has different temperature ranges; multiple strains cannot be plotted together on shared axes.

# Setup
```{r load libraries}
library(tidyverse)
library(metR)
library(knitr)
# library(broom) # unused library loaded
library(cowplot)  # for get_legend() necessary for combining plots and use ggdraw
library(grid) # for textGrob necessary for combining axis

# library(commonR)  Natasha defined her own theme_pres(), used here and in other .Rmd;  sloppy handling of functions, variables and environments
```

```{r define paths}

PlotsOut <- file.path("..", "Figures")
```

```{r create ggplot theme}
theme_thesis <- function() {
  ggplot2::theme_bw() +
  ggplot2::theme(text = element_text(family = "sans"), 
        axis.text=element_text(size=11), axis.title=element_text(size=13),
        legend.text=element_text(size=11), legend.title=element_text(size=13),
        strip.background = element_rect(colour="black", fill=NA),
        strip.text = element_text(size=11), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.title.y.right = element_text(size = 12), 
        axis.title.x.top = element_text(size = 12)
  )}
```



```{r import GAM data}

DampingIndex_Pred <- readRDS(file = file.path("..", "Data", "ProcessedData", "DampingIndex_GAMPred.Rds" )) 

head(DampingIndex_Pred)

```

# Damping Across Taxa

```{r import & wrangle sstate data}
sstate_damping_2025 <- readRDS(file = file.path("..", "Data", "ProcessedData", "LOW_SState_Damping_2025.Rds" ))

sstate_damping_pre2025 <- readRDS(file = file.path("..", "Data", "ProcessedData", "LOW_SState_Damping.Rds" )) %>% 
  mutate(NaCl_mM = 1) %>% 
  rename(LightLevel_ue = 'Light_Level')

sstate_damping <- bind_rows (sstate_damping_pre2025, sstate_damping_2025)%>%
  mutate(Strain = factor(Strain), 
  ThermalTolerance = case_when(
     Strain %in% c("Fragilariopsis_cylindrus",
                   "Chlamydomonas_ICEMDV",
                   "Chlamydomonas_priscuii",
                   "Chlamydomonas_malina",
                   "Chlamydomonas_klinobasis") ~ "Polar",
     Strain %in% c("Chlamydomonas_reinhardtii",
                   "Chlamydomonas_euryale",
                   "Chlorella_vulgaris",
                   "Thalassiosira_pseudonana") ~ "Temperate"),  
  Type = case_when(
     Strain %in% c("Fragilariopsis_cylindrus",
                   "Thalassiosira_pseudonana") ~ "Diatom", 
     Strain %in% c("Chlamydomonas_ICEMDV",
                   "Chlamydomonas_priscuii", 
                   "Chlamydomonas_malina", 
                   "Chlamydomonas_reinhardtii", 
                   "Chlamydomonas_euryale", 
                   "Chlamydomonas_klinobasis", 
                   "Chlorella_vulgaris") ~ "Green Algae"),
  l.strain = case_when(
    Strain == "Chlamydomonas_ICEMDV" ~ "~italic('C. ICEMDV')", 
    Strain == "Chlamydomonas_malina" ~"~italic('C. malina')",
    Strain == "Chlamydomonas_priscuii" ~ "~italic('C. priscuii')",
    Strain == "Chlamydomonas_reinhardtii" ~ "~italic('C. reinhardtii')",
    Strain == "Chlorella_vulgaris" ~ "~italic('C. vulgaris')",
    Strain == "Fragilariopsis_cylindrus" ~ "~italic('F. cylindrus')", 
    Strain == "Thalassiosira_pseudonana" ~ "~italic('T. pseudonana')",
    Strain == "Chlamydomonas_klinobasis" ~ "~italic('C. klinobasis')",            Strain == "Chlamydomonas_euryale" ~ "~italic('C. euryale')"),
  l.temp = paste0("Grown~at~", GrowthTemp_C,"*degree*C"),
  FacetLabel_temp = paste0("grown~at~", GrowthTemp_C,"*degree*C"),
  )

```


```{r Compute flash spacing}

# Computing flash spacing averages for secondary axes of heatmaps for algae
spacing.algae <- sstate_damping %>%
  filter(Type == "Green Algae") %>%
  group_by(PulseSpace_s) %>%
  summarise(
    LightLevel_ue.mean = mean(LightLevel_ue, na.rm = TRUE),
    LightLevel_ue.sd = sd(LightLevel_ue, na.rm = TRUE),
    .groups = "drop"
  )

# Computing flash spacing averages for secondary axes of heatmaps for diatoms
spacing.diatom <- sstate_damping %>%
  filter(Type == "Diatom") %>%
  group_by(PulseSpace_s) %>%
  summarise(
    LightLevel_ue.mean = mean(LightLevel_ue, na.rm = TRUE),
    LightLevel_ue.sd = sd(LightLevel_ue, na.rm = TRUE),
    .groups = "drop"
  )
```



```{r Plot Algae DampingIndex GAM}

# Run this chunk each time for each strain to generate strain specific hypervolume by changing the value for Gamstrain and selecting the corresponding "scale_x_continuous" in ggplot below. 
# Example scale_x_continuous(limits = c(10, 26), breaks = seq(10, 26, by = 4), expand = c(0, 0)) + # For C.vulgaris 

# plotting all strains and adding a facet-wrap does not work as each strain has different x-axis ranges.
GamStrain = "Chlamydomonas_euryale"
#Chlamydomonas_ICEMDV                     #remember to change scale_x_continuous
#Chlamydomonas_malina                     #remember to change scale_x_continuous                  
#Chlamydomonas_priscuii                   #remember to change scale_x_continuous
#Chlorella_vulgaris                       #remember to change scale_x_continuous
#Chlamydomonas_reinhardtii                #remember to change scale_x_continuous
#Chlamydomonas_klinobasis                 #remember to change scale_x_continuous
#Chlamydomonas_euryale                    #remember to change scale_x_continuous

ErrorTol = 170


DampingIndex_Hyper <- DampingIndex_Pred %>% 
  filter(Strain %in% GamStrain) %>%
  group_by(Strain, GrowthTemp_C, NaCl_mM) %>% 
  mutate(IsFitGood = case_when(se.fit/fit*100 < ErrorTol ~ (1),#Assigns '0' to faulty data where Standarderror larger +/- Errortolerance
                    TRUE ~ (0))) %>% 
  filter(IsFitGood == 1) %>%  # filters faulty data where Standarderror larger +/- Errortolerance
  ungroup()
 
       
maxDI <- DampingIndex_Hyper %>%
  filter(fit == max(fit, na.rm = TRUE))
maxDI <- maxDI$fit

minDI <- DampingIndex_Hyper %>%
  filter(fit == min(fit, na.rm = TRUE))
minDI <- minDI$fit


ContourBreaks <- DampingIndex_Hyper %>%
  group_by(Strain) %>% 
  mutate(upper10 = quantile(fit, 0.9, na.rm = TRUE),
         upper25 = quantile(fit, 0.75, na.rm = TRUE),
         median50 = quantile(fit, 0.5, na.rm = TRUE),
         lower25 = quantile(fit, 0.25, na.rm = TRUE),
         lower10 = quantile(fit, 0.1, na.rm = TRUE)) %>%
  ungroup() %>%
  dplyr::select(Strain, upper10, upper25, median50, lower25, lower10) %>%
  distinct() %>%
  pivot_longer(cols = c(upper10, upper25, median50, lower25, lower10), names_to = 'quantile', values_to = 'quantileValue') %>%
  mutate(quantileValue = round(quantileValue, digits = 1))


###plot
#https://stackoverflow.com/questions/28775036/ggplot-line-graph-with-na-values
#https://stackoverflow.com/questions/6906661/ggplot2-make-missing-value-in-geom-tile-not-blank
#https://stackoverflow.com/questions/24822621/multiple-legends-in-a-plot-with-geom-tile
#https://stackoverflow.com/questions/43395219/plotting-multiple-layers-with-geom-raster-or-geom-tile-or-geom-rect
#https://stackoverflow.com/questions/59949635/ggplot-align-grid-lines-with-end-of-geom-tile
#https://stackoverflow.com/questions/22945651/remove-space-between-plotted-data-and-the-axes
#https://stackoverflow.com/questions/5293715/how-to-use-greek-symbols-in-ggplot2


ContourLines <- ContourBreaks 
ContourLines <- ContourLines$quantileValue 

(ContourDI <- ggplot(na.omit(DampingIndex_Hyper,spacing.algae), aes(Temp_C, LightLevel_ue, z = fit, fill = fit)) +
  geom_tile() +
 scale_fill_gradientn(name = expression(atop("Damping", "Index")), colors = pals::ocean.matter(14), limits  = range(0, 16.5), na.value = "pink") +  # Range is the damping index range plotted. 
    metR::geom_contour2(aes(label = stat(level)),
                        breaks = ContourLines, # can replace 'ContourLines' with distinct breaks that match the legend ex breaks = c(0, 4, 8, 12, 16)
                        skip = 0,
                        color = "white",
                        linemitre = 10,
                        linewidth = 0.5,
                        stroke = 0.5,
                        size = 3.3,
                        label.placer = label_placer_fraction(frac = 0.4)) +
    # geom_text_contour(breaks = ContourLines, stroke = 0.1, size = 3.3,
    #               label.placer = label_placer_fraction(frac = 0.4)) +
    # scale_x_continuous(limits = c(10, 26), breaks = seq(10, 26, by = 4), expand = c(0, 0)) + # For C.vulgaris
   # scale_x_continuous(limits = c(4, 12), breaks = seq(4, 12, by = 4), expand = c(0, 0)) + # For C.ICE, C.malina & C.priscuii
       # scale_x_continuous(limits = c(12, 24), breaks = seq(12, 24, by = 4), expand = c(0, 0)) + # For C.reinhardtii
      # scale_x_continuous(limits = c(0, 12), breaks = seq(0, 12, by = 4), expand = c(0, 0)) + # For C.klinobasis
     scale_x_continuous(limits = c(16, 28), breaks = seq(16, 28, by = 4), expand = c(0, 0)) + #  For Chlamydomonas_euryale
    scale_y_continuous(limits = c(0, 0.7), breaks = seq(0.2, 0.6, by = 0.2), expand=c(0,0), 
                       sec.axis = dup_axis(breaks = c(spacing.algae$LightLevel_ue.mean), 
                                           labels = c(spacing.algae$PulseSpace_s),
                                           name = sym("Flash Spacing (s)"))) +
    coord_cartesian(expand = FALSE, clip = "on") +
    # labs(title = GamStrain) +
    xlab("Measurement Temperature (°C)") +
    ylab("Light Level (µmol photons"~ m^-2~ ~s^-1~ ")") +
    # ylab(NULL) +
  ggh4x::facet_nested_wrap(vars(l.strain, NaCl_mM), nrow=1, scales="free_x",
                             labeller = label_parsed) +
  geom_vline(aes(xintercept = GrowthTemp_C + 0.025), 
             col= "black", linetype = 2, linewidth = 1.7)  +
  # # ggh4x::force_panelsizes(cols = c(10, 8, 8, 8, 18, 16, 12), respect = TRUE, total_width = unit(12, "cm")) +
  theme_thesis()+
    theme(legend.position = "right") +
   theme(panel.spacing.x = unit(1, "mm")))
        
        # ,
        # legend.position = c(0.89, 0.07), legend.justification = c(1,0)))

ContourDI

# Dynamically name the plot with strain variable, 'ti' is the GAM parameter that generated the data which modeled the interactions of 'measurement temperature' and 'Equivalent PAR' as opposed to 'te' which includes the interactions and main effects of the predictors
assign(paste0("ti_plot_", GamStrain), ContourDI)

```

#Combine Algae plots into 1 figure and save
```{r Combine and save Algae Damping Index GAMs}

Filename <- "AlgaeGAM"


# Remove individual legends and axes titles
p1 <- ti_plot_Chlamydomonas_ICEMDV + theme(legend.position = "none",
    axis.title.y.right = element_blank()) +
  labs(x = NULL, y = NULL)
p2 <- ti_plot_Chlamydomonas_malina + theme(legend.position = "none",
    axis.title.y.right = element_blank()) +
  labs(x = NULL, y = NULL)
p3 <- ti_plot_Chlamydomonas_priscuii+ theme(legend.position = "none",
    axis.title.y.right = element_blank()) +
  labs(x = NULL, y = NULL)
p4 <- ti_plot_Chlamydomonas_reinhardtii + theme(legend.position = "none",
    axis.title.y.right = element_blank()) +
  labs(x = NULL, y = NULL) 
p5 <- ti_plot_Chlorella_vulgaris + theme(legend.position = "none",
    axis.title.y.right = element_blank()) +
  labs(x = NULL, y = NULL)



# Extract vertical legend from p1
legend_grob <- get_legend(
  p1 + theme(legend.position = "right", legend.direction = "vertical")
)

legend_only <- gridExtra::arrangeGrob(legend_grob)

# Define width of each plot based on °C span
temp_widths <- c(8, 8, 8, 12, 16)

# Normalize widths
rel_widths <- c(1, 1, 1, 1.5, 2)

top_row <- cowplot::plot_grid(p1, p2, p3, nrow = 1, rel_widths = rel_widths[1:3])
bottom_row <- cowplot::plot_grid(
  p4, p5, legend_only,
  nrow = 1,
  rel_widths = c(rel_widths[4], rel_widths[5], 0.1)  # Adjust legend width if needed to get bottom row x-axis to near same scale as top row x-axis
)

CombineAlgae <- cowplot::plot_grid(
  top_row,
  bottom_row,
  ncol = 1,
  rel_heights = c(1, 1)
)
CombineAlgae


# full layout with consolidated axes titles
# ggdraw uses the 'cowplot' package
AlgaeGAM <- ggdraw() +
  draw_plot(CombineAlgae, x = 0.1, y = 0.1, width = 0.85, height = 0.85) +
  draw_label("Measurement Temperature (°C)", x = 0.5, y = 0.04, size = 20) +
  draw_label("Equivalent PAR (µmol photons"~ m^-2~ ~s^-1~ ")", x = 0.05, y = 0.5, angle = 90, size = 20) +
  draw_label("Flash Spacing (s)", x = 0.97, y = 0.5, angle = 270, size = 20)
    
AlgaeGAM

# must use cowplot to save as x and y axis are lost using ggsave
cowplot::save_plot(
  filename = file.path(PlotsOut, paste0(Filename, ".png")),
  plot = AlgaeGAM,
  base_height = 10,
  base_width = 15,
  dpi = 300,
  bg ="white" # need a white background for axis visibility
)
```


### Plot individual Diatoms
```{r Plot Diatoms DampingIndex GAM}

# Run this chunk each time for each strain to generate strain specific hypervolume by changing the value for Gamstrain, selecting the corresponding "scale_x_continuous" in ggplot below and selecting the growth temp in GamTemp. 
#Example scale_x_continuous(limits = c(10, 28), breaks = seq(10, 28, by = 4), expand = c(0, 0)) + # For T.pseudonana GamTemp = 22

# select growth temp; F.c was grown at 0 and 6°C and T.p was grown at 22°C


# plotting all strains and adding a facet-wrap does not work as each strain has different x-axis ranges.

 GamStrain = "Fragilariopsis_cylindrus"
#Fragilariopsis_cylindrus                     #remember to change scale_x_continuous
#Thalassiosira_pseudonana                     #remember to change scale_x_continuous

# select growth temp; F.c was grown at 0 and 6°C and T.p was grown at 22°C
GamTemp = 6 #0 #6 #22 for T.p

ErrorTol = 310


DampingIndex_Hyper <- DampingIndex_Pred %>% 
  filter(Strain %in% GamStrain) %>%
  filter(GrowthTemp_C %in% GamTemp) %>% 
  group_by(Strain, GrowthTemp_C) %>% 
  mutate(IsFitGood = case_when(se.fit/fit*100 < ErrorTol ~ (1),#Assigns '0' to faulty data where Standarderror larger +/- Errortolerance
                    TRUE ~ (0))) %>% 
  filter(IsFitGood == 1) %>%  # filters faulty data where Standarderror larger +/- Errortolerance
  mutate(FacetLabel = paste(l.strain, FacetLabel_temp, sep = " ~ ")) %>% 
  ungroup()
 
       
maxDI <- DampingIndex_Hyper %>%
  filter(fit == max(fit, na.rm = TRUE))
maxDI <- maxDI$fit

minDI <- DampingIndex_Hyper %>%
  filter(fit == min(fit, na.rm = TRUE))
minDI <- minDI$fit


ContourBreaks <- DampingIndex_Hyper %>%
  group_by(Strain) %>% 
  mutate(upper10 = quantile(fit, 0.9, na.rm = TRUE),
         upper25 = quantile(fit, 0.75, na.rm = TRUE),
         median50 = quantile(fit, 0.5, na.rm = TRUE),
         lower25 = quantile(fit, 0.25, na.rm = TRUE),
         lower10 = quantile(fit, 0.1, na.rm = TRUE)) %>%
  ungroup() %>%
  dplyr::select(Strain, upper10, upper25, median50, lower25, lower10) %>%
  distinct() %>%
  pivot_longer(cols = c(upper10, upper25, median50, lower25, lower10), names_to = 'quantile', values_to = 'quantileValue') %>%
  mutate(quantileValue = round(quantileValue, digits = 1))


###plot
#https://stackoverflow.com/questions/28775036/ggplot-line-graph-with-na-values
#https://stackoverflow.com/questions/6906661/ggplot2-make-missing-value-in-geom-tile-not-blank
#https://stackoverflow.com/questions/24822621/multiple-legends-in-a-plot-with-geom-tile
#https://stackoverflow.com/questions/43395219/plotting-multiple-layers-with-geom-raster-or-geom-tile-or-geom-rect
#https://stackoverflow.com/questions/59949635/ggplot-align-grid-lines-with-end-of-geom-tile
#https://stackoverflow.com/questions/22945651/remove-space-between-plotted-data-and-the-axes
#https://stackoverflow.com/questions/5293715/how-to-use-greek-symbols-in-ggplot2


ContourLines <- ContourBreaks 
ContourLines <- ContourLines$quantileValue 

(ContourDI <- ggplot(na.omit(DampingIndex_Hyper, spacing.diatom), aes(Temp_C, LightLevel_ue, z = fit, fill = fit)) +
  geom_tile() +
  scale_fill_gradientn(name = expression(atop("Damping", "Index")), colors = pals::ocean.matter(14), limits  = range(0, 16.5), na.value = "pink") +  # Range is the damping index range plotted. 
    metR::geom_contour2(aes(label = stat(level)),
                        breaks = ContourLines, # can replace 'ContourLines' with distinct breaks that match the legend ex breaks = c(0, 4, 8, 12, 16)
                        skip = 0,
                        color = "white",
                        linemitre = 10,
                        linewidth = 0.5,
                        stroke = 0.5,
                        size = 3.3,
                        label.placer = label_placer_fraction(frac = 0.4)) +
    # geom_text_contour(breaks = ContourLines, stroke = 0.1, size = 3.3,
    #               label.placer = label_placer_fraction(frac = 0.4)) +
    # scale_x_continuous(limits = c(10, 28), breaks = seq(10, 28, by = 4), expand = c(0, 0)) + # For T.pseudonana
    scale_x_continuous(limits = c(0, 10), breaks = seq(2, 10, by = 4), expand = c(0, 0)) + # For F.c
    scale_y_continuous(limits = c(0, 0.7), breaks = seq(0.2, 0.6, by = 0.2), expand=c(0,0), 
                       sec.axis = dup_axis(breaks = c(spacing.diatom$LightLevel_ue.mean), 
                                           labels = c(spacing.diatom$PulseSpace_s),
                                           name = sym("Flash Spacing (s)"))) +
 coord_cartesian(expand = FALSE, clip = "on") +
  #   geom_text(aes(x = 0.1, y = 0.60, label = paste("Grown at", GrowthTemp_C, "°C", sep = " "), 
  #                 show.legend = FALSE),
  #   color = "white", 
  #   hjust = 0,
  #   vjust = 1,
  #   size = 5.0
  # )  +
    xlab("Measurement Temperature (°C)") +
    ylab("Equivalent PAR (µmol photons"~ m^-2~ ~s^-1~ ")") +
    # ylab(NULL) +
  ggh4x::facet_nested_wrap(vars(FacetLabel), nrow=1, scales="free_x",
                             labeller = label_parsed) +
  geom_vline(aes(xintercept = GrowthTemp_C + 0.025),
             col= "black", linetype = 2, linewidth = 1.7) +
  # ggh4x::force_panelsizes(cols = c(10, 8, 8, 8, 18, 16, 12), respect = TRUE, total_width = unit(12, "cm")) +
  theme_thesis()+
    theme(legend.position = "right")  +
  theme(panel.spacing.x = unit(4, "mm")))

ContourDI

# Dynamically name the plot with strain variable, ti is the GAM parameter that generated the data which modeled the interactions of 'measurement temperature' and 'Equivalent PAR' as opposed to te which includes the interactions and main effects of the predictors
assign(paste0("ti_plot_", GamStrain, GamTemp), ContourDI)

```


#Combine Diatom plots into 1 figure and save
```{r Combine and save Diatom Damping Index GAM plots}

Filename <- "DiatomsGAM"

# Remove individual legends and axes titles
p1 <- ti_plot_Fragilariopsis_cylindrus0 + theme(legend.position = "none",
    axis.title.y.right = element_blank()) +
  labs(x = NULL, y = NULL)
p2 <- ti_plot_Fragilariopsis_cylindrus6 + theme(legend.position = "none",
    axis.title.y.right = element_blank()) +
  labs(x = NULL, y = NULL)
p3 <- ti_plot_Thalassiosira_pseudonana22 + theme(legend.position = "none",
    axis.title.y.right = element_blank()) +
  labs(x = NULL, y = NULL)


# Extract vertical legend from p1
legend_grob <- get_legend(
  p1 + theme(legend.position = "right", legend.direction = "vertical")
)


legend_only <- gridExtra::arrangeGrob(legend_grob)

# Define width of each plot based on °C span
temp_widths <- c(10, 10, 18)

# Normalize widths
rel_widths <- c(1, 1, 1.8)

top_row <- cowplot::plot_grid(p1, p2, nrow = 1, rel_widths = rel_widths[1:2])
bottom_row <- cowplot::plot_grid(
  p3, legend_only,
  nrow = 1,
  rel_widths = c(rel_widths[3], 0.2)  # Adjust legend width if needed to get bottom row x-axis to near same scale as top row x-axis
)

CombineDiatoms <- cowplot::plot_grid(
  top_row,
  bottom_row,
  ncol = 1,
  rel_heights = c(1, 1)
)
CombineDiatoms


# full layout with consolidated axes titles
# ggdraw uses the 'cowplot' package
DiatomGAM <- ggdraw() +
  draw_plot(CombineDiatoms, x = 0.1, y = 0.1, width = 0.85, height = 0.85) +
  draw_label("Measurement Temperature (°C)", x = 0.5, y = 0.04, size = 20) +
  draw_label("Equivalent PAR (µmol photons"~ m^-2~ ~s^-1~ ")", x = 0.05, y = 0.5, angle = 90, size = 20) +
  draw_label("Flash Spacing (s)", x = 0.97, y = 0.5, angle = 270, size = 20)
    
DiatomGAM

# must use cowplot to save as x and y axis are lost using ggsave
cowplot::save_plot(
  filename = file.path(PlotsOut, paste0(Filename, ".png")),
  plot = DiatomGAM,
  base_height = 10,
  base_width = 15,
  dpi = 300,
  bg ="white" # need a white background for axis visibility
)
```


