---
title: "OttawaImport"
author:
- Natasha Ryan
- Douglas A. Campbell
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
    code_folding: hide
    keep_md: yes
    fig_caption: yes
    toc: TRUE
    toc_float: TRUE   
csl: plos-one.csl
editor_options: 
  chunk_output_type: inline
---

*Some of the code used to create this R Notebook was refurbished from "PlateAbImport.Rmd" written by Maximilian Berthold, Douglas A. Campbell, Melissa L. Rioux, Sarah J Gore, and Alyson MacCormack.*

```{r load libraries}
library(tidyverse)
library(lubridate)
library(googlesheets4)
library(googledrive)
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_chunk$set(fig.path='Output/')
```

#Setup

### Set Project Variables 
```{r set project variables}
Project <- "LOW"
TargetData <- "UOttawa"
DataOut <- file.path("..","Data", "CleanData")
CalibData <- file.path("..","Data", "CalibData")


DataIn <- file.path("..","Data", "RawData", TargetData, fsep = .Platform$file.sep)
DataRefDecay <- file.path("..","Data", "RawData",TargetData, "DecayRef", fsep = .Platform$file.sep)
DataRLC <- file.path("..","Data", "RawData",TargetData, "RLC", fsep = .Platform$file.sep)

FileID <- "fit"

FileEncode <- "UTF-8" #guess_encoding says 'ASCII"

Delimiter <- ","

HeaderRows <- 0

#conversions
# us_s = 1000000
photons_umol = 6.022E17
A2_m2 = 1E20
```


### Metadata Import
```{r import metadata}
googlesheets4::gs4_deauth()

OttawaMetadata <- read_sheet("https://docs.google.com/spreadsheets/d/1m8Pjwj4DL10AwdsQo2G3boymSWG1Oq-gJDzjU0csRKs/edit?usp=sharing")
```


### Create Functions
```{r warning=FALSE, message=FALSE, echo=FALSE}
read.delim_plus <- function(flnm, file_encode, delimiter, header_rows){read.delim(flnm, fileEncoding = file_encode, sep = delimiter,  skip = header_rows, row.names = NULL) %>% mutate(Filename = flnm)
}

conversion <- function(data){
  data %>%
    group_by(PulseSpace_s, Temp_C) %>%
    mutate(Light_Level = (1/PulseSpace_s) / (mean(Sig) / A2_m2) / photons_umol) %>%
    mutate(Light_Level = round(Light_Level, 5)) %>%
    ungroup()
}
```


# No Decay Data

### Import
```{r list PSI files for file import}
SolisenseFiles <- list.files(path = DataIn, pattern = FileID, full.names = TRUE)
SolisenseFiles

unique(duplicated(SolisenseFiles))
```

```{r read Solisense files}
SolFits <- SolisenseFiles %>%
  map_df(~read.delim_plus(flnm =., file_encode = FileEncode, delimiter = Delimiter, header_rows = HeaderRows)) #purrr::map to read all files
```

### Tidy 
```{r tidy dataset}
SolFitsTrim <- SolFits %>%
  filter(!grepl("----", DATE)) %>%
  select(-c("RFID_User_Data", "Barcode_Data","PIF", "LEDSel", "Dur.Wat", "Lon", "Lat", "GPS_stat","X","Light_1", "Light_2", "Light_3","Light_4","Light_5","Light_6")) %>%
  mutate(Filename = str_remove(string = Filename, pattern = "_fit.csv")) %>%
  mutate(Filename = str_remove(string = Filename, pattern = "../Data/RawData/UOttawa/")) %>%
  separate(Filename, into=c("YYYYMMDD", "CultureID", "Temp_C", "PulseSpace_s"), sep = "([\\/\\_])", remove = FALSE) %>%
  mutate(YYYYMMDD = lubridate::ymd(YYYYMMDD),
          TIME = as.character(TIME)) %>%
   rename(ObsDate = DATE,
         ObsTime = TIME,
         FvFm = "Fv.Fm") %>%
  mutate(Temp_C = str_remove(Temp_C, "C")) %>%
  type_convert() %>%
  mutate(PulseSpace_s = str_remove(PulseSpace_s, "s")) %>%
  mutate(PulseSpace_s = as.numeric(PulseSpace_s)) %>%
  mutate(ObsDateTime = lubridate::ymd_hms(paste(ObsDate, ObsTime))) %>%
  relocate(ObsDateTime, .after = ObsTime) %>%
  relocate(CultureID, .before = ObsDate) %>%
  type_convert()

SolFitsTrim <-SolFitsTrim %>%
  group_by(Filename) %>%
  mutate(Flashnumber = row_number()) %>%
  relocate(Flashnumber, .before=p) %>%
  mutate(Etime_s = (as.numeric(Time__mSec) - min(as.numeric(Time__mSec), na.rm = TRUE))/1000, .before=Flashnumber) %>%
  ungroup()
```


```{r correct file issues}
ProblemFiles <- unique(SolFitsTrim$Filename[SolFitsTrim$Flashnumber >32])

ProblemList <- SolFitsTrim %>% 
  filter(Filename %in% c(ProblemFiles)) %>% 
  group_by(Filename) %>%
  mutate(TimeDiff = c(NA, diff(ObsTime))) %>% #shows actual spacing between flashes
  ungroup() %>%
  doBy::split_by("Filename") #creates list of problem files
```

From ProblemList, use spacing between flashes to decipher runs:
Keep last 32 flashes in:
  20230918_MaPo1001_4C_16s        - 33 flashes 
  20230919_MaPo1004_4C_8s         - 33 flashes 
  20230918_MaPo1002_8C_16s        - 36 flashes 
  20230919_MaPo1004_8C_16s        - 36 flashes 
  20230919_MaPo1001_4C_16s        - 48 flashes
  20230918_MaPo1002_4C_16s        - 33 flashes
  20230920_MaPo1005_12C_16s       - 37 flashes
  20230919_MaPo1002_12C_8s        - 33 flashes
  20230918_MaPo1001_12C_16s       - triple, contains 3 runs, 1-32 = 8s, 33-64 = 2s, and 65-96 = 16s
  
Other adjustments: 
  20230918_MaPo1001_4C_1s         - double, keep first run
  20230920_MaPo1005_24C_16s       - keep 3-33


```{r manually trim data}
last32 <- SolFitsTrim %>%
  filter(!Filename %in% c("20230918_MaPo1001_4C_1s", "20230920_MaPo1005_24C_16s")) %>%
  group_by(Filename) %>%
  slice(tail(row_number(),32)) %>% 
  mutate(Flashnumber = row_number()) %>%
  ungroup()

MaPo1001_4C_1s <- SolFitsTrim %>%
  filter(Filename== "20230918_MaPo1001_4C_1s") %>%
  filter(Flashnumber <= 32)

MaPo1005_24C_16s <- SolFitsTrim %>%
  filter(Filename== "20230920_MaPo1005_24C_16s") %>%
  filter(Flashnumber %in% c(3:33)) %>%
  mutate(Flashnumber = row_number())

SolFitsTrim2 <- rbind(last32, MaPo1001_4C_1s, MaPo1005_24C_16s)
```


### Normalize Fluorescence & Add Temperature Differential Column 

```{r}
SolFitsTrim2 <- SolFitsTrim2 %>%
  mutate(SState = case_when(Flashnumber %% 4 == 1 ~ 1,
                    Flashnumber %% 4 == 2 ~ 2,
                    Flashnumber %% 4 == 3 ~ 3, 
                    Flashnumber %% 4 == 0 ~ 4)) #add column for sstate


#normalize values to final stretch of 10 values from 'randomized' PSII pool, done after cleaning to ensure the last 10 values are correct 
SolFitsTrim2 <- SolFitsTrim2 %>%
  group_by(Filename) %>%
  mutate(Sigrunnorm =  Sig/mean(Sig[(length(Sig) - 10):length(Sig)], na.rm = TRUE),
         Forunnorm = Fo/mean(Fo[(length(Fo) - 10):length(Fo)], na.rm = TRUE),
         Fmrunnorm = Fm/mean(Fm[(length(Fm) - 10):length(Fm)], na.rm = TRUE),
         FvFmrunnorm = FvFm/mean(FvFm[(length(FvFm) - 10):length(FvFm)], na.rm = TRUE)) %>%
  ungroup()


# Add columns for growth temperature and difference between measurement & growth temps
SolFitsTrim2 <- SolFitsTrim2 %>%
  mutate(GrowthTemp_C = 
           ifelse(CultureID %in% c("MaPo1001", "MaPo1002", "MaPo1003", "MaPo1004"), 4,
           ifelse(CultureID %in% c("MaPo1005", "MaPo1006"), 24,
                  NA))) %>%
  mutate(TempDiff_C = Temp_C - GrowthTemp_C)
```


### Group by Taxon 
```{r}
SolFitsPriscuii <- SolFitsTrim2 %>%
  filter(CultureID %in% c("MaPo1001", "MaPo1004"))

SolFitsICEMDV <- SolFitsTrim2 %>%
  filter(CultureID %in% "MaPo1002")

SolFitsMalina <- SolFitsTrim2 %>%
  filter(CultureID %in% "MaPo1003")

SolFitsReinhardtii <- SolFitsTrim2 %>%
  filter(CultureID %in% c("MaPo1005", "MaPo1006"))
```


### Calculate Light Levels
```{r}
# add column for light level (µmol photons/m2/s)
SolFitsPriscuii <- conversion(SolFitsPriscuii) 
SolFitsICEMDV <- conversion(SolFitsICEMDV) 
SolFitsMalina <- conversion(SolFitsMalina) 
SolFitsReinhardtii <- conversion(SolFitsReinhardtii) 
```


### Test Plots

```{r}
SolFitsICEMDV %>%
  ggplot()+
  geom_line(aes(x=Flashnumber, y=FvFmrunnorm, colour = as.factor(Filename))) +
  facet_grid(rows = vars(PulseSpace_s), cols = vars(TempDiff_C)) +
  theme_bw() +
  theme(legend.position = "none")

SolFitsMalina %>%
  ggplot()+
  geom_line(aes(x=Flashnumber, y=FvFmrunnorm, colour = as.factor(Filename))) +
  facet_grid(rows = vars(PulseSpace_s), cols = vars(TempDiff_C)) +
  theme_bw() +
  theme(legend.position = "none")

SolFitsPriscuii %>%
  ggplot()+
  geom_line(aes(x=Flashnumber, y=FvFmrunnorm, colour = as.factor(Filename))) +
  facet_grid(rows = vars(PulseSpace_s), cols = vars(TempDiff_C)) +
  theme_bw() +
  theme(legend.position = "none")

SolFitsReinhardtii %>%
  ggplot()+
  geom_line(aes(x=Flashnumber, y=FvFmrunnorm, colour = as.factor(Filename))) +
  facet_grid(rows = vars(PulseSpace_s), cols = vars(TempDiff_C)) +
  theme_bw() +
  theme(legend.position = "none")
```


```{r save data}
saveRDS(SolFitsICEMDV, file.path(DataOut, paste(Project, "Chlamydomonas", "ICEMDV", "SolFitsTrim.Rds", sep = "_"), fsep = .Platform$file.sep))

saveRDS(SolFitsMalina, file.path(DataOut, paste(Project, "Chlamydomonas", "malina", "SolFitsTrim.Rds", sep = "_"), fsep = .Platform$file.sep))

saveRDS(SolFitsPriscuii, file.path(DataOut, paste(Project, "Chlamydomonas", "priscuii", "SolFitsTrim.Rds", sep = "_"), fsep = .Platform$file.sep))

saveRDS(SolFitsReinhardtii, file.path(DataOut, paste(Project, "Chlamydomonas", "reinhardtii", "SolFitsTrim.Rds", sep = "_"), fsep = .Platform$file.sep))
```

# RLC Data
### Import
```{r list RLC files}
# RLCFiles <- list.files(path = DataRLC, pattern = FileID, full.names = TRUE)
# RLCFiles
# 
# unique(duplicated(RLCFiles))
```

```{r create RLC df}
# RLCFits <- RLCFiles %>%
#   map_df(~read.delim_plus(flnm =., file_encode = FileEncode, delimiter = Delimiter, header_rows = HeaderRows))
```

### Tidy
```{r tidy RLC fits}
# #for NoDecayDataset
# RLCFitsTrim <- RLCFits |>
#   filter(!grepl("----", DATE)) |> # remove rows with "----" 
#   select(-c("RFID_User_Data", "Barcode_Data","PIF", "LEDSel", "Dur.Wat", "Lon", "Lat", "GPS_stat","X"))  # remove superfluous columns
# 
# RLCFitsTrim <- RLCFitsTrim |>
#   mutate(Filename = str_remove(string = Filename, pattern = "_RLC_fit.csv")) |>
#     mutate(Filename = str_remove(string = Filename, pattern = "../Data/RawData/")) |>
#   separate(Filename, into=c("Location", "Type", "YYYYMMDD", "CultureID", "Temp_C"), sep = "([\\/\\_])", remove = FALSE) %>%
#   mutate(YYYYMMDD = lubridate::ymd(YYYYMMDD),
#           TIME = as.character(TIME)) %>%
#    rename(ObsDate = DATE,
#          ObsTime = TIME,
#          FvFm = "Fv.Fm") |>
#   mutate(Temp_C = str_remove(Temp_C, "C")) |>
#   type_convert() |>
#   mutate(ObsDateTime = lubridate::ymd_hms(paste(ObsDate, ObsTime))) %>%
#   relocate(ObsDateTime, .after = ObsTime) %>%
#   relocate(CultureID, .before = ObsDate) |>
#   type_convert()
# 
# 
# RLCFitsTrim <- RLCFitsTrim |>
#   group_by(Filename) |>
#   arrange(ObsTime) |>
#   mutate(Etime_s = (as.numeric(Time__mSec) - min(as.numeric(Time__mSec), na.rm = TRUE))/1000) %>%
#   ungroup()
# 
# RLCFitsTrim <- RLCFitsTrim |>
#   rowwise() |>
#   mutate(PAR_ue = sum(Light_1, Light_2, Light_3, Light_4, Light_5, Light_6)) |>
#   group_by(Filename) |>
#   mutate(Dark1s = case_when(row_number() == 1 ~ 0,
#                             row_number() %% 2 == 0 & row_number() > 1 ~ 0,
#                             row_number() %% 2 == 1 & row_number() > 1 ~ 1)) |>
#   ungroup()
```


