---
title: "OttawaImport"
author:
- Natasha Ryan
- Douglas A. Campbell
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
    code_folding: hide
    keep_md: yes
    fig_caption: yes
    toc: TRUE
    toc_float: TRUE   
csl: plos-one.csl
editor_options: 
  chunk_output_type: inline
---

*Some of the code used to create this R Notebook was refurbished from "PlateAbImport.Rmd" written by Maximilian Berthold, Douglas A. Campbell, Melissa L. Rioux, Sarah J Gore, and Alyson MacCormack.*

```{r load libraries}
library(tidyverse)
library(lubridate)
library(googlesheets4)
library(googledrive)
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_chunk$set(fig.path='Output/')
```

#Setup

### Set Project Variables 
```{r set project variables}
Project <- "LOW"
TargetData <- "Ottawa_2025"
DataOut <- file.path("..","Data", "CleanData")
CalibData <- file.path("..","Data", "CalibData")


DataIn <- file.path("..","Data", "RawData", TargetData, fsep = .Platform$file.sep)
#DataRefDecay <- file.path("..","Data", "RawData",TargetData, "DecayRef", fsep = .Platform$file.sep)
DataRLC <- file.path("..","Data", "RawData",TargetData, "RLC", fsep = .Platform$file.sep)

FileID <- "fit"

FileEncode <- "UTF-8" #guess_encoding says 'ASCII"

Delimiter <- ","

HeaderRows <- 0

#conversions
# us_s = 1000000
photons_umol = 6.022E17
A2_m2 = 1E20

#set colours
TempColours = c("black","violet",  "blue", "darkgreen", "green", "orange", "red", "darkred")

names(TempColours) <- c(2, 4, 8, 12, 16, 20, 24, 28)

TempColours
```


### Metadata Import
```{r import metadata}
googlesheets4::gs4_deauth()

OttawaMetadata <- read_sheet("https://docs.google.com/spreadsheets/d/1m8Pjwj4DL10AwdsQo2G3boymSWG1Oq-gJDzjU0csRKs/edit?usp=sharing")

OttawaMetadata <- OttawaMetadata |>
  filter(!is.na(PrimaryOperator))|>
  select(-c(InnocDate, Notes, `...19`))
```


### Create Functions
```{r warning=FALSE, message=FALSE, echo=FALSE}
read.delim_plus <- function(flnm, file_encode, delimiter, header_rows){read.delim(flnm, fileEncoding = file_encode, sep = delimiter,  skip = header_rows, row.names = NULL) %>% mutate(Filename = flnm)
}

conversion <- function(data){
  data %>%
    group_by(Filename, PulseSpace_s, Temp_C) %>%
    mutate(LightLevel_ue = (1/PulseSpace_s) / (mean(Sig) / A2_m2) / photons_umol) %>%
    mutate(LightLevel_ue = round(LightLevel_ue, 5)) %>%
    ungroup()
}
```


# No Decay Data

### Import
```{r list PSI files for file import}
SolisenseFiles <- list.files(path = DataIn, pattern = FileID, full.names = TRUE, recursive = FALSE)

SolisenseFiles 

unique(duplicated(SolisenseFiles))

SolisenseFilesExtra <- list.files(path = "../Data/RawData/Extrapriscui16C_2025", pattern = FileID, full.names = TRUE, recursive = FALSE)

SolisenseFilesExtra
```

```{r read Solisense files}
SolFits <- SolisenseFiles %>%
  map_df(~read.delim_plus(flnm =., file_encode = FileEncode, delimiter = Delimiter, header_rows = HeaderRows))
```

```{r read Solisense files extra}
SolFitsExtra <- SolisenseFilesExtra %>%
  map_df(~read.delim_plus(flnm =., file_encode = FileEncode, delimiter = Delimiter, header_rows = HeaderRows))
```

### Tidy 
```{r tidy dataset}
SolFitsTrim <- SolFits %>%
  filter(!grepl("----", DATE)) |>
  select(-c("RFID_User_Data", "Barcode_Data","PIF", "Lon", "Lat", "GPS_stat", "LEDSel",  "Light_1", "Light_2", "Light_3","Light_4","Light_5","Light_6", "Dur.Wat", "X", "StDev..":"Flag.6" )) |>
  mutate(Filename = str_remove(string = Filename, pattern = "_fit.csv")) |>
  mutate(Filename = str_remove(string = Filename, pattern = "../Data/RawData/Ottawa_2025/")) |>
  separate(Filename, into=c("YYYYMMDD", "CultureID", "Temp_C", "PulseSpace_s"), sep = "([\\/\\_])", remove = FALSE) |>
  mutate(YYYYMMDD = lubridate::ymd(YYYYMMDD),
          TIME = as.character(TIME)) |>
   rename(ObsDate = DATE,
         ObsTime = TIME,
         FvFm = "Fv.Fm") |>
  mutate(Temp_C = str_remove(Temp_C, "C")) %>%
  type_convert() %>%
  mutate(PulseSpace_s = str_remove(PulseSpace_s, "s")) %>%
  mutate(PulseSpace_s = as.numeric(PulseSpace_s)) %>%
  mutate(ObsDateTime = lubridate::ymd_hms(paste(ObsDate, ObsTime))) %>%
  relocate(ObsDateTime, .after = ObsTime) %>%
  relocate(CultureID, .before = ObsDate) %>%
  type_convert()

SolFitsTrim <-SolFitsTrim %>%
  group_by(Filename) %>%
  mutate(Flashnumber = row_number()) %>%
  relocate(Flashnumber, .before=p) %>%
  mutate(Etime_s = (as.numeric(Time__mSec) - min(as.numeric(Time__mSec), na.rm = TRUE))/1000,
         .before=Flashnumber) %>% #add col for time elapsed 
  mutate(TimeDiff_s = c(NA, diff(Time__mSec)/1000)) %>%
  ungroup()
```

```{r tidy extra dataset}
SolFitsTrimExtra <- SolFitsExtra %>%
  filter(!grepl("----", DATE)) |>
  select(-c("RFID_User_Data", "Barcode_Data","PIF", "Lon", "Lat", "GPS_stat", "LEDSel",  "Light_1", "Light_2", "Light_3","Light_4","Light_5","Light_6", "Dur.Wat", "X")) |>
  mutate(Filename = str_remove(string = Filename, pattern = "_fit.csv")) |>
  mutate(Filename = str_remove(string = Filename, pattern = "../Data/RawData/Extrapriscui16C_2025/")) |>
  separate(Filename, into=c("YYYYMMDD", "CultureID", "Temp_C", "PulseSpace_s"), sep = "([\\/\\_])", remove = FALSE) |>
  mutate(YYYYMMDD = lubridate::ymd(YYYYMMDD),
          TIME = as.character(TIME)) |>
   rename(ObsDate = DATE,
         ObsTime = TIME,
         FvFm = "Fv.Fm") |>
  mutate(Temp_C = str_remove(Temp_C, "C")) %>%
  type_convert() %>%
  mutate(PulseSpace_s = str_remove(PulseSpace_s, "s")) %>%
  mutate(PulseSpace_s = as.numeric(PulseSpace_s)) %>%
  mutate(ObsDateTime = lubridate::ymd_hms(paste(ObsDate, ObsTime))) %>%
  relocate(ObsDateTime, .after = ObsTime) %>%
  relocate(CultureID, .before = ObsDate) %>%
  type_convert()

SolFitsTrimExtra <-SolFitsTrimExtra %>%
  group_by(Filename) %>%
  mutate(Flashnumber = row_number()) %>%
  relocate(Flashnumber, .before=p) %>%
  mutate(Etime_s = (as.numeric(Time__mSec) - min(as.numeric(Time__mSec), na.rm = TRUE))/1000,
         .before=Flashnumber) %>% #add col for time elapsed 
  mutate(TimeDiff_s = c(NA, diff(Time__mSec)/1000)) %>%
  ungroup()
```


```{r test flash number}
# SolFitsTrimTest <- SolFitsTrim |>
#   filter(Flashnumber > 32)

#Temporary Flashnumber Fix

SolFitsTrim2 <- SolFitsTrim |>
  filter(Flashnumber <= 20)

SolFitsTrimExtra2 <- SolFitsTrimExtra |>
  filter(Flashnumber <= 20)

```


### Normalize Fluorescence & Add Temperature Differential Column 

```{r}

NormStartFlash = 12

SolFitsTrim2 <- SolFitsTrim2 %>%
  mutate(SState = case_when(Flashnumber %% 4 == 1 ~ 1,
                    Flashnumber %% 4 == 2 ~ 2,
                    Flashnumber %% 4 == 3 ~ 3, 
                    Flashnumber %% 4 == 0 ~ 4)) #add column for sstate

SolFitsTrimExtra2 <- SolFitsTrimExtra2 %>%
  mutate(SState = case_when(Flashnumber %% 4 == 1 ~ 1,
                    Flashnumber %% 4 == 2 ~ 2,
                    Flashnumber %% 4 == 3 ~ 3, 
                    Flashnumber %% 4 == 0 ~ 4))
#normalize values to final stretch of 10 values from 'randomized' PSII pool, done after cleaning to ensure the last 10 values are correct 
# SolFitsTrim2 <- SolFitsTrim2 %>%
#   filter(Filename != "20251121_MaPo2015_28C_4s") |>
#   group_by(Filename) %>%
#   mutate(Sigrunnorm =  Sig/mean(Sig[(length(Sig) - 10):length(Sig)], na.rm = TRUE),
#          Forunnorm = Fo/mean(Fo[(length(Fo) - 10):length(Fo)], na.rm = TRUE),
#          Fmrunnorm = Fm/mean(Fm[(length(Fm) - 10):length(Fm)], na.rm = TRUE),
#          FvFmrunnorm = FvFm/mean(FvFm[(length(FvFm) - 10):length(FvFm)], na.rm = TRUE)) %>%
#   ungroup()

#Ottawa_2025 did not run all captures for 32 flashes, so using last 10 points average may be problematic.  Normalize to flashes from 9 to end of flash train.

SolFitsTrim2 <- SolFitsTrim2 %>%
  filter(Filename != "20251121_MaPo2015_28C_4s") |>
  group_by(Filename) %>%
  mutate(FilenameLength = length(Sig),
         Sigrunnorm =  Sig/mean(Sig[NormStartFlash]:Sig[length(Sig)], na.rm = TRUE),
         Forunnorm = Fo/mean(Fo[NormStartFlash]:Fo[length(Fo)], na.rm = TRUE),
         Fmrunnorm = Fm/mean(Fm[NormStartFlash]:Fm[length(Fm)], na.rm = TRUE),
         FvFmrunnorm = FvFm/mean(FvFm[NormStartFlash]:FvFm[length(FvFm)], na.rm = TRUE)) %>%
  ungroup()

SolFitsTrimExtra2 <- SolFitsTrimExtra2 %>%
  filter(Filename != "20251121_MaPo2015_28C_4s") |>
  group_by(Filename) %>%
  mutate(FilenameLength = length(Sig),
         Sigrunnorm =  Sig/mean(Sig[NormStartFlash]:Sig[length(Sig)], na.rm = TRUE),
         Forunnorm = Fo/mean(Fo[NormStartFlash]:Fo[length(Fo)], na.rm = TRUE),
         Fmrunnorm = Fm/mean(Fm[NormStartFlash]:Fm[length(Fm)], na.rm = TRUE),
         FvFmrunnorm = FvFm/mean(FvFm[NormStartFlash]:FvFm[length(FvFm)], na.rm = TRUE)) %>%
  ungroup()

```


```{r, metadata merge}
SolFitsMeta <- left_join(SolFitsTrim2, OttawaMetadata, by = join_by(CultureID == ID))

SolFitsMetaExtra <- left_join(SolFitsTrimExtra2, OttawaMetadata, by = join_by(CultureID == ID))
```



Add columns for growth temperature and difference between measurement & growth temps
```{r TempDiff_C}

SolFitsMeta <- SolFitsMeta %>%
  mutate(TempDiff_C = Temp_C - GrowthTemp_C)

SolFitsMetaExtra <- SolFitsMetaExtra %>%
  mutate(TempDiff_C = Temp_C - GrowthTemp_C)

```


### Calculate equivalent Light Levels using conversion function
```{r}
# add column for light level (µmol photons/m2/s)
SolFitsMeta <- conversion(SolFitsMeta) 
SolFitsMetaExtra <- conversion(SolFitsMetaExtra) 

```


```{r test plots}

SolFitsMeta |>
  filter(Strain == "Chlamydomonas_priscui") |>
  ggplot() +
  geom_point(aes(x = Flashnumber, y = FvFmrunnorm, colour = as.factor(Temp_C)), size = 0.5) +
  scale_colour_manual(values = TempColours) +
  facet_grid(rows = vars(GrowthTemp_C,CultureID), cols = vars(Strain, PulseSpace_s)) +
  theme_bw()

SolFitsMeta |>
  filter(Strain == "Chlamydomonas_klinobasis") |>
  ggplot() +
  geom_point(aes(x = Flashnumber, y = FvFmrunnorm, colour = as.factor(Temp_C)), size = 0.1) +
  scale_colour_manual(values = TempColours) +
  facet_grid(rows = vars(GrowthTemp_C,CultureID), cols = vars(Strain, PulseSpace_s)) +
  theme_bw()

SolFitsMeta |>
  filter(Strain == "Chlamydomonas_malina") |>
  ggplot() +
  geom_point(aes(x = Flashnumber, y = FvFmrunnorm, colour = as.factor(Temp_C)), size = 0.1) +
  scale_colour_manual(values = TempColours) +
  facet_grid(rows = vars(GrowthTemp_C,CultureID), cols = vars(Strain, PulseSpace_s)) +
  theme_bw()

SolFitsMeta |>
  filter(Strain == "Chlamydomonas_euryale") |>
  ggplot() +
  geom_point(aes(x = Flashnumber, y = FvFmrunnorm, colour = as.factor(Temp_C)), size = 0.1) +
  scale_colour_manual(values = TempColours) +
  facet_grid(rows = vars(GrowthTemp_C,CultureID), cols = vars(Strain, PulseSpace_s)) +
  theme_bw()

SolFitsMeta |>
  filter(Strain == "Chlamydomonas_reinhardtii") |>
  ggplot() +
  geom_point(aes(x = Flashnumber, y = FvFmrunnorm, colour = as.factor(Temp_C)), size = 0.1) +
  scale_colour_manual(values = TempColours) +
  facet_grid(rows = vars(GrowthTemp_C,CultureID), cols = vars(Strain, PulseSpace_s)) +
  theme_bw()

SolFitsMetaExtra |>
  ggplot() +
  geom_point(aes(x = Flashnumber, y = FvFmrunnorm, colour = as.factor(Temp_C)), size = 0.1) +
  scale_colour_manual(values = TempColours) +
  facet_grid(rows = vars(GrowthTemp_C,CultureID), cols = vars(Strain, PulseSpace_s)) +
  theme_bw()
```
#Inspection of plots for Problem Runs
Filtered to 24 flashes

Cp

Possibly MaPo2004_16C_1s, 2 odd points  21 & 23 
Ck
nothing obvious

Cm
Possibly MaPo2011_4C_16s, points 21-24?

Ce
MaPo2015_28C_1s; last point high 24

Cr:
MaPo2002_28C_1 s, another run appended? points 19-24

SolFitsMetaExtra
All fine

Filtered to 20 flashes
Cr:
MaPo2002_28C_1 s, another run appended? points 19-24


```{r problemplot_filter}
SolFitsMeta <- SolFitsMeta |>
  filter(!grepl("MaPo2002_28C_1 s", Filename)) #remove for now
  
```

```{r merge extra plots}
SolFitsMeta <- rbind(SolFitsMeta, SolFitsMetaExtra)
```


```{r test plots single flash space}

PulseSpaceSet = 2

SolFitsMeta |>
  filter(Strain == "Chlamydomonas_priscui") |>
  filter(PulseSpace_s == PulseSpaceSet) |>
  ggplot() +
  geom_line(aes(x = Flashnumber, y = FvFmrunnorm, colour = as.factor(NaCl_mM))) +
  facet_grid(rows = vars(TempDiff_C, Temp_C), cols = vars(Strain, PulseSpace_s, NaCl_mM)) +
  theme_bw()

SolFitsMeta |>
  filter(Strain == "Chlamydomonas_klinobasis") |>
  filter(PulseSpace_s == PulseSpaceSet) |>
  ggplot() +
  geom_line(aes(x = Flashnumber, y = FvFmrunnorm, colour = as.factor(NaCl_mM))) +
  facet_grid(rows = vars(TempDiff_C, Temp_C), cols = vars(Strain, PulseSpace_s, NaCl_mM)) +
  theme_bw()

SolFitsMeta |>
  filter(Strain == "Chlamydomonas_malina") |>
  filter(PulseSpace_s == PulseSpaceSet) |>
  ggplot() +
  geom_line(aes(x = Flashnumber, y = FvFmrunnorm, colour = as.factor(NaCl_mM))) +
  facet_grid(rows = vars(TempDiff_C, Temp_C), cols = vars(Strain, PulseSpace_s, NaCl_mM)) +
  theme_bw()

SolFitsMeta |>
  filter(Strain == "Chlamydomonas_euryale") |>
  filter(PulseSpace_s == PulseSpaceSet) |>
  ggplot() +
  geom_line(aes(x = Flashnumber, y = FvFmrunnorm, colour = as.factor(NaCl_mM))) +
  facet_grid(rows = vars(TempDiff_C, Temp_C), cols = vars(Strain, PulseSpace_s,NaCl_mM)) +
  theme_bw()

SolFitsMeta |>
  filter(Strain == "Chlamydomonas_reinhardtii") |>
  filter(PulseSpace_s == PulseSpaceSet) |>
  ggplot() +
  geom_line(aes(x = Flashnumber, y = FvFmrunnorm, colour = as.factor(NaCl_mM))) +
  facet_grid(rows = vars(TempDiff_C, Temp_C), cols = vars(Strain, PulseSpace_s, NaCl_mM)) +
  theme_bw()
```

```{r save SolFitsMeta}
saveRDS(SolFitsMeta, file.path(DataOut, paste(Project, "Ottawa_2025", "SolFitsMeta.Rds", sep = "_"), fsep = .Platform$file.sep))
```



# RLC Data
### Import RLC; missing 8C & 12C files?
```{r list RLC files}
RLCFiles <- list.files(path = DataRLC, pattern = FileID, full.names = TRUE)
RLCFiles

unique(duplicated(RLCFiles))
```

```{r create RLC df}
RLCFits <- RLCFiles %>%
  map_df(~read.delim_plus(flnm =., file_encode = FileEncode, delimiter = Delimiter, header_rows = HeaderRows))
```

### Tidy RLC
```{r tidy RLC fits}

RLCFitsTrim <- RLCFits |>
  filter(!grepl("----", DATE)) |> # remove rows with "----"
  select(-c("RFID_User_Data", "Barcode_Data","PIF","Lon", "Lat", "GPS_stat", "LEDSel", "Dur.Wat", "StDev..", "StDev...25", "StDev...28", "StDev...28", "StDev...31", "StDev...34", "StDev...37", "StDev...40", "Flag", "Flag.1", "Flag.2","Flag.3", "Flag.4","Flag.5", "Flag.6", "X"))  # remove superfluous columns


RLCFitsTrim <- RLCFitsTrim |>
  mutate(Filename = str_remove(string = Filename, pattern = "_RLC_fit.csv")) |>
  mutate(Filename = str_remove(string = Filename, pattern = "../Data/RawData/"))|>
  mutate(Filename = str_remove(string = Filename, pattern = "Ottawa_2025/"))|>
  separate(Filename, into=c("Type", "YYYYMMDD", "CultureID", "Temp_C"), sep = "([\\/\\_])", remove = FALSE) |>
  mutate(YYYYMMDD = lubridate::ymd(YYYYMMDD),
         TIME = as.character(TIME)) |>
  rename(ObsDate = DATE,
         ObsTime = TIME,
         FvFm = "Fv.Fm") |>
  mutate(Temp_C = str_remove(Temp_C, "C")) |>
  type_convert() |>
  mutate(ObsDateTime = lubridate::ymd_hms(paste(ObsDate, ObsTime))) |>
  relocate(ObsDateTime, .after = ObsTime) |>
  relocate(CultureID, .before = ObsDate) |>
  type_convert()


RLCFitsTrim <- RLCFitsTrim |>
  group_by(Filename) |>
  arrange(ObsTime) |>
  mutate(Etime_s = (as.numeric(Time__mSec) - min(as.numeric(Time__mSec),
                                                 na.rm = TRUE))/1000) |>
  ungroup()


RLCFitsTrim <- RLCFitsTrim |>
  rowwise() |>
  mutate(PAR_ue = Light_1) |>
  group_by(Filename) |>
  mutate(Dark1s = case_when(row_number() == 1 ~ 0,
                            row_number() %% 2 == 0 & row_number() > 1 ~ 0,
                            row_number() %% 2 == 1 & row_number() > 1 ~ 1)) |>
  ungroup()
```

# RLC test plots
```{r RLCFitsTrim_testplots}
RLCFitsTrim |>
  filter(Dark1s == 0) |>
  ggplot() +
  geom_point(aes(x = PAR_ue, y = FvFm)) +
  facet_grid(rows = vars(Temp_C)) +
  theme_bw()
```

#Merge RLC with MetaData

```{r, metadata merge}
RLCFitsMeta <- left_join(RLCFitsTrim, OttawaMetadata, by = join_by(CultureID == ID))
```

# RLCFitsMeta test plots
```{r RLCFitsMeta_testplots}
RLCFitsMeta |>
  filter(Dark1s == 0) |>
  ggplot() +
  geom_point(aes(x = PAR_ue, y = FvFm)) +
  facet_grid(rows = vars(Temp_C), cols = vars(Strain)) +
  theme_bw()

RLCFitsMeta |>
  filter(Dark1s == 0) |>
  filter(Strain == "Chlamydomonas_priscui") |>
  ggplot() +
  geom_point(aes(x = PAR_ue, y = FvFm)) +
  facet_grid(rows = vars(Temp_C), cols = vars(Strain, NaCl_mM)) +
  theme_bw()

RLCFitsMeta |>
  filter(Dark1s == 0) |>
  filter(Strain == "Chlamydomonas_klinobasis") |>
  ggplot() +
  geom_point(aes(x = PAR_ue, y = FvFm)) +
  facet_grid(rows = vars(Temp_C), cols = vars(Strain, NaCl_mM)) +
  theme_bw()

RLCFitsMeta |>
  filter(Dark1s == 0) |>
  filter(Strain == "Chlamydomonas_malina") |>
  ggplot() +
  geom_point(aes(x = PAR_ue, y = FvFm)) +
  facet_grid(rows = vars(Temp_C), cols = vars(Strain, NaCl_mM)) +
  theme_bw()

RLCFitsMeta |>
  filter(Dark1s == 0) |>
  filter(Strain == "Chlamydomonas_reinhardtii") |>
  ggplot() +
  geom_point(aes(x = PAR_ue, y = FvFm)) +
  facet_grid(rows = vars(Temp_C), cols = vars(Strain, NaCl_mM)) +
  theme_bw()

RLCFitsMeta |>
  filter(Dark1s == 0) |>
  filter(Strain == "Chlamydomonas_euryale") |>
  ggplot() +
  geom_point(aes(x = PAR_ue, y = FvFm)) +
  facet_grid(rows = vars(Temp_C), cols = vars(Strain, NaCl_mM)) +
  theme_bw()
```


```{r save RLCFitsMeta}
saveRDS(RLCFitsMeta, file.path(DataOut, paste(Project, "Ottawa_2025", "RLCFitsMeta.Rds", sep = "_"), fsep = .Platform$file.sep))
```

