---
title: "OttawaImport"
author:
- Natasha Ryan
- Douglas A. Campbell
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
    code_folding: hide
    keep_md: yes
    fig_caption: yes
    toc: TRUE
    toc_float: TRUE   
csl: plos-one.csl
editor_options: 
  chunk_output_type: inline
---

*Some of the code used to create this R Notebook was refurbished from "PlateAbImport.Rmd" written by Maximilian Berthold, Douglas A. Campbell, Melissa L. Rioux, Sarah J Gore, and Alyson MacCormack.*

```{r load libraries}
library(tidyverse)
library(lubridate)
library(googlesheets4)
library(googledrive)
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_chunk$set(fig.path='Output/')
```

# Set Project Variables 
```{r set project variables}
Project <- "LOW"
TargetData <- "UOttawa"
DataOut <- file.path("..","Data", "CleanData")
CalibData <- file.path("..","Data", "CalibData")


DataIn <- file.path("..","Data", "RawData", TargetData, fsep = .Platform$file.sep)
DataRefDecay <- file.path("..","Data", "RawData",TargetData, "DecayRef", fsep = .Platform$file.sep)
DataRLC <- file.path("..","Data", "RawData",TargetData, "RLC", fsep = .Platform$file.sep)

FileID <- "fit"

FileEncode <- "UTF-8" #guess_encoding says 'ASCII"

Delimiter <- ","

HeaderRows <- 0
```

```{r conversions}
us_s = 1000000
photons_umol = 6.022E17
A2_m2 = 1E20
```

```{r read ActPAR calibration files}
#ActPARCal <- readRDS("~/Dropbox/CampbellLabProtocols/ChlorophyllFluorescence/SolisenseInformation/SolisenseInformation_DCCalibParam.Rds")

ActPARCrossCal <- list.files(path = CalibData, full.names = TRUE) %>%
     map_df(~readRDS(file  = .))

#intercept set to 0 in lm in SolisenseInformation.Rproj/SolisenseCalibCompare.Rmd
ActPARCrossCal <- ActPARCrossCal |>
  rename(#Intercept = `estimate_(Intercept)`,
         Slope = `estimate_LIFT_Gen_Developer.cal`,
         #Intercept_SE = `std.error_(Intercept)`,
         Slope_SE = `std.error_LIFT_Gen_Developer.cal`)
```

# Metadata Import
```{r import metadata}
googlesheets4::gs4_deauth()

OttawaMetadata <- read_sheet("https://docs.google.com/spreadsheets/d/1m8Pjwj4DL10AwdsQo2G3boymSWG1Oq-gJDzjU0csRKs/edit?usp=sharing")
```

# File Import
```{r list PSI files for file import}
SolisenseFiles <- list.files(path = DataIn, pattern = FileID, full.names = TRUE)
SolisenseFiles

RLCFiles <- list.files(path = DataRLC, pattern = FileID, full.names = TRUE)
RLCFiles

unique(duplicated(SolisenseFiles))
unique(duplicated(RLCFiles))
```


```{r warning=FALSE, message=FALSE, echo=FALSE}
read.delim_plus <- function(flnm, file_encode, delimiter, header_rows){read.delim(flnm, fileEncoding = file_encode, sep = delimiter,  skip = header_rows, row.names = NULL) %>% mutate(Filename = flnm)
}
```

# Create Data Frame 
purrr::map to read all files
```{r read Solisense files}
SolFits <- SolisenseFiles %>%
  map_df(~read.delim_plus(flnm =., file_encode = FileEncode, delimiter = Delimiter, header_rows = HeaderRows))

RLCFits <- RLCFiles %>%
  map_df(~read.delim_plus(flnm =., file_encode = FileEncode, delimiter = Delimiter, header_rows = HeaderRows))

```

# Tidy 
```{r tidy dataset}
SolFitsTrim <- SolFits %>%
  filter(!grepl("----", DATE)) %>%
  select(-c("RFID_User_Data", "Barcode_Data","PIF", "LEDSel", "Dur.Wat", "Lon", "Lat", "GPS_stat","X","Light_1", "Light_2", "Light_3","Light_4","Light_5","Light_6")) %>%
  mutate(Filename = str_remove(string = Filename, pattern = "_fit.csv")) %>%
  mutate(Filename = str_remove(string = Filename, pattern = "../Data/RawData/")) %>%
  separate(Filename, into=c("Location", "YYYYMMDD", "CultureID", "Temp_C", "PulseSpace_s"), sep = "([\\/\\_])", remove = FALSE) %>%
  mutate(YYYYMMDD = lubridate::ymd(YYYYMMDD),
          TIME = as.character(TIME)) %>%
   rename(ObsDate = DATE,
         ObsTime = TIME,
         FvFm = "Fv.Fm") %>%
  mutate(Temp_C = str_remove(Temp_C, "C")) %>%
  type_convert() %>%
  mutate(PulseSpace_s = str_remove(PulseSpace_s, "s")) %>%
  mutate(PulseSpace_s = str_remove(PulseSpace_s, "S")) %>%
  mutate(PulseSpace_s = as.numeric(PulseSpace_s)) %>%
  mutate(ObsDateTime = lubridate::ymd_hms(paste(ObsDate, ObsTime))) %>%
  relocate(ObsDateTime, .after = ObsTime) %>%
  relocate(CultureID, .before = ObsDate) %>%
  type_convert()

SolFitsTrim <-SolFitsTrim %>%
  group_by(Filename) %>%
  arrange(ObsTime) |>
  mutate(Flashnumber = row_number()) %>%
  mutate(Flashnumber2 = rank(ObsTime)) |>
  relocate(Flashnumber, .before=p) %>%
  mutate(Etime_s = (as.numeric(Time__mSec) - min(as.numeric(Time__mSec), na.rm = TRUE))/1000, .before=Flashnumber) %>%
  ungroup()
```


```{r tidy RLC fits}
#for NoDecayDataset
RLCFitsTrim <- RLCFits |>
  filter(!grepl("----", DATE)) |> # remove rows with "----" 
  select(-c("RFID_User_Data", "Barcode_Data","PIF", "LEDSel", "Dur.Wat", "Lon", "Lat", "GPS_stat","X"))  # remove superfluous columns

RLCFitsTrim <- RLCFitsTrim |>
  mutate(Filename = str_remove(string = Filename, pattern = "_RLC_fit.csv")) |>
    mutate(Filename = str_remove(string = Filename, pattern = "../Data/RawData/")) |>
  separate(Filename, into=c("Location", "Type", "YYYYMMDD", "CultureID", "Temp_C"), sep = "([\\/\\_])", remove = FALSE) %>%
  mutate(YYYYMMDD = lubridate::ymd(YYYYMMDD),
          TIME = as.character(TIME)) %>%
   rename(ObsDate = DATE,
         ObsTime = TIME,
         FvFm = "Fv.Fm") |>
  mutate(Temp_C = str_remove(Temp_C, "C")) |>
  type_convert() |>
  mutate(ObsDateTime = lubridate::ymd_hms(paste(ObsDate, ObsTime))) %>%
  relocate(ObsDateTime, .after = ObsTime) %>%
  relocate(CultureID, .before = ObsDate) |>
  type_convert()


RLCFitsTrim <- RLCFitsTrim |>
  group_by(Filename) |>
  arrange(ObsTime) |>
  mutate(Etime_s = (as.numeric(Time__mSec) - min(as.numeric(Time__mSec), na.rm = TRUE))/1000) %>%
  ungroup()

RLCFitsTrim <- RLCFitsTrim |>
  rowwise() |>
  mutate(PAR_ue = sum(Light_1, Light_2, Light_3, Light_4, Light_5, Light_6)) |>
  group_by(Filename) |>
  mutate(Dark1s = case_when(row_number() == 1 ~ 0,
                            row_number() %% 2 == 0 & row_number() > 1 ~ 0,
                            row_number() %% 2 == 1 & row_number() > 1 ~ 1)) |>
  ungroup()
```

# Chlamy File Issues

```{r}
#to check spacing to find correct runs 
# SolFitsTest <- SolFitsTrim %>%
#   filter(Filename == "Problem File") %>%
#   mutate(TimeDiff = c(NA, diff(Etime_s)))
```

20230919_MaPo1001_4C_16s (48 flashes, keep last 32)
20230918_MaPo1002_4C_16S (1 extra, keep last 32)
20230918_MaPo1002_4C_16s (1 extra, keep last 32)
20230919_MaPo1002_12C_8s (1 extra, keep last 32)
20230919_MaPo1004_4C_8s (1 extra, keep last 32)
20230920_MaPo1005_12C_16s (37 flashes, keep last 32)
20230918_MaPo1002_8C_16s (36 flashes keep last 32)
20230919_MaPo1004_8C_16s (4 extra, keep last 32)
20230918_MaPo1001_12C_16s (triple, contains 3 runs, 1-32 is 8s, 33-64 is 2s, and 65-96 is 16s)

20230920_MaPo1005_24C_16s (1 extra, keep last 31)
20230918_MaPo1001_4C_1s (double, keep first run)


```{r manually trim data}
SolFitsTrim_last32 <- SolFitsTrim %>%
  filter(!Filename %in% c('UOttawa/20230920_MaPo1005_24C_16s', 'UOttawa/20230918_MaPo1001_4C_1s')) %>%
  group_by(Filename) %>%
  slice(tail(row_number(),32)) %>% 
  mutate(Flashnumber2 = row_number()) %>%
  ungroup()

SolFitsTrim_MaPo1005_24C_16s <- SolFitsTrim %>%
  filter(Filename == 'UOttawa/20230920_MaPo1005_24C_16s') %>%
  slice(tail(row_number(),31)) %>% 
  mutate(Flashnumber2 = row_number()) %>%
  ungroup()

SolFitsTrim_MaPo1001_4C_1s <- SolFitsTrim %>%
  filter(Filename== "UOttawa/20230918_MaPo1001_4C_1s") %>%
  filter(Flashnumber <= 32)

SolFitsTrim2 <- rbind(SolFitsTrim_last32, SolFitsTrim_MaPo1001_4C_1s, SolFitsTrim_MaPo1005_24C_16s)
```

# Normalizing 
```{r}
# Moved this here so the fixed files would be normalized with the correct values 
SolFitsTrim2 <-SolFitsTrim2 %>%
  group_by(Filename) %>%
  arrange(ObsTime) |>
  mutate(Sigrunnorm =  Sig/mean(Sig[(length(Sig) - 10):length(Sig)], na.rm = TRUE),
         Forunnorm = Fo/mean(Fo[(length(Fo) - 10):length(Fo)], na.rm = TRUE),
         Fmrunnorm = Fm/mean(Fm[(length(Fm) - 10):length(Fm)], na.rm = TRUE),
         FvFmrunnorm = FvFm/mean(FvFm[(length(FvFm) - 10):length(FvFm)], na.rm = TRUE)) |>
  ungroup()

#add column for nominal SState for PSII that 'advance' with flash
SolFitsTrim2 <- SolFitsTrim2 |>
  mutate(SState = case_when(Flashnumber2 %% 4 == 1 ~ 1,
                    Flashnumber2 %% 4 == 2 ~ 2,
                    Flashnumber2 %% 4 == 3 ~ 3, 
                    Flashnumber2 %% 4 == 0 ~ 4))
```


```{r}
#filter bad runs 
SolFitsTrim2 <- SolFitsTrim2 %>%
  filter(!Filename %in% c("UOttawa/20230920_MaPo1006_24C_16s", "UOttawa/20230918_MaPo1001_12C_16s", "UOttawa/20230919_MaPo1001_8C_16s", "UOttawa/20230919_MaPo1002_4C_16s", "UOttawa/20230919_MaPo1002_8C_16s"))
```

```{r}
SolFitsTrim2  |>
  filter(CultureID == "MaPo1001") %>%
  filter(PulseSpace_s == 16) %>%
  ggplot()+
  geom_line(aes(x=Flashnumber2, y=FvFmrunnorm, colour = as.factor(PulseSpace_s))) +
  facet_wrap(vars(Filename)) +
  #facet_grid(rows= vars(Temp_C), cols = vars(PulseSpace_s)) +
  #facet_grid(rows = vars(PulseSpace_s, Temp_C), cols = vars(CultureID, ObsDate)) +
  theme_bw()
```



# Group by Taxon 

```{r}
SolFitsPriscuii <- SolFitsTrim2 %>%
  filter(CultureID %in% c("MaPo1001", "MaPo1004"))

SolFitsICEMDV <- SolFitsTrim2 %>%
  filter(CultureID %in% "MaPo1002")

SolFitsMalina <- SolFitsTrim2 %>%
  filter(CultureID %in% "MaPo1003")

SolFitsReinhardtii <- SolFitsTrim2 %>%
  filter(CultureID %in% c("MaPo1005", "MaPo1006"))
```

Overall: 

2 full runs at each temp/spacing combination except:
    - 20230919_MaPo1002_12C_16s only has 31 flashes
    - 20230918_MaPo1001_12C_8s only has 30 flashes     
    - 1 run of Chlamydomonas malina (MaPo1003 at 8C, 8s)
    - Chlamydomonas priscuii has 3 runs for each 
        - except 4C, 16s, which only has 2 
    
  

# Test Plots
```{r}
SolFitsMalina %>%
  filter(PulseSpace_s == 16) %>%
  ggplot() +
  geom_line(aes(x=Flashnumber2, y=FvFmrunnorm, colour = as.factor(CultureID))) +
  facet_wrap(vars(Filename)) +
  theme_bw()
```


```{r}
SolFitsICEMDV %>%
  ggplot() +
  geom_line(aes(x=Flashnumber2, y=FvFm)) +
  facet_grid(rows = vars(PulseSpace_s), cols = vars(Temp_C)) +
  labs(title = "Chlamydomonas_ICEMDV") +
  theme_bw()

SolFitsMalina %>%
  ggplot() +
  geom_line(aes(x=Flashnumber2, y=FvFm)) +
  facet_grid(rows = vars(PulseSpace_s), cols = vars(Temp_C)) +
  labs(title = "Chlamydomonas_malina") +
  theme_bw()

SolFitsReinhardtii %>%
  ggplot() +
  geom_line(aes(x=Flashnumber2, y=FvFm, colour = as.factor(CultureID))) +
  facet_grid(rows = vars(PulseSpace_s), cols = vars(Temp_C)) +
  labs(title = "Chlamydomonas_reinhardtii") +
  theme_bw()
```

```{r}
RLCFitsTrim %>%
  ggplot()+
  geom_line(aes(x = PAR_ue, y = FvFm)) +
  facet_grid(rows = vars(Temp_C), cols = vars(CultureID)) +
  labs(title = TargetData) +
  theme_bw()
```

```{r}
#facetbyspacing
SolFitsTrim2 %>%
  filter(CultureID == "MaPo1001") |>
  filter(ObsDate == "2023-09-18") |>
  filter(PulseSpace_s %in% c(1, 8)) |>
  filter(Temp_C %in% c(4, 12)) |>
  ggplot()+
  geom_line(aes(x=Flashnumber2, y= FvFm)) +
  facet_grid(rows = vars(Temp_C), cols = vars(PulseSpace_s)) +
  labs(title = paste("Antarctic Lake", "Chlamydomonas"), x = "Flash Number", y = "Maximum Quantum Yield of PSII") +
  theme_bw()

#ggsave(
 # "AntarcticChlamyFlashCyclesDecay.png",
 # plot = last_plot(),
 # path = file.path("..", "Output")
#)
```

```{r save data}
saveRDS(SolFitsICEMDV, file.path(DataOut, paste(Project, "Chlamydomonas", "ICEMDV", "SolFitsTrim.Rds", sep = "_"), fsep = .Platform$file.sep))

saveRDS(SolFitsMalina, file.path(DataOut, paste(Project, "Chlamydomonas", "malina", "SolFitsTrim.Rds", sep = "_"), fsep = .Platform$file.sep))

saveRDS(SolFitsPriscuii, file.path(DataOut, paste(Project, "Chlamydomonas", "priscuii", "SolFitsTrim.Rds", sep = "_"), fsep = .Platform$file.sep))

saveRDS(SolFitsReinhardtii, file.path(DataOut, paste(Project, "Chlamydomonas", "reinhardtii", "SolFitsTrim.Rds", sep = "_"), fsep = .Platform$file.sep))
```





