---
title: "Recombination Probabilities"
author:
- Douglas A. Campbell
- Natasha Ryan
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
    code_folding: hide
    keep_md: yes
    fig_caption: yes
    toc: TRUE
    toc_float: TRUE   
csl: plos-one.csl
editor_options: 
  chunk_output_type: inline
---



```{r load libraries}

library(tidyverse)
#library(readxl)
#library(stringr)
library(broom)
library(minpack.lm)
#library(OneR)
#library(changepoint)
#library(chngpt)
library(broom)
#library(Rwave)
#library(WaveletComp)

```

# Set Project Variables
```{r set project variables}
Project <- "LOW"
DataIn <- file.path("..","Data", "CleanData")
CalibData <- file.path("..","Data", "CalibData")

FileID <- "SolFitsTrim"

#Data with decay poor; go back to Solisense and re-fit to generate more files with decay data
#FileID <- "SolFitsDecayTrim"

```


```{r start variables}
PSIIo <- 1000 #PSII #possible number from Solisense
sigmaPSII_m2 <- 250e-20 #m2 photon-1 #actual number from Solisense
FlashletEx_uE <- 30000 #umol photons m-2 s-1 set by Solisense
Flashlet_s <- 1.2e-6 #duration of flashlet set by Solisense
FlashTrain <- 100 #number of sequential flashlets in the induction phase of flash set by Solisense
Dark_s <- 2e-6 #duration of dark period b/t sequential flashlets set by Solisense
Photons_umol <- 6.33E17 #conversion photons in a umol

SNL <- 0.005 #value for the signal to noise ratio; available from Solisense


Recomb <- 0.2 #fractional probability of average Recombination loss from PSII advancing through SState; could be a vector of 4 values, 1 for each SState

RecombFit <- 0 #goal fitted value for average Recombination probability; more sophisticated model might generate vector of 4 values

qp <- 0.9 #fraction of PSII closed during an excitation flashtrain; available from Solisense

ExcitPSIIPhotons = FlashletEx_uE * Photons_umol * sigmaPSII_m2 * FlashTrain * Flashlet_s

ExcitPSIIPhotons  #could be substrate for Poisson distribution modelling of fraction of 'missed' PSII that do not get excited during a flash

tau_s = 1e-3 #lifetime for re-opening of PSII after excitation flash train; available from Solisense

PSIIreopen = PSIIo - (PSIIo * exp(-1/tau_s  * FlashTrain * (Flashlet_s + Dark_s)))

PSIIreopen
```

```{r import clean data}

FitFilesSol <- list.files(path = DataIn, pattern = FileID, full.names = TRUE)

read.RDS <- function(flnm){readRDS(flnm) |>
    mutate(filename = flnm)
}

#test for duplicate file names
unique(duplicated(FitFilesSol))

LOW_SolFits <- FitFilesSol |>
  purrr::map_df(~read.RDS(flnm = .))

head(LOW_SolFits)

LOW_SolFits %>%
  ggplot() +
  geom_point(aes(x = Flashnumber, y = Fm)) +
  facet_grid(rows = vars(PulseSpace_s), cols = vars(Temp_C)) +
  theme_bw()
  
```

```{r nominal SState}
LOW_SolFits <- LOW_SolFits |>
  mutate(SState = case_when(Flashnumber %% 4 == 1 ~ 1,
                    Flashnumber %% 4 == 2 ~ 2,
                    Flashnumber %% 4 == 3 ~ 3, 
                    Flashnumber %% 4 == 0 ~ 4))

LOW_SolFits %>%
  ggplot() +
  geom_point(aes(x = Flashnumber, y = FvFmrunnorm, colour = as.factor(SState))) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  facet_grid(rows = vars(PulseSpace_s), cols = vars(Temp_C)) +
  theme_bw()

LOW_SolFits %>%
  ggplot() +
  geom_point(aes(x = Flashnumber, y = Sigrunnorm, colour = as.factor(SState))) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  facet_grid(rows = vars(PulseSpace_s), cols = vars(Temp_C)) +
  theme_bw()

LOW_SolFits %>%
  ggplot() +
  geom_point(aes(x = Flashnumber, y = Forunnorm, colour = as.factor(SState))) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  facet_grid(rows = vars(PulseSpace_s), cols = vars(Temp_C)) +
  theme_bw()

```

```{r exp decay function}
exp_eqn <- function(x, Intercept, mu){Intercept*exp(x*mu)
  }

expplateau_eqn <- function(x, Intercept, Plateau, mu){Plateau + (Intercept - Plateau)*exp(x*mu)
}

#Create single fitting function to make it easier to optimize?
expplateauSState_eqn <- function(sstate, x, Plateau, Intercept1, Intercept2, Intercept3, Intercept4,  mu1, mu2, mu3, mu4){
  case_when(sstate == 1 ~ Plateau + (Intercept1 - Plateau)*exp(x*mu1),
            sstate == 2 ~ Plateau + (Intercept2 - Plateau)*exp(x*mu2),
            sstate == 3 ~ Plateau + (Intercept3 - Plateau)*exp(x*mu3),
            sstate == 4 ~ Plateau + (Intercept4 - Plateau)*exp(x*mu4)
  )
  }
```



```{r FvFm decay to plateau models}

SolFitsModels <- LOW_SolFits |>
 # filter(SState == 1) |>
  nest(.by = c("Filename", "Temp_C", "PulseSpace_s", "SState")) |>
  mutate(FvFmSStates_model = map(data, possibly(~nlsLM(.$FvFmrunnorm ~ expplateau_eqn(x = .$Flashnumber, Intercept, Plateau = 1,  mu),
                                      data = .x,
                                      start = list(Intercept = 1.1, mu = 0.01),
                                      control = nls.lm.control(maxiter = 100)
                                      )
                                      ), NULL)
         ) |>
  mutate(FvFmSStates_Tidy = map(FvFmSStates_model, broom::tidy),
        FvFmSStates_Glance = map(FvFmSStates_model, broom::glance),
        FvFmSStates_Augment = map(FvFmSStates_model, broom::augment))

 #Not working

# SolFitsModels2 <- LOW_SolFits |>
#   nest(.by = c("Filename", "Temp_C","PulseSpace_s",  "SState")) |>
#   mutate(FvFmSStates4_model = map2(.x = SState,  .y = data, ~nlsLM(.y$FvFmrunnorm ~ expplateauSState_eqn(sstate = .x,  x = .y$Flashnumber, Plateau = 1, Intercept1, Intercept2, Intercept3, Intercept4,  mu1, mu2, mu3, mu4),
#                                       data = .y,
#                                       start = list(Intercept1 = 1, Intercept2 = 1, Intercept3 = 1, Intercept4 = 1, mu1 = 0.1, mu2 = 0.1, mu3 = 0.1, mu4 = 0.1),
#                                       control = nls.lm.control(maxiter = 100)
#                                       )
#                                       )
#   )
#                                    
                                   
                                   
  #                                  , NULL)
  #        ) |>
  # mutate(FvFmSStates_Tidy = map(FvFmSStates_model, broom::tidy),
  #       FvFmSStates_Glance = map(FvFmSStates_model, broom::glance),
  #       FvFmSStates_Augment = map(FvFmSStates_model, broom::augment))

 
# Adding legend for lines
#https://community.rstudio.com/t/adding-manual-legend-to-ggplot2/41651/2

# line_colours = c("ETRqpOxbo" = "green",
#             "fitted" = "black",
#             "residual" = "red")

SolFitsModels %>%
  unnest(FvFmSStates_Augment) |>
  ggplot() +
  geom_point(aes(x = Flashnumber, y = FvFmrunnorm, colour = as.factor(SState))) +
  geom_line(aes(x = Flashnumber, y = .fitted, colour = as.factor(SState))) +
  facet_grid(cols = vars(Temp_C), rows = vars(PulseSpace_s)) +
  theme_bw()

SolFitsModels %>%
  unnest(FvFmSStates_Tidy) |>
  filter(term == "mu") |>
  ggplot() +
  geom_point(aes(x = PulseSpace_s, y = estimate, colour = as.factor(SState))) +
  geom_line(aes(x = PulseSpace_s, y = estimate, colour = as.factor(SState))) +
  facet_grid(cols = vars(Temp_C)) +
  theme_bw()

SolFitsModels %>%
  unnest(FvFmSStates_Tidy) |>
  filter(term == "Intercept") |>
  ggplot() +
  geom_point(aes(x = PulseSpace_s, y = log10(estimate), colour = as.factor(SState))) +
  geom_line(aes(x = PulseSpace_s, y = log10(estimate), colour = as.factor(SState))) +
  facet_grid(cols = vars(Temp_C)) +
  theme_bw()

```


```{r filter for reasonable mu and intercept}
SolFitsModelsFilter <- SolFitsModels |>
  unnest(FvFmSStates_Tidy) |>
  filter(estimate > -2 & estimate < 2) |>
  select(-c("statistic")) |>
  pivot_wider(names_from = "term", values_from = c("estimate", "std.error", "p.value"))

SolFitsModelsFilter |>
  ggplot() +
  geom_point(aes(x = PulseSpace_s, y = estimate_mu, colour = as.factor(SState))) +
  geom_line(aes(x = PulseSpace_s, y = estimate_mu, colour = as.factor(SState))) +
  facet_grid(cols = vars(Temp_C)) +
  theme_bw()

SolFitsModelsFilter |>
  ggplot() +
  geom_point(aes(x = PulseSpace_s, y = estimate_Intercept, colour = as.factor(SState))) +
  geom_line(aes(x = PulseSpace_s, y = estimate_Intercept, colour = as.factor(SState))) +
  facet_grid(cols = vars(Temp_C)) +
  theme_bw()

# |>
#   group_by(Temp_C, SState) |>
#   mutate(SState_mu_av = mean(estimate_mu),
#          SState_intercept_av = mean(estimate_Intercept)) |>
#   ungroup()

```



Measure fraction of PSII closed after a flashtrain; that gives fraction of PSII that 'miss' advancing
Use Poisson distribution to estimate probability of double turnover during tau
Check on simulataneous rates arithmetic

```{r simulation}
Simulate <- tibble(Flashnumber = c(0:59),
                  SState = case_when(Flashnumber %% 4 == 1 ~ 1,
                    Flashnumber %% 4 == 2 ~ 2,
                    Flashnumber %% 4 == 3 ~ 3, 
                    Flashnumber %% 4 == 0 ~ 4),
                   PSIIadvance = PSIIo * qp^Flashnumber) |> #actually multiple issues encompassed by this estimate
  mutate(YSPSII = case_when(SState == 1 ~ 1.05,  #these values extractable from data
                           SState == 2 ~ 0.95,
                           SState == 3 ~ 0.95,
                           SState == 4 ~ 1.05)) |>
  mutate(meanYSPSII = mean(YSPSII)) |>
  mutate(FVFM = ((PSIIadvance * YSPSII) + ((PSIIo - PSIIadvance) * meanYSPSII))/PSIIo) |>
  mutate(YSPSIIupper = meanYSPSII + (meanYSPSII * SNL),
         YSPSIIlower = meanYSPSII - (meanYSPSII * SNL)) |>
  mutate(SignifFlash = if_else(FVFM <= YSPSIIupper & FVFM >= YSPSIIlower, 0, 1))

Simulate <- Simulate |>
   mutate(FlashThreshold =  as.numeric(filter(Simulate, SignifFlash == 1) |>
  slice_tail(n = 1) |>
   select(Flashnumber))
   )

                  #PSIIreopen = PSIIclose - (PSIIclose * exp(-1/tau_s  * FlashTrain * (Flashlet_s + Dark_s))) # have measured qP C from Solisense
                   #PSIInet = PSIIclose - PSIIreopen) |>


```


```{r simulate plot}
Simulate |>
  ggplot() +
  geom_point(aes(x = Flashnumber, y = PSIIadvance/PSIIo)) +
  geom_point(aes(x = Flashnumber, y = FVFM, colour = as.factor(SignifFlash))) +
  geom_line(aes(x = Flashnumber, y = YSPSIIupper), linetype = "dashed", colour = "green") +
  geom_line(aes(x = Flashnumber, y = YSPSIIlower), linetype = "dashed", colour = "green") +
  geom_vline(xintercept = Simulate$FlashThreshold[1], linetype = "dashed", colour = "darkgreen") +
  geom_text(aes(x = FlashThreshold[1] + 1, y = 0.6, label = paste("Significance Threshold", SNL, FlashThreshold[1], sep = " ")), colour = "darkgreen") +
  
  
  theme_bw()
```

Is there a flaw, b/c recombination 'reopens' PSII and will be caught by the qP closure parameter at the end of the flashtrain?
Is so will need to explicitly model PSII closure as:
(photons m-2 s-1) * (m2 photon) closure
1/tau reopening
Rate constants for 'known' processes; recombination then the third factor influencing reopening.
It is possible that recombination would show up as one of the three 'tau' for reopening.
Alternate extraction of amplitude of recombination from 'amplitudes' for fast, medium & slow kinetic phases of reopening.

```{r simulate with recombination}
Simulate <- Simulate |>
  #mutate(PSIIadvancerecomb = PSIIadvance - (PSIIadvance * Recomb), .after = PSIIreopen) |>
  mutate(PSIIadvancerecomb = PSIIadvance - (PSIIadvance * Recomb), .after = PSIIadvance) |>

  mutate(FMrecomb = ((PSIIadvancerecomb * YSPSII) + ((PSIIo - PSIIadvancerecomb) * meanYSPSII))/PSIIo) |>
  mutate(SignifFlashrecomb = if_else(FMrecomb <= YSPSIIupper & FMrecomb >= YSPSIIlower, 0, 1)) |>
  mutate(FlashrecombThreshold =  as.numeric(filter(Simulate, SignifFlashrecomb == 1) |>
  slice_tail(n = 1) |>
   select(Flashnumber))
  )

#remove noise factor for simplicity for now
# mutate(PSIIadvancerecombnoise = PSIIadvancerecomb * (1 + rnorm(n=10, sd = SNL))) |>

#Issue with PSIIo does not take into account re-closure of PSIIo upon subsequent flashes, but they will be out of sync

  Simulate |> 
  ggplot() +
    geom_point(aes(x = Flashnumber, y = PSIIadvancerecomb/PSIIo), colour = "red") +
    geom_point(aes(x = Flashnumber, y = FMrecomb, colour = as.factor(SignifFlashrecomb))) +
    geom_line(aes(x = Flashnumber, y = YSPSIIupper), linetype = "dashed", colour = "green") +
    geom_line(aes(x = Flashnumber, y = YSPSIIlower), linetype = "dashed", colour = "green") +
    geom_vline(xintercept = Simulate$FlashrecombThreshold[1], linetype = "dashed", colour = "red") +
    geom_text(aes(x = FlashrecombThreshold[1] + 1, y = 0.5, label = paste("Signif Thresh:", FlashrecombThreshold[1], "SNL:", SNL, "qP:", qp,  "Recomb:", Recomb,  sep = " "))) +
  geom_vline(xintercept = Simulate$FlashThreshold[1], linetype = "dashed", colour = "darkgreen") +
  geom_text(aes(x = FlashThreshold[1] + 1, y = 0.6,  label = paste("Signif Thresh:", FlashThreshold[1], "SNL:", SNL, "qP:", qp, sep = " ")), colour = "darkgreen") +
  theme_bw()
```
# Ask Zbigniew about estimation of PSII closure in Solisense software;
Is it more sophisticated than:
[PSII]o x sigmaPSII x I  -> [PSII]closed
[PSII]closed x 1/tau -> [PSII]open

If not, then estimate gap b/t 'simple' qp and 'actual' qp and attribute to Recombination?

# Set up optimization function to chose level of Recombination that matches data, with inputs of 'known' qP for expected SState cycling in the absence of Recombination
https://rstudio-pubs-static.s3.amazonaws.com/140054_deff0e911755406a87a2492d46f86d90.html

Consider fitting curves with eqns and then optimizing parameters of equation.


```{r find recomb}
RecombFit <- 0.5

Simulate <- Simulate |>
   mutate(FMfit = (((PSIIadvance - (PSIIadvance * RecombFit)) * YSPSII) + ((PSIIo - (PSIIadvance - (PSIIadvance * RecombFit))) * meanYSPSII))/PSIIo) |>
    mutate(FMgap2 = (FMfit - FMrecomb)^2,
         FMsum2 = sum(FMgap2))

#remove noise factor for simplicity
# * (1 + rnorm(n=10, sd = SNL))

 Simulate |> 
  ggplot() +
   geom_point(aes(x = Flashnumber, y = FMrecomb, colour = as.factor(SignifFlashrecomb))) +
   geom_line(aes(x = Flashnumber, y = YSPSIIupper), linetype = "dashed", colour = "green") +
   geom_line(aes(x = Flashnumber, y = YSPSIIlower), linetype = "dashed", colour = "green") +
   geom_point(aes(x = Flashnumber, y = FMfit)) +
   geom_point(aes(x = Flashnumber, y = FM)) +
  theme_bw()
 
```

fmgap <- function(FM, FMrecombnoise) {
        100 - 0.5 * price #a linear function
}
revenues <- function(price) {
        price * sales(price) #a non-linear function
}
par(mfrow = c(1, 2))
curve(sales, from = 0, to = 300, xname = 'Price', ylab = 'Sales', main = 'Sales')
curve(revenues, from = 0, to = 300, xname = 'Price', ylab = 'Revenues', main = 'Revenues')

par(mfrow = c(1, 1))
optimize(revenues, interval = c(50, 150), maximum = TRUE)

f <- function(x) {
  dfSum <- df[1,1]
  for (i in 1:nrow(df)) {
    dfSum = dfSum + df[i,1]*x - df[i,3]}
  dfSum
}

# solve (or optimize)...
s <- uniroot(f, lower=0, upper=1)

```{r minimize sum of squares}

RecombFit <- 0.01
Simulate <- Simulate |>
   mutate(FMfit = (((PSIIadvance - (PSIIadvance * RecombFit)) * YSPSII) + ((PSIIo - (PSIIadvance - (PSIIadvance * RecombFit))) * meanYSPSII))/PSIIo) |>
    mutate(FMgap2 = (FMfit - FMrecomb)^2,
         FMsum2 = sum(FMgap2))

OutputVector <- c()
RecombVec <- seq(from = 0.05, to = 0.95, by = 0.05)
for(i in c(1:19)) {
  FindRecomb = RecombVec[i]
  Simulate <- Simulate |>
   mutate(FMfit = (((PSIIadvance - (PSIIadvance * FindRecomb)) * YSPSII) + ((PSIIo - (PSIIadvance - (PSIIadvance * FindRecomb))) * meanYSPSII))/PSIIo) |>
    mutate(FMgap2 = (FMfit - FMrecomb)^2,
         FMsum2 = sum(FMgap2))
  OutputVector <- c(OutputVector,  Simulate$FMsum2[1])
  OptIndex = which(OutputVector == min(OutputVector, na.rm = TRUE))
  OptRecomb <- RecombVec[OptIndex]
}

Simulate2 <- Simulate |>
   mutate(FMfit = (((PSIIadvance - (PSIIadvance * OptRecomb)) * YSPSII) + ((PSIIo - (PSIIadvance - (PSIIadvance * OptRecomb))) * meanYSPSII))/PSIIo) |>
    mutate(FMgap2 = (FMfit - FMrecomb)^2,
         FMsum2 = sum(FMgap2))

 Simulate2 |> 
  ggplot() +
   geom_point(aes(x = Flashnumber, y = FMrecomb, colour = as.factor(SignifFlashrecomb))) +
   geom_line(aes(x = Flashnumber, y = YSPSIIupper), linetype = "dashed", colour = "green") +
   geom_line(aes(x = Flashnumber, y = YSPSIIlower), linetype = "dashed", colour = "green") +
   geom_point(aes(x = Flashnumber, y = FMfit)) +
   geom_point(aes(x = Flashnumber, y = FM)) +
  theme_bw()
```

