---
title: "Sackville Import"
author:
- Natasha Ryan
- Maximilian Berthold
- Douglas A. Campbell
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
    code_folding: hide
    keep_md: yes
    fig_caption: yes
    toc: TRUE
    toc_float: TRUE   
csl: plos-one.csl
editor_options: 
  chunk_output_type: inline
---

*Some of the code used to create this R Notebook was refurbished from "PlateAbImport.Rmd" written by Maximilian Berthold, Douglas A. Campbell, Melissa L. Rioux, Sarah J Gore, and Alyson MacCormack.*

```{r load libraries}
library(tidyverse)
library(lubridate)
library(googlesheets4)
library(googledrive)
```


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_chunk$set(fig.path='Output/')
```

# T Pseudonana 
### Setup
```{r set project variables}
Project <- "LOW"
TargetData <- "Sackville"
Target_WI23 <- "Pilot/Protocol2"
TargetTaxa <- "Tpseudonana"
DataOut <- file.path("..","Data", "CleanData")
CalibData <- file.path("..","Data", "CalibData")

DataIn <- file.path("..","Data", "RawData", TargetData, TargetTaxa, fsep = .Platform$file.sep)
DataRefDecay <- file.path("..","Data", "RawData",TargetTaxa, "DecayRef", fsep = .Platform$file.sep)
DataRLC <- file.path("..","Data", "RawData",TargetTaxa, "RLC", fsep = .Platform$file.sep)

FileID <- "fit"

FileEncode <- "UTF-8"
#guess_encoding says 'ASCII"

Delimiter <- ","

HeaderRows <- 0
```


```{r data read adds filename and cdate, warning=FALSE, message=FALSE, echo=FALSE}
read.delim_plus <- function(flnm, file_encode, delimiter, header_rows){read.delim(flnm, fileEncoding = file_encode, sep = delimiter,  skip = header_rows, row.names = NULL) %>% mutate(Filename = flnm)
}
```

### Metadata Import
```{r import metadata}
#SackvilleMetadata <- read_sheet("")
```

### No Decay
```{r list PSI files for file import}
SolisenseFiles <- list.files(path = DataIn, pattern = FileID, full.names = TRUE)
SolisenseFiles

unique(duplicated(SolisenseFiles)) #test for duplicate file names
```

```{r merging WI23 files}
#adding the files from WI23 since the FA24 Tpseudonana data collection was limites
DataIn_WI23 <- file.path("..","Data", "RawData", TargetData, Target_WI23, fsep = .Platform$file.sep)
SolisenseFiles_WI23 <- list.files(path = DataIn_WI23, pattern = FileID, full.names = TRUE)
SolisenseFiles_WI23

unique(duplicated(SolisenseFiles_WI23)) #test for duplicate file names
```

```{r read Solisense files}
SolFits <- SolisenseFiles %>%
  map_df(~read.delim_plus(flnm =., file_encode = FileEncode, delimiter = Delimiter, header_rows = HeaderRows))

SolFits_WI23 <- SolisenseFiles_WI23 %>%
  map_df(~read.delim_plus(flnm =., file_encode = FileEncode, delimiter = Delimiter, header_rows = HeaderRows))
```


```{r tidy SolFitsTrim ND}
#for NoDecayDataset
SolFitsTrim <- SolFits %>%
  filter(!grepl("----", DATE)) |> # remove rows with "----" 
  select(-c("RFID_User_Data", "Barcode_Data","PIF", "LEDSel", "Dur.Wat", "Lon", "Lat", "GPS_stat","X","Light_1", "Light_2", "Light_3","Light_4","Light_5","Light_6", "Tau1QA","Alp1QA","Alp1PQ","Tau1PQ")) %>% # remove superfluous columns
  mutate(Filename = str_remove(string = Filename, pattern = "_fit.csv")) %>%
  mutate(Filename = str_remove(string = Filename, pattern = "../Data/RawData/Sackville/Tpseudonana/")) %>% separate(Filename, into=c("YYYYMMDD", "CultureID", "Temp_C", "PulseSpace_s"), sep = "([\\/\\_])", remove = FALSE) %>%
  mutate(YYYYMMDD = lubridate::ymd(YYYYMMDD),
          TIME = as.character(TIME)) %>%
   rename(ObsDate = DATE,
         ObsTime = TIME,
         FvFm = "Fv.Fm") |>
  mutate(Temp_C = str_remove(Temp_C, "C")) |>
  type_convert() %>%
  mutate(PulseSpace_s = str_remove(PulseSpace_s, "S")) %>%  #s vs S
  mutate(PulseSpace_s = str_remove(PulseSpace_s, "s")) %>% 
  filter(PulseSpace_s %in% c(1, 2, 4, 8, 16)) |>  #remove rows with other PulseSpace_s
  mutate(PulseSpace_s = as.numeric(PulseSpace_s)) %>%
  mutate(ObsDateTime = lubridate::ymd_hms(paste(ObsDate, ObsTime))) %>%
  relocate(ObsDateTime, .after = ObsTime) %>%
  relocate(CultureID, .before = ObsDate) |>
  type_convert()

SolFitsTrim <-SolFitsTrim %>%
  group_by(Filename) %>%
  mutate(Flashnumber = row_number()) %>%
  relocate(Flashnumber, .before=p) %>%
  mutate(Etime_s = (as.numeric(Time__mSec) - min(as.numeric(Time__mSec), na.rm = TRUE))/1000, .before=Flashnumber) %>%
  ungroup()
```

```{r correct file issues}
ProblemFiles <- unique(SolFitsTrim$Filename[SolFitsTrim$Flashnumber >32])

ProblemList <- SolFitsTrim %>% 
  filter(Filename %in% c(ProblemFiles)) %>% 
  group_by(Filename) %>%
  mutate(TimeDiff = c(NA, diff(ObsTime))) %>% #shows actual spacing between flashes
  ungroup() %>%
  doBy::split_by("Filename") #creates list of problem files
```

From ProblemList, use spacing between flashes to decipher runs:
  20230911_NaOm1663_22C_2S        - 58 flashes, no good runs 
  20230911_NaOm1663_22C_4S        - 34 flashes, only first 29 are good
  
  20230912_NaOm1663_18C_ 8S, 4S, 2S, & 1S     - 96 flashes, saved exact same run 3x; odd spikes
  

```{r manual trim}
first32 <- SolFitsTrim %>%
  filter(!Filename %in% c("20230911_NaOm1663_22C_4S", '20230911_NaOm1663_22C_2S'))  %>%
  group_by(Filename) %>%
  filter(Flashnumber %in% c(1:32)) %>%
  mutate(Flashnumber = row_number()) %>%
  ungroup()

NaOm1663_22C_4S <- SolFitsTrim %>%
  filter(Filename == '20230911_NaOm1663_22C_4S') %>%
  filter(Flashnumber %in% c(1:29)) %>%
  mutate(Flashnumber = row_number())

SolFitsTrim2 <- rbind(first32, NaOm1663_22C_4S)
```


```{r tidy clean data}
SolFitsTrim2 <- SolFitsTrim2 %>%
  mutate(SState = case_when(Flashnumber %% 4 == 1 ~ 1,
                    Flashnumber %% 4 == 2 ~ 2,
                    Flashnumber %% 4 == 3 ~ 3, 
                    Flashnumber %% 4 == 0 ~ 4)) #add column for sstate

#normalize values to final stretch of 10 values from 'randomized' PSII pool, done after cleaning to ensure the last 10 values are correct 
SolFitsTrim2 <- SolFitsTrim2 %>%
  group_by(Filename) %>%
  mutate(Sigrunnorm =  Sig/mean(Sig[(length(Sig) - 10):length(Sig)], na.rm = TRUE),
         Forunnorm = Fo/mean(Fo[(length(Fo) - 10):length(Fo)], na.rm = TRUE),
         Fmrunnorm = Fm/mean(Fm[(length(Fm) - 10):length(Fm)], na.rm = TRUE),
         FvFmrunnorm = FvFm/mean(FvFm[(length(FvFm) - 10):length(FvFm)], na.rm = TRUE)) %>%
  ungroup()
```

### WI 23 Data
```{r tidy SolFitsTrim WI23}
SolFitsTrim_WI23 <- SolFits_WI23 %>%
  filter(!grepl("----", DATE)) |> # remove rows with "----" 
  select(-c("RFID_User_Data", "Barcode_Data","PIF", "LEDSel", "Dur.Wat", "Lon", "Lat", "GPS_stat","X","Light_1", "Light_2", "Light_3","Light_4","Light_5","Light_6")) %>% # remove superfluous columns
  mutate(Filename = str_remove(string = Filename, pattern = "_fit.csv")) %>%
  mutate(Filename = str_remove(string = Filename, pattern = "../Data/RawData/Sackville/Pilot/Protocol2/")) %>%
  separate(Filename, into=c("YYYYMMDD", "CultureID", "Temp_C", "PulseSpace_s"), sep = "([\\/\\_])", remove = FALSE) %>%
  mutate(YYYYMMDD = lubridate::ymd(YYYYMMDD),
          TIME = as.character(TIME)) %>%
   rename(ObsDate = DATE,
         ObsTime = TIME,
         FvFm = "Fv.Fm") |>
  mutate(Temp_C = str_remove(Temp_C, "C")) |>
  type_convert() %>%
  mutate(PulseSpace_s = str_remove(PulseSpace_s, "S")) |>  #s vs S
  filter(PulseSpace_s %in% c(1, 2, 4, 8)) |>  #remove rows with other PulseSpace_s
  mutate(PulseSpace_s = as.numeric(PulseSpace_s)) %>%
  mutate(ObsDateTime = lubridate::ymd_hms(paste(ObsDate, ObsTime))) %>%
  relocate(ObsDateTime, .after = ObsTime) %>%
  relocate(CultureID, .before = ObsDate) |>
  type_convert()


SolFitsTrim_WI23 <-SolFitsTrim_WI23 %>%
  group_by(Filename) %>%
  mutate(Flashnumber = row_number()) %>%
  relocate(Flashnumber, .before=p) %>%
  mutate(Etime_s = (as.numeric(Time__mSec) - min(as.numeric(Time__mSec), na.rm = TRUE))/1000, .before=Flashnumber) %>%
  ungroup()
```


```{r correct file issues}
ProblemFiles_WI23 <- unique(SolFitsTrim_WI23$Filename[SolFitsTrim_WI23$Flashnumber >32])
ProblemFiles_WI23 #no problem files so can continue
```

```{r tidy clean data WI23}
SolFitsTrim_WI23 <- SolFitsTrim_WI23 %>%
  mutate(SState = case_when(Flashnumber %% 4 == 1 ~ 1,
                    Flashnumber %% 4 == 2 ~ 2,
                    Flashnumber %% 4 == 3 ~ 3, 
                    Flashnumber %% 4 == 0 ~ 4)) #add column for sstate

#normalize values to final stretch of 10 values from 'randomized' PSII pool, done after cleaning to ensure the last 10 values are correct 
SolFitsTrim_WI23 <- SolFitsTrim_WI23 %>%
  group_by(Filename) %>%
  mutate(Sigrunnorm =  Sig/mean(Sig[(length(Sig) - 10):length(Sig)], na.rm = TRUE),
         Forunnorm = Fo/mean(Fo[(length(Fo) - 10):length(Fo)], na.rm = TRUE),
         Fmrunnorm = Fm/mean(Fm[(length(Fm) - 10):length(Fm)], na.rm = TRUE),
         FvFmrunnorm = FvFm/mean(FvFm[(length(FvFm) - 10):length(FvFm)], na.rm = TRUE)) %>%
  ungroup()
```

### Merge Files 
```{r}
SolFitsTrim_Combined <- rbind(SolFitsTrim2, SolFitsTrim_WI23)
```

```{r add deviation from growth temp}
SolFitsTrim_Combined <- SolFitsTrim_Combined %>%
  mutate(GrowthTemp_C = 18) %>%
  mutate(TempDiff_C = Temp_C - GrowthTemp_C) 
```

```{r}
#Remove trials with overinduction 
SolFitsTrim_Combined2 <- SolFitsTrim_Combined %>% 
  filter(!YYYYMMDD == "2024-03-11") 
```


```{r save SolFitsTrim data}
saveRDS(SolFitsTrim_Combined2, file.path(DataOut, paste(Project, "Thalassiosira_pseudonana", "SolFitsTrim.Rds", sep = "_"), fsep = .Platform$file.sep))
```



### Test Plots 

```{r Fm vs Flashnumber(by T)}
SolFitsTrim_Combined2 %>%
  ggplot()+
  geom_line(aes(x=Flashnumber, y=Fmrunnorm, colour = as.factor(CultureID))) +
  facet_grid(rows = vars(PulseSpace_s), cols = vars(TempDiff_C)) +
  theme_bw()
```


# Chlorella
### Setup
```{r set project variables}
TargetTaxa2 <- "Chlorella"
DataIn_Chlorella <- file.path("..","Data", "RawData", TargetData, TargetTaxa2, fsep = .Platform$file.sep)
```

### No Decay
```{r list PSI files for file import}
ChlorellaFiles <- list.files(path = DataIn_Chlorella, pattern = FileID, full.names = TRUE)
ChlorellaFiles

unique(duplicated(ChlorellaFiles)) #test for duplicate file names
```

```{r read Chlorella files}
ChlorellaFits <- ChlorellaFiles %>%
  map_df(~read.delim_plus(flnm =., file_encode = FileEncode, delimiter = Delimiter, header_rows = HeaderRows))
```


```{r tidy SolFitsTrim ND}
#for NoDecayDataset
ChlorellaFitsTrim <- ChlorellaFits %>%
  filter(!grepl("----", DATE)) |> # remove rows with "----" 
  select(-c("RFID_User_Data", "Barcode_Data","PIF", "LEDSel", "Dur.Wat", "Lon", "Lat", "GPS_stat","X","Light_1", "Light_2", "Light_3","Light_4","Light_5","Light_6")) %>% # remove superfluous columns
  mutate(Filename = str_remove(string = Filename, pattern = "_fit.csv")) %>%
  mutate(Filename = str_remove(string = Filename, pattern = "../Data/RawData/Sackville/Chlorella/")) %>% separate(Filename, into=c("YYYYMMDD", "CultureID", "Temp_C", "PulseSpace_s"), sep = "([\\/\\_])", remove = FALSE) %>%
  mutate(YYYYMMDD = lubridate::ymd(YYYYMMDD),
          TIME = as.character(TIME)) %>%
   rename(ObsDate = DATE,
         ObsTime = TIME,
         FvFm = "Fv.Fm") |>
  mutate(Temp_C = str_remove(Temp_C, "C")) |>
  type_convert() |>
  mutate(PulseSpace_s = str_remove(PulseSpace_s, "S")) |>  #s vs S
  filter(PulseSpace_s %in% c(1, 2, 4, 8, 16)) |>  #remove rows with other PulseSpace_s
  mutate(PulseSpace_s = as.numeric(PulseSpace_s)) %>%
  mutate(ObsDateTime = lubridate::ymd_hms(paste(ObsDate, ObsTime))) %>%
  relocate(ObsDateTime, .after = ObsTime) %>%
  relocate(CultureID, .before = ObsDate) |>
  type_convert()

ChlorellaFitsTrim <- ChlorellaFitsTrim %>%
  group_by(Filename) %>%
  mutate(Flashnumber = row_number()) %>%
  relocate(Flashnumber, .before=p) %>%
  mutate(Etime_s = (as.numeric(Time__mSec) - min(as.numeric(Time__mSec), na.rm = TRUE))/1000, .before=Flashnumber) %>%
  ungroup()
```

```{r correct chlorella file issues}
ChlorellaProblemFiles <- unique(ChlorellaFitsTrim$Filename[ChlorellaFitsTrim$Flashnumber >32])
ChlorellaProblemFiles #no issues so I can proceed
```

```{r tidy clean chlorella data}
ChlorellaFitsTrim <-ChlorellaFitsTrim %>%
  mutate(SState = case_when(Flashnumber %% 4 == 1 ~ 1,
                    Flashnumber %% 4 == 2 ~ 2,
                    Flashnumber %% 4 == 3 ~ 3, 
                    Flashnumber %% 4 == 0 ~ 4)) #add column for sstate

#normalize values to final stretch of 10 values from 'randomized' PSII pool, done after cleaning to ensure the last 10 values are correct 
ChlorellaFitsTrim <-ChlorellaFitsTrim %>%
  group_by(Filename) %>%
  mutate(Sigrunnorm =  Sig/mean(Sig[(length(Sig) - 10):length(Sig)], na.rm = TRUE),
         Forunnorm = Fo/mean(Fo[(length(Fo) - 10):length(Fo)], na.rm = TRUE),
         Fmrunnorm = Fm/mean(Fm[(length(Fm) - 10):length(Fm)], na.rm = TRUE),
         FvFmrunnorm = FvFm/mean(FvFm[(length(FvFm) - 10):length(FvFm)], na.rm = TRUE)) %>%
  ungroup()
```

```{r add deviation from growth temp}
ChlorellaFitsTrim <-ChlorellaFitsTrim %>%
  mutate(GrowthTemp_C = 18) %>%
  mutate(TempDiff_C = Temp_C - GrowthTemp_C) %>%
  filter(!Filename == "20240312_NaOm2987_26C_1S")
```


### Test Plots
```{r Fm vs Flashnumber(by T)}
ChlorellaFitsTrim %>%
  ggplot()+
  geom_line(aes(x=Flashnumber, y=FvFmrunnorm, colour = as.factor(Filename))) +
  facet_grid(rows = vars(PulseSpace_s), cols = vars(TempDiff_C)) +
  theme_bw() +
  theme(legend.position = "none")
```

```{r save Chlorella data}
saveRDS(ChlorellaFitsTrim, file.path(DataOut, paste(Project, "Chlorella_vulgaris", "SolFitsTrim.Rds", sep = "_"), fsep = .Platform$file.sep))
```


# Other Code 
```{r read ActPAR calibration files}
# #ActPARCal <- readRDS("~/Dropbox/CampbellLabProtocols/ChlorophyllFluorescence/SolisenseInformation/SolisenseInformation_DCCalibParam.Rds")
# 
# ActPARCrossCal <- list.files(path = CalibData, full.names = TRUE) %>%
#      map_df(~readRDS(file  = .))
# 
# #intercept set to 0 in lm in SolisenseInformation.Rproj/SolisenseCalibCompare.Rmd
# ActPARCrossCal <- ActPARCrossCal |>
#   rename(#Intercept = `estimate_(Intercept)`,
#          Slope = `estimate_LIFT_Gen_Developer.cal`,
#          #Intercept_SE = `std.error_(Intercept)`,
#          Slope_SE = `std.error_LIFT_Gen_Developer.cal`)
```

Odd issue in WithDecay files, with ChiSq appearing a second time after QARredMax column, offsetting column headers from proper data columns
```{r fix column names}
#only use this chunk if analyzing data from Data-RawData-Pilot-WithDecay-Fit
#create new data frame, had to remove extra column

#OldNames <- c("Chisq", "Fo", "Fm","Fv.Fm", "Sig", "p","Alp1QA","Tau1QA","Alp2QA","Tau2QA","Alp3QA")
#NewNames <- c("Fo", "Fm","Fv.Fm", "Sig", "p","Alp1QA","Tau1QA","Alp2QA","Tau2QA","Alp3QA","Tau3QANew") #rename_with(~ NewNames[which(OldNames == .x)], .cols = OldNames) |>

  #select(-c(Tau3QA)) |>
  #rename(Tau3QA = Tau3QANew) |>

#SolFits2 <- SolFits[-32] 
#assign correct names
#Fixnames <- colnames(SolFits)[22:32]
#rename columns 
#colnames(SolFits2)[21:31] <- c(Fixnames)
```

```{r tidy SolFitsTrim WD}
# for WithDecay Fit Dataset

#Think of better ways to do this
#Careful about RunDate, which is the label of a RunDate sub-folder.
#SolFitsTrim <- SolFits %>% 
  #filter(!grepl("----", DATE)) %>% # remove rows with "----"
 # select(-c("RFID_User_Data", "Barcode_Data","PIF", "Lon", "Lat", "GPS_stat","X")) %>% # remove superfluous columns
  #mutate(Filename = str_remove(string = Filename, pattern = "_fit.csv")) %>%
  #mutate(Filename = str_remove(string = Filename, pattern = "../Data/RawData/NoDecay/Fit/")) %>%
  #separate(Filename, into=c("YYYYMMDD", "CultureID", "Temp_C", "PulseSpace_s"), sep = "([\\/\\_])", remove = FALSE) %>%
  #separate(Filename, into = c("Project", "YYYYMMDD", "CultureID","Temp_C", "PulseSpace_s","Ex_WL", "Flashpower"), sep = "([\\/\\_])", remove = FALSE) %>%
 # mutate(YYYYMMDD = lubridate::ymd(YYYYMMDD),
         # TIME = as.character(TIME)) %>% #time-column may be read in as factor, and as.character changes it to numeric; using lubdridate::hms would only change the format to 13H 4M 2S but does not work later to merge into one DateTime-column
  #rename(ObsDate = DATE,
        # ObsTime = TIME,
       #  FvFm = "Fv.Fm") %>%
 # mutate(Ex_WL = as.factor(as.numeric(Ex_WL))) %>%
  #mutate(FvFm = as.numeric(as.character(FvFm)),
        # nm445 = as.numeric(as.character(Light_1)),
        # nm470 = as.numeric(as.character(Light_2)),
        # nm505 = as.numeric(as.character(Light_3)),
       # nm535 = as.numeric(as.character(Light_4)),
       #  nm590 = as.numeric(as.character(Light_5)),
        # IR = as.numeric(as.character(Light_6))) %>%
 # mutate(ObsDateTime = lubridate::ymd_hms(paste(ObsDate, ObsTime))) %>%
 # relocate(ObsDateTime, .after = ObsTime) %>%
 # relocate(CultureID, .before = ObsDate)

#SolFitsTrim <-SolFitsTrim %>%
#  group_by(Filename) %>%
 # mutate(Flashnumber = row_number()) %>%
 # relocate(Flashnumber, .before=p) %>%
  #mutate(Etime_s = (as.numeric(Time__mSec) - min(as.numeric(Time__mSec), na.rm = TRUE))/1000, .before=Flashnumber) %>%
  #ungroup()
```


```{r tidy RLC Fits}
# #for NoDecayDataset
# RLCFitsTrim <- RLCFits |>
#   filter(!grepl("----", DATE)) |> # remove rows with "----" 
#   select(-c("RFID_User_Data", "Barcode_Data","PIF", "LEDSel", "Dur.Wat", "Lon", "Lat", "GPS_stat","X"))  # remove superfluous columns
# 
# RLCFitsTrim <- RLCFitsTrim |>
#   mutate(Filename = str_remove(string = Filename, pattern = "_RLC_fit.csv")) |>
#     mutate(Filename = str_remove(string = Filename, pattern = "../Data/RawData/")) |>
#   separate(Filename, into=c("Location", "Type", "YYYYMMDD", "CultureID", "Temp_C"), sep = "([\\/\\_])", remove = FALSE) %>%
#   mutate(YYYYMMDD = lubridate::ymd(YYYYMMDD),
#           TIME = as.character(TIME)) %>%
#    rename(ObsDate = DATE,
#          ObsTime = TIME,
#          FvFm = "Fv.Fm") |>
#   mutate(Temp_C = str_remove(Temp_C, "C")) |>
#   type_convert() |>
#   mutate(ObsDateTime = lubridate::ymd_hms(paste(ObsDate, ObsTime))) %>%
#   relocate(ObsDateTime, .after = ObsTime) %>%
#   relocate(CultureID, .before = ObsDate) |>
#   type_convert()
# 
# 
# RLCFitsTrim <- RLCFitsTrim |>
#   group_by(Filename) |>
#   arrange(ObsTime) |>
#   mutate(Etime_s = (as.numeric(Time__mSec) - min(as.numeric(Time__mSec), na.rm = TRUE))/1000) %>%
#   ungroup()
# 
# RLCFitsTrim <- RLCFitsTrim |>
#   rowwise() |>
#   mutate(PAR_ue = sum(Light_1, Light_2, Light_3, Light_4, Light_5, Light_6)) |>
#   group_by(Filename) |>
#   mutate(Dark1s = case_when(row_number() == 1 ~ 0,
#                             row_number() %% 2 == 0 & row_number() > 1 ~ 0,
#                             row_number() %% 2 == 1 & row_number() > 1 ~ 1)) |>
#   ungroup()
#      
# 
# head(RLCFitsTrim)
```



