---
title: "Sackville Import"
author:
- Natasha Ryan
- Maximilian Berthold
- Douglas A. Campbell
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
    code_folding: hide
    keep_md: yes
    fig_caption: yes
    toc: TRUE
    toc_float: TRUE   
csl: plos-one.csl
editor_options: 
  chunk_output_type: inline
---

*Some of the code used to create this R Notebook was refurbished from "PlateAbImport.Rmd" written by Maximilian Berthold, Douglas A. Campbell, Melissa L. Rioux, Sarah J Gore, and Alyson MacCormack.*

```{r load libraries}
library(tidyverse)
library(lubridate)
library(googlesheets4)
library(googledrive)
```


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_chunk$set(fig.path='Output/')
```

#Setup

### Set Project Variables 
```{r set project variables}
Project <- "LOW"
TargetData <- "Sackville"
DataOut <- file.path("..","Data", "CleanData")
CalibData <- file.path("..","Data", "CalibData")


DataIn <- file.path("..","Data", "RawData", TargetData, fsep = .Platform$file.sep)
DataRefDecay <- file.path("..","Data", "RawData",TargetData, "DecayRef", fsep = .Platform$file.sep)
#DataRLC <- file.path("..","Data", "RawData",TargetData, "RLC", fsep = .Platform$file.sep)

FileID <- "fit"

FileEncode <- "UTF-8" #guess_encoding says 'ASCII"

Delimiter <- ","

HeaderRows <- 0

#conversions 
#us_s = 1000000
photons_umol = 6.022E17
A2_m2 = 1E20
```


### Metadata Import
```{r import metadata}
#googlesheets4::gs4_deauth()

#SackvilleMetadata <- read_sheet("https://docs.google.com/spreadsheets/d/13QTuZybpU-m63X-x8DeM3eVRDmFz8JIZTRnLWsMeIw0/edit?usp=sharing")
```


### Create Functions 
```{r warning=FALSE, message=FALSE, echo=FALSE}
read.delim_plus <- function(flnm, file_encode, delimiter, header_rows){read.delim(flnm, fileEncoding = file_encode, sep = delimiter,  skip = header_rows, row.names = NULL) %>% mutate(Filename = flnm)
}

conversion <- function(data){
  data %>%
    group_by(PulseSpace_s, Temp_C) %>%
    mutate(Light_Level = (1/PulseSpace_s) / (mean(Sig) / A2_m2) / photons_umol) %>%
    mutate(Light_Level = round(Light_Level, 5)) %>%
    ungroup()
}
```


# No Decay Data

### Import 
```{r list PSI files for file import}
SolisenseFiles <- list.files(path = DataIn, pattern = FileID, full.names = TRUE)
SolisenseFiles

unique(duplicated(SolisenseFiles)) #test for duplicate file names
```

```{r read Solisense files}
SolFits <- SolisenseFiles %>%
  map_df(~read.delim_plus(flnm =., file_encode = FileEncode, delimiter = Delimiter, header_rows = HeaderRows)) #purrr::map to read all files
```


```{r tidy SolFitsTrim ND}
#for NoDecayDataset
SolFitsTrim <- SolFits %>%
  filter(!grepl("----", DATE)) |> # remove rows with "----" 
  select(-c("RFID_User_Data", "Barcode_Data","PIF", "LEDSel", "Dur.Wat", "Lon", "Lat", "GPS_stat","X","Light_1", "Light_2", "Light_3","Light_4","Light_5","Light_6", "Tau1QA","Alp1QA","Alp1PQ","Tau1PQ")) %>% # remove superfluous columns
  mutate(Filename = str_remove(string = Filename, pattern = "_fit.csv")) %>%
  mutate(Filename = str_remove(string = Filename, pattern = "../Data/RawData/Sackville/")) %>%
  separate(Filename, into=c("YYYYMMDD", "CultureID", "Temp_C", "PulseSpace_s"), sep = "([\\/\\_])", remove = FALSE) %>%
  mutate(YYYYMMDD = lubridate::ymd(YYYYMMDD),
          TIME = as.character(TIME)) %>%
   rename(ObsDate = DATE,
         ObsTime = TIME,
         FvFm = "Fv.Fm") |>
  mutate(Temp_C = str_remove(Temp_C, "C")) |>
  type_convert() %>%
  mutate(PulseSpace_s = str_remove(PulseSpace_s, "S")) %>%  #s vs S
  mutate(PulseSpace_s = str_remove(PulseSpace_s, "s")) %>% 
  filter(PulseSpace_s %in% c(1, 2, 4, 8, 16)) |>  #remove rows with other PulseSpace_s
  mutate(PulseSpace_s = as.numeric(PulseSpace_s)) %>%
  mutate(ObsDateTime = lubridate::ymd_hms(paste(ObsDate, ObsTime))) %>%
  relocate(ObsDateTime, .after = ObsTime) %>%
  relocate(CultureID, .before = ObsDate) |>
  type_convert()

SolFitsTrim <-SolFitsTrim %>%
  group_by(Filename) %>%
  mutate(Flashnumber = row_number()) %>%
  relocate(Flashnumber, .before=p) %>%
  mutate(Etime_s = (as.numeric(Time__mSec) - min(as.numeric(Time__mSec), na.rm = TRUE))/1000, .before=Flashnumber) %>%
  ungroup()
```


```{r correct file issues}
ProblemFiles <- unique(SolFitsTrim$Filename[SolFitsTrim$Flashnumber >32])

ProblemList <- SolFitsTrim %>% 
  filter(Filename %in% c(ProblemFiles)) %>% 
  group_by(Filename) %>%
  mutate(TimeDiff = c(NA, diff(ObsTime))) %>% #shows actual spacing between flashes
  ungroup() %>%
  doBy::split_by("Filename") #creates list of problem files
```


From ProblemList, use spacing between flashes to decipher runs:
  20230911_NaOm1663_22C_2S        - 58 flashes, no good runs 
  20230911_NaOm1663_22C_4S        - 34 flashes, only first 29 are good
  
  20230912_NaOm1663_18C_ 8S, 4S, 2S, & 1S     - 96 flashes, saved exact same run 3x; bad
  

```{r manual trim}
first32 <- SolFitsTrim %>%
  filter(!Filename %in% c('20230911_NaOm1663_22C_2S', "20230911_NaOm1663_22C_4S", "20230912_NaOm1663_18C_1S",
                          "20230912_NaOm1663_18C_2S", "20230912_NaOm1663_18C_4S", "20230912_NaOm1663_18C_8S",
                          "20240312_NaOm2987_26C_1S"))  %>%
  group_by(Filename) %>%
  filter(Flashnumber %in% c(1:32)) %>%
  mutate(Flashnumber = row_number()) %>%
  ungroup()

NaOm1663_22C_4S <- SolFitsTrim %>%
  filter(Filename == '20230911_NaOm1663_22C_4S') %>%
  filter(Flashnumber %in% c(1:29)) %>%
  mutate(Flashnumber = row_number())

SolFitsTrim2 <- rbind(first32, NaOm1663_22C_4S)
```


### Normalize Fluorescence & Add Temperature Differential Column

```{r tidy clean data}
SolFitsTrim2 <- SolFitsTrim2 %>%
  mutate(SState = case_when(Flashnumber %% 4 == 1 ~ 1,
                    Flashnumber %% 4 == 2 ~ 2,
                    Flashnumber %% 4 == 3 ~ 3, 
                    Flashnumber %% 4 == 0 ~ 4)) #add column for sstate

#normalize values to final stretch of 10 values from 'randomized' PSII pool, done after cleaning to ensure the last 10 values are correct 
SolFitsTrim2 <- SolFitsTrim2 %>%
  group_by(Filename) %>%
  mutate(Sigrunnorm =  Sig/mean(Sig[(length(Sig) - 10):length(Sig)], na.rm = TRUE),
         Forunnorm = Fo/mean(Fo[(length(Fo) - 10):length(Fo)], na.rm = TRUE),
         Fmrunnorm = Fm/mean(Fm[(length(Fm) - 10):length(Fm)], na.rm = TRUE),
         FvFmrunnorm = FvFm/mean(FvFm[(length(FvFm) - 10):length(FvFm)], na.rm = TRUE)) %>%
  ungroup()

# Add columns for growth temperature and difference between measurement & growth temps
SolFitsTrim2 <- SolFitsTrim2 %>%
  mutate(GrowthTemp_C = 22) %>%
  mutate(TempDiff_C = Temp_C - GrowthTemp_C)
```

### Group by Taxon 
```{r}
SolFitsChlorella <- SolFitsTrim2 %>%
  filter(CultureID %in% c("NaOm1671", "NaOm2987","NaOm2990"))

SolFitsPseudonana <- SolFitsTrim2 %>%
  filter(CultureID %in% c("NaOm1305", "NaOm1663", "NaOm2988", "NaOm2992", "NaOm2991"))

SolFitsWeissflogii <- SolFitsTrim2 %>%
  filter(CultureID == "NaOm1293")
```

### Calculate Light Levels
```{r}
# add column for light level (µmol photons/m2/s)
SolFitsChlorella <- conversion(SolFitsChlorella)
SolFitsPseudonana <- conversion(SolFitsPseudonana) 
SolFitsWeissflogii <- conversion(SolFitsWeissflogii) 
```

### Test Plots
```{r}
SolFitsChlorella %>%
  ggplot()+
  geom_line(aes(x=Flashnumber, y=FvFmrunnorm, colour = as.factor(Filename))) +
  facet_grid(rows = vars(PulseSpace_s), cols = vars(TempDiff_C)) +
  theme_bw() +
  theme(legend.position = "none")

SolFitsPseudonana %>%
  ggplot()+
  geom_line(aes(x=Flashnumber, y=FvFmrunnorm, colour = as.factor(Filename))) +
  facet_grid(rows = vars(PulseSpace_s), cols = vars(TempDiff_C)) +
  theme_bw() +
  theme(legend.position = "none")
```


```{r save Data}
saveRDS(SolFitsChlorella, file.path(DataOut, paste(Project,"Chlorella_vulgaris", "SolFitsTrim.Rds", sep = "_"), fsep = .Platform$file.sep))

saveRDS(SolFitsPseudonana, file.path(DataOut, paste(Project,"Thalassiosira_pseudonana", "SolFitsTrim.Rds", sep = "_"), fsep = .Platform$file.sep))
```


