---
title: "TakuvikImport"
author:
- Natasha Ryan
- Maximilian Berthold
- Douglas A. Campbell
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
    code_folding: hide
    keep_md: yes
    fig_caption: yes
    toc: TRUE
    toc_float: TRUE   
csl: plos-one.csl
editor_options: 
  chunk_output_type: inline
---

*Some of the code used to create this R Notebook was refurbished from "PlateAbImport.Rmd" written by Maximilian Berthold, Douglas A. Campbell, Melissa L. Rioux, Sarah J Gore, and Alyson MacCormack.*

```{r load libraries}
library(tidyverse)
library(lubridate)
library(googlesheets4)
library(googledrive)
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_chunk$set(fig.path='Output/')
```

# Set Project Variables 
```{r set project variables}
Project <- "LOW"
TargetData <- "Takuvik"
DataOut <- file.path("..","Data", "CleanData")
CalibData <- file.path("..","Data", "CalibData")


DataIn <- file.path("..","Data", "RawData", TargetData, fsep = .Platform$file.sep)
DataRefDecay <- file.path("..","Data", "RawData",TargetData, "DecayRef", fsep = .Platform$file.sep)
DataRLC <- file.path("..","Data", "RawData",TargetData, "RLC", fsep = .Platform$file.sep)

FileID <- "fit"

FileEncode <- "UTF-8" #guess_encoding says 'ASCII"

Delimiter <- ","

HeaderRows <- 0
```

```{r conversions}
us_s = 1000000
photons_umol = 6.022E17
A2_m2 = 1E20
```

```{r read ActPAR calibration files}
#ActPARCal <- readRDS("~/Dropbox/CampbellLabProtocols/ChlorophyllFluorescence/SolisenseInformation/SolisenseInformation_DCCalibParam.Rds")

ActPARCrossCal <- list.files(path = CalibData, full.names = TRUE) %>%
     map_df(~readRDS(file  = .))

#intercept set to 0 in lm in SolisenseInformation.Rproj/SolisenseCalibCompare.Rmd
ActPARCrossCal <- ActPARCrossCal |>
  rename(#Intercept = `estimate_(Intercept)`,
         Slope = `estimate_LIFT_Gen_Developer.cal`,
         #Intercept_SE = `std.error_(Intercept)`,
         Slope_SE = `std.error_LIFT_Gen_Developer.cal`)
```

# Metadata Import
```{r import metadata}
googlesheets4::gs4_deauth()

TakuvikMetadata <- read_sheet("https://docs.google.com/spreadsheets/d/1OC22KJy-VdQU9r1-PYvQaJO42_ugdFcHmRHWbMYOHUc/edit?usp=sharing")
```

# File Import
```{r list PSI files for file import}
SolisenseFiles <- list.files(path = DataIn, pattern = FileID, full.names = TRUE)
SolisenseFiles

#test for duplicate file names
unique(duplicated(SolisenseFiles))

RLCFiles <- list.files(path = DataRLC, pattern = FileID, full.names = TRUE)
RLCFiles

```

```{r warning=FALSE, message=FALSE, echo=FALSE}
read.delim_plus <- function(flnm, file_encode, delimiter, header_rows){read.delim(flnm, fileEncoding = file_encode, sep = delimiter,  skip = header_rows, row.names = NULL) %>% mutate(Filename = flnm)
}
```

# Create Data Frame 
purrr::map to read all files
```{r read Solisense files}
SolFits <- SolisenseFiles %>%
  map_df(~read.delim_plus(flnm =., file_encode = FileEncode, delimiter = Delimiter, header_rows = HeaderRows))

RLCFits <- RLCFiles %>%
  map_df(~read.delim_plus(flnm =., file_encode = FileEncode, delimiter = Delimiter, header_rows = HeaderRows))

```


```{r tidy data}
#for NoDecay Data
SolFitsTrim <- SolFits %>%
  filter(!grepl("----", DATE)) %>% # remove rows with "----" 
  select(-c("RFID_User_Data", "Barcode_Data","PIF", "LEDSel", "Dur.Wat", "Lon", "Lat", "GPS_stat","X", "Light_2", "Light_3","Light_4","Light_5","Light_6")) #remove extra columns

PrefacesToFix <- c("20230926_SeGu1008_0C", "20230926_SeGu1008_10C", "20230927_SeGu1008_10C", "20230927_SeGu1008_6C","20230927_SeGu1008_2C","20230927_SeGu1008_0C")

SolFitsBadNames <- SolFitsTrim %>%
  filter(grepl(paste(PrefacesToFix, collapse="|"), Filename)) %>%
  mutate(across("Filename", str_replace, "1008","1006"))

SolFitsGoodNames <- SolFitsTrim %>%
  filter(!grepl(paste(PrefacesToFix, collapse="|"), Filename))

SolFitsTrim <- rbind(SolFitsBadNames, SolFitsGoodNames)

SolFitsTrim <- SolFitsTrim %>%
  mutate(Filename = str_remove(string = Filename, pattern = "_fit.csv")) %>%
  mutate(Filename = str_remove(string = Filename, pattern = "../Data/RawData/")) %>%
  separate(Filename, into=c("Location", "YYYYMMDD", "CultureID", "Temp_C", "PulseSpace_s"), sep = "([\\/\\_])", remove = FALSE) %>%
  mutate(YYYYMMDD = lubridate::ymd(YYYYMMDD)) %>%
  rename("ObsDate" = "DATE", "ObsTime" = "TIME", "FvFm" = "Fv.Fm") %>%
  mutate(Temp_C = str_remove(Temp_C, "C")) %>%
  type_convert() %>%
  mutate(PulseSpace_s = str_remove(PulseSpace_s, "s")) %>%  #s vs S
  filter(PulseSpace_s %in% c(1, 2, 4, 8, 16)) %>%  #remove rows with other PulseSpace_s
  mutate(PulseSpace_s = as.numeric(PulseSpace_s)) %>%
  mutate(ObsDateTime = lubridate::ymd_hms(paste(ObsDate, ObsTime))) %>%
  relocate(ObsDateTime, .after = ObsTime) %>%
  relocate(CultureID, .before = ObsDate) %>%
  type_convert()

#normalize values to final stretch of 10 values from 'randomized' PSII pool
SolFitsTrim <- SolFitsTrim %>%
  group_by(Filename) %>%
  #arrange(ObsTime)%>%
  mutate(Sigrunnorm =  Sig/mean(Sig[(length(Sig) - 10):length(Sig)], na.rm = TRUE),
         Forunnorm = Fo/mean(Fo[(length(Fo) - 10):length(Fo)], na.rm = TRUE),
         Fmrunnorm = Fm/mean(Fm[(length(Fm) - 10):length(Fm)], na.rm = TRUE),
         FvFmrunnorm = FvFm/mean(FvFm[(length(FvFm) - 10):length(FvFm)], na.rm = TRUE)) %>%
  mutate(Flashnumber = row_number()) %>%
  mutate(Flashnumber2 = rank(ObsTime)) %>%
  relocate(Flashnumber, .before=p) %>%
  mutate(Etime_s = (as.numeric(Time__mSec) - min(as.numeric(Time__mSec), na.rm = TRUE))/1000, .before=Flashnumber) %>%
  ungroup()
```


```{r tidy RLC Fits}
#for NoDecayDataset
RLCFitsTrim <- RLCFits |>
  filter(!grepl("----", DATE)) |> # remove rows with "----" 
  select(-c("RFID_User_Data", "Barcode_Data","PIF", "LEDSel", "Dur.Wat", "Lon", "Lat", "GPS_stat","X"))  # remove superfluous columns

RLCFitsBadNames <- RLCFitsTrim %>%
  filter(grepl(paste(PrefacesToFix, collapse="|"), Filename)) %>%
  mutate(across("Filename", str_replace, "1008","1006"))

RLCFitsGoodNames <- RLCFitsTrim %>%
  filter(!grepl(paste(PrefacesToFix, collapse="|"), Filename))

RLCFitsTrim <- rbind(RLCFitsBadNames, RLCFitsGoodNames)

RLCFitsTrim <- RLCFitsTrim |>
  mutate(Filename = str_remove(string = Filename, pattern = "_RLC_fit.csv")) |>
    mutate(Filename = str_remove(string = Filename, pattern = "../Data/RawData/")) |>
  separate(Filename, into=c("Location", "Type", "YYYYMMDD", "CultureID", "Temp_C"), sep = "([\\/\\_])", remove = FALSE) %>%
  mutate(YYYYMMDD = lubridate::ymd(YYYYMMDD),
          TIME = as.character(TIME)) %>%
   rename(ObsDate = DATE,
         ObsTime = TIME,
         FvFm = "Fv.Fm") |>
  mutate(Temp_C = str_remove(Temp_C, "C")) |>
  type_convert() |>
  mutate(ObsDateTime = lubridate::ymd_hms(paste(ObsDate, ObsTime))) %>%
  relocate(ObsDateTime, .after = ObsTime) %>%
  relocate(CultureID, .before = ObsDate) |>
  type_convert()


RLCFitsTrim <- RLCFitsTrim |>
  group_by(Filename) |>
  arrange(ObsTime) |>
  mutate(Etime_s = (as.numeric(Time__mSec) - min(as.numeric(Time__mSec), na.rm = TRUE))/1000) %>%
  ungroup()

RLCFitsTrim <- RLCFitsTrim |>
  rowwise() |>
  mutate(PAR_ue = sum(Light_1, Light_2, Light_3, Light_4, Light_5, Light_6)) |>
  group_by(Filename) |>
  mutate(Dark1s = case_when(row_number() == 1 ~ 0,
                            row_number() %% 2 == 0 & row_number() > 1 ~ 0,
                            row_number() %% 2 == 1 & row_number() > 1 ~ 1)) |>
  ungroup()
```


#Takuvik File Issues 
```{r}
SolFitsTrim %>%
  filter(Flashnumber > 32)
```


20230926_SeGu1001_6C_16s  - last 32 are a run, but spacing is 19s? bad run, computer should have replacement 

**To fix**
20230925_SeGu1005_2C_16s  - keep flashes 58-89

code all of those to keep the last 32 
20230925_SeGu1002_2C_16s  - remove first flash 
20230926_SeGu1001_10C_8s  - remove first flash 
20230927_SeGu1001_0C_16s  - remove first flash
20230927_SeGu1006_0C_16s  - remove first flash
20230925_SeGu1003_2C_16   - keep last 32
20230925_SeGu1004_2C_16s  - keep last 32
20230925_SeGu1006_2C_16s  - keep last 32
20230926_SeGu1001_0C_16s  - keep last 32
20230926_SeGu1006_10C_16s - keep last 32
20230927_SeGu1001_10C_16s - keep last 32
20230926_SeGu1008_6C_16s  - keep last 32
20230927_SeGu1006_6C_16s  - keep last 32
20230925_SeGu1005_2C_8s   - 2 full runs, keep second 
20230927_SeGu1006_2C_16s  - contains 2 full runs, keep second 

```{r manual trim}
SolFitsTrim_last32 <- SolFitsTrim %>%
  filter(!Filename %in% c('Takuvik/20230925_SeGu1005_2C_16s', 'Takuvik/20230926_SeGu1001_6C_16s')) %>% 
  group_by(Filename) %>%
  slice(tail(row_number(),32)) %>% 
  mutate(Flashnumber2 = row_number()) %>%
  ungroup()

SolFitsTrimSeGu1005_2C_16s <- SolFitsTrim %>%
  filter(Filename == 'Takuvik/20230925_SeGu1005_2C_16s') %>%
  filter(Flashnumber %in% c(58:89)) %>%
  mutate(Flashnumber2 = row_number())
  
SolFitsTrim2 <- rbind(SolFitsTrim_last32, SolFitsTrimSeGu1005_2C_16s)
```

# Group by Taxon 

```{r}
SolFitsChaeotoceros <- SolFitsTrim2 %>%
  filter(CultureID %in% c("SeGu1002", "SeGu1007", "SeGu1004", "SeGu1005"))

SolFitsFragilariopsis <- SolFitsTrim2 %>%
  filter(CultureID %in% c("SeGu1001","SeGu1003","SeGu1006","SeGu1008"))
```

Fragilariopsis
4 trials for each interval at 0C and 10C
    - except 20230927_SeGu1006_10C_16s only has 31
6 runs at 2C for 1,2,4,8s
    - 7 for 16s
5 runs at 6C for 1,2,4, 16s

Chaeotoceros (only done at 2C)
3 runs at 1, 2, 4, 16s 
2 runs at 8s 


# Test Plot
```{r}
SolFitsFragilariopsis %>%
  filter(Temp_C %in% c("0","10")) %>%
  filter(PulseSpace_s %in% c("1","4")) %>%
  filter(CultureID == "SeGu1001") |>
  filter(ObsDate == "2023-09-26") |>
  ggplot() +
  geom_line(aes(x=Flashnumber2, y=FvFm)) +
  facet_grid(rows = vars(Temp_C), cols = vars(PulseSpace_s)) +
  labs(title = "Fragilariopsis_cylindrus", x = "Flash Number", y = "Maximum Quantum Yield of PSII") +
  theme_bw()


ggsave("FragilariopsisFlashCyclesDecay.png",
  plot = last_plot(),
  path = file.path("..", "Output"))
```

```{r}
SolFitsChaeotoceros %>%
  ggplot() +
  geom_line(aes(x=Flashnumber2, y=FvFm, colour = as.factor(CultureID))) +
  facet_grid(rows = vars(PulseSpace_s), cols = vars(Temp_C)) +
  labs(title = "Chaeotoceros_gelidus") +
  theme_bw()
```

```{r}
RLCFitsTrim %>%
  ggplot()+
  geom_line(aes(x = PAR_ue, y = FvFm)) +
  facet_grid(rows = vars(Temp_C), cols = vars(CultureID)) +
  labs(title = TargetData) +
  theme_bw()
```


```{r save Data}
saveRDS(SolFitsChaeotoceros, file.path(DataOut, paste(Project, TargetData, "Chaeotoceros", "SolFitsTrim.Rds", sep = "_"), fsep = .Platform$file.sep))

saveRDS(SolFitsFragilariopsis, file.path(DataOut, paste(Project, TargetData, "Fragilariopsis", "SolFitsTrim.Rds", sep = "_"), fsep = .Platform$file.sep))
```

