---
title: "TakuvikImport"
author:
- Natasha Ryan
- Maximilian Berthold
- Douglas A. Campbell
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
    code_folding: hide
    keep_md: yes
    fig_caption: yes
    toc: TRUE
    toc_float: TRUE   
csl: plos-one.csl
editor_options: 
  chunk_output_type: inline
---

*Some of the code used to create this R Notebook was refurbished from "PlateAbImport.Rmd" written by Maximilian Berthold, Douglas A. Campbell, Melissa L. Rioux, Sarah J Gore, and Alyson MacCormack.*

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
knitr::opts_chunk$set(fig.path='Output/')
```

```{r load libraries}
library(tidyverse)
library(lubridate)
library(googlesheets4)
library(googledrive)
```

#Setup

### Set Project Variables 
```{r set project variables}
Project <- "LOW"
TargetData <- "Takuvik"
DataOut <- file.path("..","Data", "CleanData")
CalibData <- file.path("..","Data", "CalibData")


DataIn <- file.path("..","Data", "RawData", TargetData, fsep = .Platform$file.sep)
DataRefDecay <- file.path("..","Data", "RawData",TargetData, "DecayRef", fsep = .Platform$file.sep)
DataRLC <- file.path("..","Data", "RawData",TargetData, "RLC", fsep = .Platform$file.sep)

FileID <- "fit"

FileEncode <- "UTF-8" #guess_encoding says 'ASCII"

Delimiter <- ","

HeaderRows <- 0

#conversions 
#us_s = 1000000
#photons_umol = 6.022E17
#A2_m2 = 1E20
```


```{r read ActPAR calibration files}
# #ActPARCal <- readRDS("~/Dropbox/CampbellLabProtocols/ChlorophyllFluorescence/SolisenseInformation/SolisenseInformation_DCCalibParam.Rds")
# 
# ActPARCrossCal <- list.files(path = CalibData, full.names = TRUE) %>%
#      map_df(~readRDS(file  = .))
# 
# #intercept set to 0 in lm in SolisenseInformation.Rproj/SolisenseCalibCompare.Rmd
# ActPARCrossCal <- ActPARCrossCal |>
#   rename(#Intercept = `estimate_(Intercept)`,
#          Slope = `estimate_LIFT_Gen_Developer.cal`,
#          #Intercept_SE = `std.error_(Intercept)`,
#          Slope_SE = `std.error_LIFT_Gen_Developer.cal`)
```

### Metadata Import
```{r import metadata}
googlesheets4::gs4_deauth()

TakuvikMetadata <- read_sheet("https://docs.google.com/spreadsheets/d/1OC22KJy-VdQU9r1-PYvQaJO42_ugdFcHmRHWbMYOHUc/edit?usp=sharing")
```

### Create Functions 
```{r warning=FALSE, message=FALSE, echo=FALSE}
read.delim_plus <- function(flnm, file_encode, delimiter, header_rows){read.delim(flnm, fileEncoding = file_encode, sep = delimiter,  skip = header_rows, row.names = NULL) %>% mutate(Filename = flnm)
}
```


# No Decay Data

### Import 
```{r list PSI files for file import}
SolisenseFiles <- list.files(path = DataIn, pattern = FileID, full.names = TRUE)
SolisenseFiles

unique(duplicated(SolisenseFiles)) #test for duplicate file names
```

```{r read Solisense files}
SolFits <- SolisenseFiles %>%
  map_df(~read.delim_plus(flnm =., file_encode = FileEncode, delimiter = Delimiter, header_rows = HeaderRows)) #purrr::map to read all files
```

### Tidy
```{r correct prefaces}
#correct errors in file prefaces 
ProblemFiles <- c("20230926_SeGu1008_0C", "20230926_SeGu1008_10C", "20230927_SeGu1008_10C", "20230927_SeGu1008_6C","20230927_SeGu1008_2C","20230927_SeGu1008_0C")

SolFitsBadNames <- SolFits %>%
  filter(grepl(paste(ProblemFiles, collapse="|"), Filename)) %>%
  mutate(Filename = str_replace(Filename, "1008", "1006")) #isolate and correct incorrect files 

SolFitsGoodNames <- SolFits %>%
  filter(!grepl(paste(ProblemFiles, collapse="|"), Filename)) #isolate unaffected files 

SolFits <- rbind(SolFitsBadNames, SolFitsGoodNames) #merge
```

```{r tidy data}
#for NoDecay Data
SolFitsTrim <- SolFits %>%
  filter(!grepl("----", DATE)) %>% # remove rows with "----" 
  select(-c("RFID_User_Data", "Barcode_Data","PIF", "LEDSel", "Dur.Wat", "Lon", "Lat", "GPS_stat","X","Light_1", "Light_2", "Light_3","Light_4","Light_5","Light_6")) %>% #remove extra columns
  mutate(Filename = str_remove(string = Filename, pattern = "_fit.csv")) %>%
  mutate(Filename = str_remove(string = Filename, pattern = "../Data/RawData/Takuvik/")) %>%
  separate(Filename, into=c("YYYYMMDD", "CultureID", "Temp_C", "PulseSpace_s"), sep = "([\\/\\_])", remove = FALSE) %>%
  mutate(YYYYMMDD = lubridate::ymd(YYYYMMDD)) %>%
  rename("ObsDate" = "DATE", "ObsTime" = "TIME", "FvFm" = "Fv.Fm") %>%
  mutate(Temp_C = str_remove(Temp_C, "C")) %>%  
  type_convert() %>%
  mutate(PulseSpace_s = str_remove(PulseSpace_s, "s")) %>%
  mutate(PulseSpace_s = as.numeric(PulseSpace_s)) %>%
  mutate(ObsDateTime = lubridate::ymd_hms(paste(ObsDate, ObsTime))) %>%
  relocate(ObsDateTime, .after = ObsTime) %>%
  relocate(CultureID, .before = ObsDate) %>%
  type_convert() 

SolFitsTrim <- SolFitsTrim %>%
  group_by(Filename) %>%
  mutate(Flashnumber = row_number()) %>% #add flashnumber column 
  #mutate(Flashnumber2 = rank(ObsTime)) %>%
  relocate(Flashnumber, .before=p) %>%
  mutate(Etime_s = (as.numeric(Time__mSec) - min(as.numeric(Time__mSec), na.rm = TRUE))/1000, .before=Flashnumber) %>% #add col for time elapsed 
  ungroup()
```

```{r correct file issues}
ProblemFiles <- unique(SolFitsTrim$Filename[SolFitsTrim$Flashnumber >32])

ProblemList <- SolFitsTrim %>% 
  filter(Filename %in% c(ProblemFiles)) %>% 
  group_by(Filename) %>%
  mutate(TimeDiff = c(NA, diff(ObsTime))) %>% #shows actual spacing between flashes
  ungroup() %>%
  doBy::split_by("Filename") #creates list of problem files
```

From ProblemList, use spacing between flashes to decipher runs:
Keep last 32 flashes in:
  20230925_SeGu1003_2C_16s        - 49 flashes
  20230925_SeGu1004_2C_16s        - 37 flashes
  20230926_SeGu1006_10C_16s       - 34 flashes 
  20230927_SeGu1006_0C_16s        - 33 flashes 
  20230927_SeGu1006_6C_16s        - 38 flashes
  20230925_SeGu1002_2C_16s        - 33 flashes
  20230925_SeGu1006_2C_16s        - 34 flashes 
  20230925_SeGu1008_2C_16s        - 34 flashes
  20230926_SeGu1001_0C_16s        - 34 flashes 
  20230926_SeGu1001_10C_8s        - 33 flashes
  20230926_SeGu1003_6C_16s        - 70 flashes
  20230926_SeGu1006_6C_16s        - 40 flashes
  20230926_SeGu1008_6C_16s        - 39 flashes
  20230927_SeGu1001_0C_16s        - 33 flashes
  20230927_SeGu1001_10C_16s       - 34 flashes 

Other adjustments: 
  20230927_SeGu1006_2C_16s        - 65 flashes, keep 2-33
  20230925_SeGu1005_2C_16s        - 120 flashes, use 58-89
  20230926_SeGu1001_6C_16s        - remove, incorrect spacing
  20230925_SeGu1005_2C_8s         - remove, file error 



```{r manual trim}
last32 <- SolFitsTrim %>%
  filter(!Filename %in% c('20230927_SeGu1006_2C_16s', '20230925_SeGu1005_2C_16s', '20230926_SeGu1001_6C_16s', '20230925_SeGu1005_2C_8s')) %>%
  group_by(Filename) %>%
  slice(tail(row_number(),32)) %>% 
  mutate(Flashnumber = row_number()) %>%
  ungroup()

SeGu1005_2C_16s <- SolFitsTrim %>%
  filter(Filename == '20230925_SeGu1005_2C_16s') %>%
  filter(Flashnumber %in% c(58:89)) %>%
  mutate(Flashnumber = row_number())

SeGu1006_2C_16s <- SolFitsTrim %>%
  filter(Filename == '20230927_SeGu1006_2C_16s') %>%
  filter(Flashnumber %in% c(2:33)) %>%
  mutate(Flashnumber = row_number())
  
SolFitsTrim2 <- rbind(last32,SeGu1005_2C_16s, SeGu1006_2C_16s)
```

### Normalize
Now that the files are clean: 

```{r tidy clean data}
SolFitsTrim2 <- SolFitsTrim2 %>%
  mutate(SState = case_when(Flashnumber %% 4 == 1 ~ 1,
                    Flashnumber %% 4 == 2 ~ 2,
                    Flashnumber %% 4 == 3 ~ 3, 
                    Flashnumber %% 4 == 0 ~ 4)) #add column for sstate

#normalize values to final stretch of 10 values from 'randomized' PSII pool, done after cleaning to ensure the last 10 values are correct 
SolFitsTrim2 <- SolFitsTrim2 %>%
  group_by(Filename) %>%
  mutate(Sigrunnorm =  Sig/mean(Sig[(length(Sig) - 10):length(Sig)], na.rm = TRUE),
         Forunnorm = Fo/mean(Fo[(length(Fo) - 10):length(Fo)], na.rm = TRUE),
         Fmrunnorm = Fm/mean(Fm[(length(Fm) - 10):length(Fm)], na.rm = TRUE),
         FvFmrunnorm = FvFm/mean(FvFm[(length(FvFm) - 10):length(FvFm)], na.rm = TRUE)) %>%
  ungroup()
```


### Group by Taxon 
```{r}
SolFitsChaeotoceros <- SolFitsTrim2 %>%
  filter(CultureID %in% c("SeGu1002", "SeGu1007", "SeGu1004", "SeGu1005"))

SolFitsFragilariopsis <- SolFitsTrim2 %>%
  filter(CultureID %in% c("SeGu1001","SeGu1006"))
#Note SeGu1003 and 1008 were excluded due to culture issues 
```

```{r add deviation from growth temp}
SolFitsFragilariopsis <- SolFitsFragilariopsis %>%
  mutate(GrowthTemp_C = ifelse(CultureID == "SeGu1001", 0, ifelse(CultureID == "SeGu1006", 6, NA)))%>%
  mutate(TempDiff_C = Temp_C - GrowthTemp_C)
```

```{r split by growth temp}
SolFitsFragilariopsisG0 <- SolFitsFragilariopsis %>%
  filter(GrowthTemp_C == 0)

SolFitsFragilariopsisG6 <- SolFitsFragilariopsis %>%
  filter(GrowthTemp_C == 6)
```

```{r}
SolFitsFragilariopsis %>%
  ggplot()+
  geom_line(aes(x=Flashnumber, y=Fmrunnorm, colour = as.factor(Filename))) +
  facet_grid(rows = vars(PulseSpace_s), cols = vars(Temp_C)) +
  theme_bw() +
  theme(legend.position = "none")
```


```{r save Data}
saveRDS(SolFitsChaeotoceros, file.path(DataOut, paste(Project,"Chaeotoceros_gelidus", "SolFitsTrim.Rds", sep = "_"), fsep = .Platform$file.sep))

saveRDS(SolFitsFragilariopsisG0, file.path(DataOut, paste(Project,"Fragilariopsis_cylindrus","G0", "SolFitsTrim.Rds", sep = "_"), fsep = .Platform$file.sep))

saveRDS(SolFitsFragilariopsisG6, file.path(DataOut, paste(Project,"Fragilariopsis_cylindrus","G6", "SolFitsTrim.Rds", sep = "_"), fsep = .Platform$file.sep))
```


# RLC Data
### Import
```{r list RLC files for import}
RLCFiles <- list.files(path = DataRLC, pattern = FileID, full.names = TRUE)
RLCFiles
```

```{r create df }
RLCFits <- RLCFiles %>%
  map_df(~read.delim_plus(flnm =., file_encode = FileEncode, delimiter = Delimiter, header_rows = HeaderRows))
```

### Tidy
```{r tidy RLC Fits}
#for NoDecayDataset
RLCFitsTrim <- RLCFits |>
  filter(!grepl("----", DATE)) |> # remove rows with "----" 
  select(-c("RFID_User_Data", "Barcode_Data","PIF", "LEDSel", "Dur.Wat", "Lon", "Lat", "GPS_stat","X"))  # remove superfluous columns

RLCFitsBadNames <- RLCFitsTrim %>%
  filter(grepl(paste(PrefacesToFix, collapse="|"), Filename)) %>%
  mutate(across("Filename", str_replace, "1008","1006"))

RLCFitsGoodNames <- RLCFitsTrim %>%
  filter(!grepl(paste(PrefacesToFix, collapse="|"), Filename))

RLCFitsTrim <- rbind(RLCFitsBadNames, RLCFitsGoodNames)

RLCFitsTrim <- RLCFitsTrim |>
  mutate(Filename = str_remove(string = Filename, pattern = "_RLC_fit.csv")) |>
    mutate(Filename = str_remove(string = Filename, pattern = "../Data/RawData/")) |>
  separate(Filename, into=c("Location", "Type", "YYYYMMDD", "CultureID", "Temp_C"), sep = "([\\/\\_])", remove = FALSE) %>%
  mutate(YYYYMMDD = lubridate::ymd(YYYYMMDD),
          TIME = as.character(TIME)) %>%
   rename(ObsDate = DATE,
         ObsTime = TIME,
         FvFm = "Fv.Fm") |>
  mutate(Temp_C = str_remove(Temp_C, "C")) |>
  type_convert() |>
  mutate(ObsDateTime = lubridate::ymd_hms(paste(ObsDate, ObsTime))) %>%
  relocate(ObsDateTime, .after = ObsTime) %>%
  relocate(CultureID, .before = ObsDate) |>
  type_convert()


RLCFitsTrim <- RLCFitsTrim |>
  group_by(Filename) |>
  arrange(ObsTime) |>
  mutate(Etime_s = (as.numeric(Time__mSec) - min(as.numeric(Time__mSec), na.rm = TRUE))/1000) %>%
  ungroup()

RLCFitsTrim <- RLCFitsTrim |>
  rowwise() |>
  mutate(PAR_ue = sum(Light_1, Light_2, Light_3, Light_4, Light_5, Light_6)) |>
  group_by(Filename) |>
  mutate(Dark1s = case_when(row_number() == 1 ~ 0,
                            row_number() %% 2 == 0 & row_number() > 1 ~ 0,
                            row_number() %% 2 == 1 & row_number() > 1 ~ 1)) |>
  ungroup()
```


```{r}
RLCFitsTrim %>%
  ggplot()+
  geom_line(aes(x = PAR_ue, y = FvFm)) +
  facet_grid(rows = vars(Temp_C), cols = vars(CultureID)) +
  labs(title = TargetData) +
  theme_bw()
```

# Other Code 
Joining Chl and species from metadata with SolFits to normalize for chl
```{r}
# ChlData <- TakuvikMetadata %>%
#   select(c("ExpDate","ID","Strain", "Chl_ugmL")) %>%
#   drop_na() %>%
#   rename("YYYYMMDD"= "ExpDate", "Species"="Strain") %>%
#   mutate(YYYYMMDD = lubridate::ymd(YYYYMMDD)) %>%
#   unite("StrainCode", c("YYYYMMDD","ID"))
# 
# SolFitsTrim3 <- SolFitsTrim2 %>%
#   unite("StrainCode", c("YYYYMMDD","CultureID"), remove = FALSE)
# 
# JoinedwithChl <- left_join(SolFitsTrim3, ChlData)
```

