---
title: "Wavelet Transformation"
author:
- Natasha Ryan
- Maximilian Berthold
- Douglas Campbell
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
    code_folding: hide
    keep_md: yes
    fig_caption: yes
    toc: TRUE
    toc_float: TRUE   
csl: plos-one.csl
editor_options: 
  chunk_output_type: inline
---

# Introduction
This RMD computes a wave transformation and extracts the duration of fluorescence oscillations for the cleaned data generated from OttawaImport, TakuvikImport, and SackvilleImport. 

```{r load libraries}
library(tidyverse)
library(doBy)
library(WaveletComp)
```


# Set Project Variables
```{r set project variables}
Project <- "LOW"
DataIn <- file.path("..","Data", "CleanData")
DataOut <- file.path("..","Data", "ProcessedData")
FileID <- "SolFitsMeta"
siglvl <- 0.05 #specifies statistical significance level
Fc <- "Fragilariopsis_cylindrus"
Ci <- "Chlamydomonas_ICEMDV"
Cm <- "Chlamydomonas_malina"
Cp <- "Chlamydomonas_priscuii"
Cr <- "Chlamydomonas_reinhardtii"
Cv <- "Chlorella_vulgaris"
Tp <- "Thalassiosira_pseudonana"
Ce <- "Chlamydomonas_euryale"
Ck <- "Chlamydomonas_klinobasis"
```


```{r create functions}
#function for wavelet transformation 
wavelet <- function(data_list){
  wavelet_list <- lapply(data_list, function(data) {
    WaveletComp::analyze.wavelet(data, "FvFmrunnorm",
                                 method = "white.noise",
                                 loess.span = 0.85, 
                                 dt = 1, dj=1/500, 
                                 n.sim = 100, 
                                 verbose = FALSE)
  })
return(wavelet_list)
}


#function for wavelet reconstruction, applies reconstruct() function from WaveletComp to all wavelets in a list
reconstructwave <- function(wavelet_list) {
  reconstructed_list <- lapply(wavelet_list, function(wave) {
    WaveletComp::reconstruct(wave, siglvl = siglvl, 
                             #plot.waves = TRUE,
                             plot.rec = FALSE,
                             lwd = 1.5, 
                             col=c("#0E1126","#429F78"), 
                             legend.coords = "topright",
                             legend.text = c("Original", "Reconstructed"),
                             timelab = "Flash Number", 
                             verbose = FALSE)
 })
  return(reconstructed_list) 
}


#function to extract the number of oscillating flashes from the reconstruction 
damping <- function(reconstructed) {
  result_list <- lapply(reconstructed, function(condition) {
    return(condition$series)
  }) #extract reconstructed wave data from 'reconstruct' objects 
  result_df <- data.frame(condition = character(0), damping.index = integer(0), stringsAsFactors = FALSE)
  
  for (condition in names(result_list)) {
    extract <- result_list[[condition]] 
    
    extract$difference <- c(NA, diff(extract$`FvFmrunnorm.r`)) #calculate difference in reconstructed FvFm values between successive rows 
    
    cycling.flashes <- which(extract$difference != 0) #extract row index where FvFm is oscillating
    
    damping.index <- max(cycling.flashes) #extract the highest row index (last row before FvFm damps)
    
    result_df <- rbind(result_df, data.frame(condition = condition, damping.index = damping.index))
  }
  return(result_df)
}

# Returns conditions with significant periodicity at 4 
wt.significance <- function(wavelets) {
  result_list <- lapply(names(wavelets), function(condition_name) {
    condition <- wavelets[[condition_name]]
    Power.avg <- condition$Power.avg
    Power.avg.pval <- condition$Power.avg.pval
    Period <- condition$Period
    Condition <- rep(condition_name, length(Power.avg))  
    df <- data.frame(Power.avg, Power.avg.pval, Period, Condition)
    
    return(df)
  })

  result_df <- bind_rows(result_list) %>%
    filter(Power.avg.pval <= siglvl) %>%
    filter(between(Period, 3.9, 4.1))
  
    significantconditions <- unique(result_df$Condition)
  
  return(significantconditions)
}


#plot of wavelet power averages across time, indicates which conditions show a significant periodicity of 4
plot.wt.avg <- function(wavelet_list) {
  
  for (condition in names(wavelet_list)) {
    extract <- wavelet_list[[condition]] 
    
    WaveletComp::wt.avg(extract, siglvl = siglvl,
      sigcol = "#71CF61", 
      maximum.level = 1,
      averagelab = "Average Wavelet Power",
      periodlab = "Period",
      legend.coords = "topright",
      main = condition)}
}

#image plot of the wavelet power spectrum of a time series
plot.wt.image <- function(wavelet_list) {
  
  for (condition in names(wavelet_list)) {
    extract <- wavelet_list[[condition]] 
    
    WaveletComp::wt.image(extract,
         plot.coi = FALSE,
         siglvl = siglvl,
         color.key = "i",
         color.palette = "rainbow(n.levels, start = 0, end = .7)",
         timelab = "Flash Number",
         spec.time.axis = list(at = c(4,8,12,16,20,24,28,32)),
         periodlab = "Period",
         main = condition)}
}
```



# Ottawa_2025 

### Ck
```{r import Ck data }
Ck_data <- readRDS(file.path(DataIn, "LOW_Ottawa_2025_SolFitsMeta.Rds")) |>
  filter(Strain == "Chlamydomonas_klinobasis")

Ck_data <- summaryBy(FvFmrunnorm ~ PulseSpace_s + Temp_C + Flashnumber + LightLevel_ue + NaCl_mM, data=Ck_data, FUN=mean, keep.names = TRUE) %>%
  splitBy(formula = ~ PulseSpace_s + Temp_C + LightLevel_ue + NaCl_mM) %>%
  setNames(., paste0("Ck_", gsub("\\|", "_", names(.))))
```

```{r wavelet transformation Ck}
Ck_wavelets <- wavelet(Ck_data)
```

```{r evaluate significance Ck}
Ck_significant <- wt.significance(Ck_wavelets)
#plot.wt.avg(Ci_wavelets)
```

```{r reconstruct waves Ck}
Ck_reconstruction <- reconstructwave(Ck_wavelets)
```

```{r damping index Ck}
Ck_damping <- damping(Ck_reconstruction)  %>%
  mutate(damping.index = ifelse(!condition %in% Ck_significant, 0, damping.index)) %>%
  rename(DampingIndex = damping.index) %>%
  separate(condition, into=c("Strain", "PulseSpace_s", "Temp_C", "LightLevel_ue", "NaCl_mM"), sep = "_") |>
  mutate(Temp_C = as.numeric(Temp_C),
         PulseSpace_s = as.numeric(PulseSpace_s),
         NaCl_mM = as.numeric(NaCl_mM),
         LightLevel_ue = as.numeric(LightLevel_ue)) |>
  mutate(Strain = Ck, GrowthTemp_C = 4, 
         TempDiff_C = Temp_C - GrowthTemp_C)
  
```

# Ck TestPlot of damping index

```{r Ck_dampingindex}
Ck_damping |>
  ggplot() +
  geom_line(aes(x = PulseSpace_s, y = DampingIndex, colour = as.factor(NaCl_mM))) +
  facet_grid(rows = vars(Temp_C, TempDiff_C), cols = vars(Strain)) +
  theme_bw()
```


### Cm
```{r import Cm data }
Cm_data <- readRDS(file.path(DataIn, "LOW_Ottawa_2025_SolFitsMeta.Rds")) |>
  filter(Strain == "Chlamydomonas_malina")

Cm_data <- summaryBy(FvFmrunnorm ~ PulseSpace_s + Temp_C + Flashnumber + LightLevel_ue + NaCl_mM, data=Cm_data, FUN=mean, keep.names = TRUE) %>%
  splitBy(formula = ~ PulseSpace_s + Temp_C + LightLevel_ue + NaCl_mM) %>%
  setNames(., paste0("Cm_", gsub("\\|", "_", names(.))))
```

```{r wavelet transformation Cm}
Cm_wavelets <- wavelet(Cm_data)
```

```{r evaluate significance Cm}
Cm_significant <- wt.significance(Cm_wavelets)
#plot.wt.avg(Ci_wavelets)
```

```{r reconstruct waves Cm}
Cm_reconstruction <- reconstructwave(Cm_wavelets)
```

```{r damping index Cm}
Cm_damping <- damping(Cm_reconstruction)  %>%
  mutate(damping.index = ifelse(!condition %in% Cm_significant, 0, damping.index)) %>%
  rename(DampingIndex = damping.index) %>%
  separate(condition, into=c("Strain", "PulseSpace_s", "Temp_C", "LightLevel_ue", "NaCl_mM"), sep = "_") |>
  mutate(Temp_C = as.numeric(Temp_C),
         PulseSpace_s = as.numeric(PulseSpace_s),
         NaCl_mM = as.numeric(NaCl_mM),
         LightLevel_ue = as.numeric(LightLevel_ue)) |>
  mutate(Strain = Cm, GrowthTemp_C = 4, 
         TempDiff_C = Temp_C - GrowthTemp_C)
  
```

# Cm TestPlot of damping index

```{r Cm_dampingindex}
Cm_damping |>
  ggplot() +
  geom_line(aes(x = PulseSpace_s, y = DampingIndex, colour = as.factor(NaCl_mM))) +
  facet_grid(rows = vars(Temp_C), cols = vars(Strain)) +
  theme_bw()
```

### Cp
```{r import Cm data }
Cp_data <- readRDS(file.path(DataIn, "LOW_Ottawa_2025_SolFitsMeta.Rds")) |>
  filter(Strain == "Chlamydomonas_priscuii")

Cp_data <- summaryBy(FvFmrunnorm ~ PulseSpace_s + Temp_C + Flashnumber + LightLevel_ue + NaCl_mM, data=Cp_data, FUN=mean, keep.names = TRUE) %>%
  splitBy(formula = ~ PulseSpace_s + Temp_C + LightLevel_ue + NaCl_mM) %>%
  setNames(., paste0("Cp_", gsub("\\|", "_", names(.))))
```

```{r wavelet transformation Cp}
Cp_wavelets <- wavelet(Cp_data)
```

```{r evaluate significance Cp}
Cp_significant <- wt.significance(Cp_wavelets)
#plot.wt.avg(Ci_wavelets)
```

```{r reconstruct waves Cp}
Cp_reconstruction <- reconstructwave(Cp_wavelets)
```

```{r damping index Cp}
Cp_damping <- damping(Cp_reconstruction)  %>%
  mutate(damping.index = ifelse(!condition %in% Cp_significant, 0, damping.index)) %>%
  rename(DampingIndex = damping.index) %>%
  separate(condition, into=c("Strain", "PulseSpace_s", "Temp_C", "LightLevel_ue", "NaCl_mM"), sep = "_") |>
  mutate(Temp_C = as.numeric(Temp_C),
         PulseSpace_s = as.numeric(PulseSpace_s),
         NaCl_mM = as.numeric(NaCl_mM),
         LightLevel_ue = as.numeric(LightLevel_ue)) |>
  mutate(Strain = Cp, GrowthTemp_C = 4, 
         TempDiff_C = Temp_C - GrowthTemp_C)
  
```

# Cp TestPlot of damping index

```{r Cp_dampingindex}
Cp_damping |>
  ggplot() +
  geom_line(aes(x = PulseSpace_s, y = DampingIndex, colour = as.factor(NaCl_mM))) +
  facet_grid(rows = vars(Temp_C, TempDiff_C), cols = vars(Strain)) +
  theme_bw()
```

### Ce
```{r import Cm data }
Ce_data <- readRDS(file.path(DataIn, "LOW_Ottawa_2025_SolFitsMeta.Rds")) |>
  filter(Strain == "Chlamydomonas_euryale")

Ce_data <- summaryBy(FvFmrunnorm ~ PulseSpace_s + Temp_C + Flashnumber + LightLevel_ue + NaCl_mM, data=Ce_data, FUN=mean, keep.names = TRUE) %>%
  splitBy(formula = ~ PulseSpace_s + Temp_C + LightLevel_ue + NaCl_mM) %>%
  setNames(., paste0("Ce_", gsub("\\|", "_", names(.))))
```

```{r wavelet transformation Ce}
Ce_wavelets <- wavelet(Ce_data)
```

```{r evaluate significance Ce}
Ce_significant <- wt.significance(Ce_wavelets)
#plot.wt.avg(Ci_wavelets)
```

```{r reconstruct waves Ce}
Ce_reconstruction <- reconstructwave(Ce_wavelets)
```

```{r damping index Ce}
Ce_damping <- damping(Ce_reconstruction)  %>%
  mutate(damping.index = ifelse(!condition %in% Ce_significant, 0, damping.index)) %>%
  rename(DampingIndex = damping.index) %>%
  separate(condition, into=c("Strain", "PulseSpace_s", "Temp_C", "LightLevel_ue", "NaCl_mM"), sep = "_") |>
  mutate(Temp_C = as.numeric(Temp_C),
         PulseSpace_s = as.numeric(PulseSpace_s),
         NaCl_mM = as.numeric(NaCl_mM),
         LightLevel_ue = as.numeric(LightLevel_ue)) |>
  mutate(Strain = Ce, GrowthTemp_C = 24, 
         TempDiff_C = Temp_C - GrowthTemp_C)
  
```

# Ce TestPlot of damping index

```{r Ce_dampingindex}
Ce_damping |>
  ggplot() +
  geom_line(aes(x = PulseSpace_s, y = DampingIndex, colour = as.factor(NaCl_mM))) +
  facet_grid(rows = vars(Temp_C, TempDiff_C), cols = vars(Strain)) +
  theme_bw()
```

### Cr
```{r import Cm data }
Cr_data <- readRDS(file.path(DataIn, "LOW_Ottawa_2025_SolFitsMeta.Rds")) |>
  filter(Strain == "Chlamydomonas_reinhardtii")

Cr_data <- summaryBy(FvFmrunnorm ~ PulseSpace_s + Temp_C + Flashnumber + LightLevel_ue + NaCl_mM, data=Cr_data, FUN=mean, keep.names = TRUE) %>%
  splitBy(formula = ~ PulseSpace_s + Temp_C + LightLevel_ue + NaCl_mM) %>%
  setNames(., paste0("Cr_", gsub("\\|", "_", names(.))))
```

```{r wavelet transformation Cr}
Cr_wavelets <- wavelet(Cr_data)
```

```{r evaluate significance Cr}
Cr_significant <- wt.significance(Cr_wavelets)
#plot.wt.avg(Ci_wavelets)
```

```{r reconstruct waves Cr}
Cr_reconstruction <- reconstructwave(Cr_wavelets)
```

```{r damping index Cr}
Cr_damping <- damping(Cr_reconstruction)  %>%
  mutate(damping.index = ifelse(!condition %in% Cr_significant, 0, damping.index)) %>%
  rename(DampingIndex = damping.index) %>%
  separate(condition, into=c("Strain", "PulseSpace_s", "Temp_C", "LightLevel_ue", "NaCl_mM"), sep = "_") |>
  mutate(Temp_C = as.numeric(Temp_C),
         PulseSpace_s = as.numeric(PulseSpace_s),
         NaCl_mM = as.numeric(NaCl_mM),
         LightLevel_ue = as.numeric(LightLevel_ue)) |>
  mutate(Strain = Cr, GrowthTemp_C = 24, 
         TempDiff_C = Temp_C - GrowthTemp_C)
  
```

# Cr TestPlot of damping index

```{r Cr_dampingindex}
Cr_damping |>
  ggplot() +
  geom_line(aes(x = PulseSpace_s, y = DampingIndex, colour = as.factor(NaCl_mM))) +
  facet_grid(rows = vars(Temp_C), cols = vars(Strain)) +
  theme_bw()
```

# Merging Results 

```{r merge}
sstate_damping <- rbind(Ck_damping, Ce_damping, Cm_damping, Cp_damping, Cr_damping) %>%
    type_convert() 


saveRDS(sstate_damping, file.path(DataOut, paste(Project, "SState_Damping_2025.RDS", sep = "_"), fsep = .Platform$file.sep))
```

```{r}
wavelets <- c(Ck_wavelets, Ce_wavelets, Cm_wavelets, Cp_wavelets, Cr_wavelets)

saveRDS(wavelets, file.path(DataOut, paste(Project, "Wavelets_2025.RDS", sep = "_"), fsep = .Platform$file.sep))
```

```{r}
reconstructions <- c(Ck_reconstruction, Ce_reconstruction, Cm_reconstruction, Cp_reconstruction, Cr_reconstruction)

saveRDS(reconstructions, file.path(DataOut, paste(Project, "Reconstructions_2025.RDS", sep = "_"), fsep = .Platform$file.sep))
```

# Figures 


```{r heat map}
#, labeller = labeller(Species=species.labs))

sstate_damping %>%
  ggplot() +
  geom_tile(aes(x = Temp_C, y = PulseSpace_s, fill = DampingIndex)) +
  viridis::scale_fill_viridis(option="mako", na.value="white") +
  facet_wrap(vars(Strain), ncol=2, scales = "free") +
  labs(x= "Measurement Temperature (°C)", y= "Spacing of Photon Delivery (s)", fill= "Duration of PSII Cycling") +
  theme_linedraw() +
  theme(text = element_text(family = "Palatino"),
        strip.background =element_rect(fill="white"),
        strip.text = element_text(colour= "black",face = "italic", size = 13),
        axis.text=element_text(size=13), axis.title=element_text(size=14),
        legend.text=element_text(size=13), legend.title=element_text(size=14),
        legend.position = c(0.75, 0.1),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

